### 2019-05-01 docs/index.Rmd

---
title: "Poseidon Adventure"
output: html_document
---

 <!-- ADD LINKS BELOW, note, using the RMD link syntax will open the presentation in the same tab, which is annoying. Use the html with the "_blank" to fix that behavior -->

These pages document initial testing of the atlantisom package in development at https://github.com/r4atlantis/atlantisom using three different [Atlantis](https://research.csiro.au/atlantis/) output datasets. Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.

### Testing the atlantisom code

<a href="https://sgaichas.github.io/poseidon-dev/TrueBioTest.html" target="_blank">Installation and initial tests - April 26, 2019, updated April 30, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NEUSExamples.html" target="_blank">Survey census visualization for NEUS -  April 30, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCAExamples.html" target="_blank">Survey census visualization for CCA - April 30, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NOBAExamples.html" target="_blank">Survey census visualization for NOBA - April 30, 2019</a>

### Create a standard survey

<a href="https://sgaichas.github.io/poseidon-dev/StandardSurveyTest.html" target="_blank">Initial survey tests - April 30, 2019, updated May 1, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NEUSStandardSurveyTest.html" target="_blank">Standard survey visualization for NEUS -  May 1, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCAStandardSurveyTest.html" target="_blank">Standard survey visualization for CCA - May 1, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NOBAStandardSurveyTest.html" target="_blank">Standard survey visualization for NOBA - May 1, 2019</a>

### More to come later!