### 2019-05-02 TrueLengthAgeCompTest.Rmd
age comp sampling up to broken function

---
title: "Testing atlantisom: generate census length and age comps"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'knitr', 'rmarkdown', 'tidyr', 'dplyr', 'ggplot2',
  'data.table', 'here', 'ggforce', 'ggthemes'
), 'packages.bib')
```

## Introduction

This page documents initial testing of the atlantisom package in development at https://github.com/r4atlantis/atlantisom using three different [Atlantis](https://research.csiro.au/atlantis/) output datasets. Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.

The purpose of atlantisom is to use existing Atlantis model output to generate input datasets for a variety of models, so that the performance of these models can be evaluated against known (simulated) ecosystem dynamics. Atlantis models can be run using different climate forcing, fishing, and other scenarios. Users of atlantisom will be able to specify fishery independent and fishery dependent sampling in space and time, as well as species-specific catchability, selectivty, and other observation processes for any Atlantis scenario. Internally consistent multispecies and ecosystem datasets with known observation error characteristics will be the atlantisom outputs, for use in individual model performance testing, comparing performance of alternative models, and performance testing of model ensembles against "true" Atlantis outputs.

Initial testing was conducted by S. Gaichas using R scripts in the R folder of this repository that are titled "PoseidonTest_[whatwastested].R". Initial tests are expanded and documented in more detail in these pages. C. Stawitz improved and streamlined the setup and intialization sections.

## Setup

First, you will want to set up libraries and install atlantisom if you haven't already. This document assumes atlantisom is already installed. For complete setup and initialization, please see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).

This document is written in in R Markdown [@R-rmarkdown], and we use several packages to produce the outputs [@R-tidyr; @R-dplyr; @R-ggplot2; @R-here; @R-ggforce; @R-ggthemes].

```{r message=FALSE, warning=FALSE}
library(tidyr)
require(dplyr)
library(ggplot2)
library(data.table)
library(here)
library(ggforce)
library(ggthemes)
library(atlantisom)
```

## Initialize input files and directories, read in "truth"

Abbreviated here; for a full explanation please see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html). This document assumes that `atlantisom::run_truth` has already completed and stored an .RData file in the atlantis output model directory.

```{r initialize}

initCCA <- FALSE
initNEUS <- TRUE
initNOBA <- FALSE

if(initCCA){
  d.name <- here("atlantisoutput","CalCurrentSummitScenario1")
  functional.groups.file <- "CalCurrentV3Groups.csv"
  biomass.pools.file <- "DIVCalCurrentV3_BIOL.nc"
  biol.prm.file <- "CalCurrentV3_Biol.prm"
  box.file <- "CalCurrentV3_utm.bgm"
  initial.conditions.file <- "DIVCalCurrentV3_BIOL.nc"
  run.prm.file <- "CalCurrentV3_run.xml"
  scenario.name <- "CCV3"
}

if(initNEUS){
  d.name <- here("atlantisoutput","NEUStest20160303")
  functional.groups.file <- "NeusGroups.csv" 
  biomass.pools.file <- ""
  biol.prm.file <- "at_biol_neus_v15_DE.prm"
  box.file <- "neus30_2006.bgm"
  initial.conditions.file <- "inneus_2012.nc"
  run.prm.file <- "at_run_neus_v15_DE.xml"
  scenario.name <- "neusDynEffort_Test1_"
}

if(initNOBA){
  d.name <- here("atlantisoutput","NOBACERESGlobalSustainability")
  functional.groups.file <- "nordic_groups_v04.csv" 
  biomass.pools.file <- "nordic_biol_v23.nc"
  biol.prm.file <- "nordic_biol_incl_harv_v_007_3.prm"
  box.file <- "Nordic02.bgm"
  initial.conditions.file <- "nordic_biol_v23.nc"
  run.prm.file <- "nordic_run_v01.xml"
  scenario.name <- "nordic_runresults_01"
}
# NOBA note: output filenames in CCA and NEUS begin with "output" and the run_truth function is written to expect this. Need to check if default Atlantis output file nomenclature has changed or if NOBA is a special case. For now, NOBA filenames have been changed to include prefix "output"
```

```{r get_names, message=FALSE, warning=FALSE}
#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

```

```{r load_Rdata, message=FALSE, warning=FALSE}

if(initCCA) {
  d.name <- here("atlantisoutput","CalCurrentSummitScenario1")
  truth.file <- "outputCCV3run_truth.RData"
  load(file.path(d.name, truth.file))
  CCAresults <- result
}

if(initNEUS) {
  d.name <- here("atlantisoutput","NEUStest20160303")
  truth.file <- "outputneusDynEffort_Test1_run_truth.RData" 
  load(file.path(d.name, truth.file))
  NEUSresults <- result
}

if(initNOBA){
  d.name <- here("atlantisoutput","NOBACERESGlobalSustainability")
  truth.file <- "outputnordic_runresults_01run_truth.RData" 
  load(file.path(d.name, truth.file))
  NOBAresults <- result
}

```

We can also read in previously generated survey census files based on true biomass results for comparison if necessary, but not yet used here (see [TrueBioTest](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html).)

```{r readsurvey1, eval=FALSE}

if(initCCA) {
  CCAsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
}

if(initNEUS) {
  NEUSsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
}

if(initNOBA) {
  NOBAsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
}

```

## Simulate a survey part 3: sample for length and age composition

This section uses the `atlantisom::create_survey()` and `atlantisom::sample_fish()` to get a biological sample dataset. From that dataset, several other functions can be run:

 *  age samples from sample_ages
 *  length samples from sample_lengths (TO BE WRITTEN)
 *  weight samples from sample_weights (TO BE WRITTEN)
 *  diet samples from sample_diet

To create a survey, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species.

The following settings should achieve a survey that samples all Atlantis model output timesteps, all fish and shark species, and all model polygons, with perfect efficiency and full selectivity for all ages:

```{r census-spec, message=FALSE, warning=FALSE}

# should return a perfectly scaled survey 
effic1 <- data.frame(species=funct.group.names,
                     efficiency=rep(1.0,length(funct.group.names)))

# should return all lengths fully sampled (Atlantis output is 10 age groups per spp)
# BUT CHECK if newer Atlantis models can do age-specific outputs
selex1 <- data.frame(species=rep(funct.group.names, each=10),
                     agecl=rep(c(1:10),length(funct.group.names)),
                     selex=rep(1.0,length(funct.group.names)*10))

# should return all model areas
boxpars <- load_box(d.name, box.file)
boxall <- c(0:(boxpars$nbox - 1))

# these are model specific, generalized above
# if(initCCA) boxall <- c(0:88) 
# if(initNEUS) boxall <- c(0:29)
# if(initNOBA) boxall <- c(0:59)

# should return all model output timesteps; need to generalize
# if(initCCA) timeall <- c(0:100) 
# if(initNEUS) timeall <- c(0:251)
# if(initNOBA) timeall <- c(0:560)

# generalized
runpar <- load_runprm(d.name, run.prm.file)
noutsteps <- runpar$tstop/runpar$outputstep
stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc

timeall <- c(0:noutsteps)

# define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
# for ecosystem indicator work test all species, e.g.
survspp <- funct.group.names

# to keep plots simpler, currently hardcoded for vertebrate/fished invert groups
# if(initCCA) survspp <- funct.group.names[c(1:44, 59:61, 65:68)] 
# if(initNEUS) survspp <- funct.group.names[1:21]
# if(initNOBA) survspp <- funct.group.names[1:36]

# for length and age groups lets just do fish and sharks
# NOBA model has InvertType, changed to GroupType in file, but check Atlantis default
if(initNOBA) funct.groups <- rename(funct.groups, GroupType = InvertType)

survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
                           funct.groups$GroupType %in% c("FISH", "SHARK")]

```

Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age and length composition.

```{r stdsurveyNbased}

# this uses result$nums, but we are not creating a biomass index this time, so we don't need a weight at age conversion

if(initCCA) datN <- CCAresults$nums
if(initNEUS) datN <- NEUSresults$nums
if(initNOBA) datN <- NOBAresults$nums

survey_testNall <- create_survey(dat = datN,
                                 time = timeall,
                                 species = survspp,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

# consider saving this interim step if it takes a long time go generate

```

Next, get true annual (cohort) age comp from this census survey based on run truth. (is there a standard Atlantis output I can compare this to as we did for biomass?)

```{r truecohortagecomp}
# what is true composition? need annual by species, use code from sample_fish
# do tidyly
dat2 <- survey_testNall %>%
  group_by(species, agecl, time) %>%
  summarize(numAtAge = sum(atoutput))

#dat<-survey_testNall
#dat2 <- aggregate(dat$atoutput,list(dat$species,dat$agecl,dat$time),sum)
#names(dat2) <- c("species","agecl","time","numAtAge")

totN <- dat2 %>%
  group_by(species, time) %>%
  summarize(totN = sum(numAtAge))

#totN <- aggregate(dat2$numAtAge,list(dat2$species,dat2$time),sum )
#names(totN) <- c("species","time","totN")

dat2totN <- merge(dat2, totN)

ageclcomp <- ggplot(dat2totN, aes(x=agecl, y=numAtAge/totN, col=time)) +
  geom_point()

ageclcomp + facet_wrap(~species, ncol=5) #, scales="free")

```

Then we use the `sample_fish` function and compare to true annual age comp calculated above as a test. These should match.

```{r comptest1, message=FALSE, warning=FALSE}
# setting the effN higher than actual numbers results in sampling all
effNall <- data.frame(species=survspp, effN=rep(1e+15, length(survspp)))
# rmultinom broke with that sample size
#comptestall <- sample_fish(survey_testNall, effNall)
#names(comptestall) <- c("species","agecl","polygon", "layer","time","numAtAgesamp")

# this one is high but not equal to total for numerous groups
effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))

comptesthigh <- sample_fish(survey_testNall, effNhigh)
names(comptesthigh) <- c("species","agecl","polygon", "layer","time","numAtAgesamp")

comptesttot <- aggregate(comptesthigh$numAtAgesamp,list(comptesthigh$species,comptesthigh$time),sum )
names(comptesttot) <- c("species","time","totsamp")

comptestprop <- merge(comptesthigh, comptesttot)

# compare individual years, these proportions at age should match
comparecomps <- ggplot() +
  geom_point(data=subset(dat2totN, time==0), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
  geom_point(data=subset(comptestprop, time==0), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

comparecomps + facet_wrap(~species, ncol=5)

#ggsave("censuscomposition_time0.png", width=11, height=11)

comparecomps <- ggplot() +
  geom_point(data=subset(dat2totN, time==100), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
  geom_point(data=subset(comptestprop, time==100), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3)+
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

comparecomps + facet_wrap(~species, ncol=5)

#ggsave("censuscomposition_time100.png", width=11, height=11)

comparecomps <- ggplot() +
  geom_point(data=subset(dat2totN, time==250), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
  geom_point(data=subset(comptestprop, time==250), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3)+
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

comparecomps + facet_wrap(~species, ncol=5)

#ggsave("censuscomposition_time250.png", width=11, height=11)

# and they do
# differences for baleen and toothed whales are due to small total numbers

```

And they did for NEUS the first time I did this, with some exceptions.

The problems start when I try to do age to length, this function returns an error at present.

**Editing to be done in atlantisom (I have made local changes to calcage2length.R that are not pushed yet because I may have introduced an error)**

```{r calcage2length-breaks, eval=FALSE}
# now turn this into length comp with calc_age2length
# function takes resn and nums as inputs so run survey and sample functions on resn as well
if(initCCA) datresn <- CCAresults$resn
if(initNEUS) datresn <- NEUSresults$resn
if(initNOBA) datresn <- NOBAresults$resn

survey_testresnall <- create_survey(dat = datresn,
                                 time = seq(0,251,5), # use timeall? census?
                                 species = survspp,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

lengthtesthigh <- sample_fish(survey_testNall, effNhigh)
# use fish sampled from N along with survey of resn in function
# calc_age2length uses nums (lengthtest) and finds matches in resn

# provides ouput for an individual species and time slice
# breaks for whole dataset (NA output all rows)

if(initCCA) truth <- CCAresults
if(initNEUS) truth <- NEUSresults
if(initNOBA) truth <- NOBAresults

length_census <- calc_age2length(structn = lengthtesthigh,
                                 resn = survey_testresnall,
                                 nums = lengthtesthigh,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

```

## References