### 2019-05-06 visualization_Rmd/CCATrueLengthAgeCompTest.Rmd
age comp census visualizations all mods

---
title: "Testing atlantisom: generate census length and age comps from California Current Atlantis"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'knitr', 'rmarkdown', 'tidyr', 'dplyr', 'ggplot2',
  'data.table', 'here', 'ggforce', 'ggthemes'
), 'packages.bib')
```

## Introduction

This page has visualizations for the CCA model example, Atlantis Summit Common Scenario 1. For full explanation of methods, see the file linked at the beginning of each section.

```{r message=FALSE, warning=FALSE}
library(tidyr)
require(dplyr)
library(ggplot2)
library(data.table)
library(here)
library(ggforce)
library(ggthemes)
library(atlantisom)
```

```{r initialize}

initCCA <- TRUE
initNEUS <- FALSE
initNOBA <- FALSE

if(initCCA){
  d.name <- here("atlantisoutput","CalCurrentSummitScenario1")
  functional.groups.file <- "CalCurrentV3Groups.csv"
  biomass.pools.file <- "DIVCalCurrentV3_BIOL.nc"
  biol.prm.file <- "CalCurrentV3_Biol.prm"
  box.file <- "CalCurrentV3_utm.bgm"
  initial.conditions.file <- "DIVCalCurrentV3_BIOL.nc"
  run.prm.file <- "CalCurrentV3_run.xml"
  scenario.name <- "CCV3"
}

if(initNEUS){
  d.name <- here("atlantisoutput","NEUStest20160303")
  functional.groups.file <- "NeusGroups.csv" 
  biomass.pools.file <- ""
  biol.prm.file <- "at_biol_neus_v15_DE.prm"
  box.file <- "neus30_2006.bgm"
  initial.conditions.file <- "inneus_2012.nc"
  run.prm.file <- "at_run_neus_v15_DE.xml"
  scenario.name <- "neusDynEffort_Test1_"
}

if(initNOBA){
  d.name <- here("atlantisoutput","NOBACERESGlobalSustainability")
  functional.groups.file <- "nordic_groups_v04.csv" 
  biomass.pools.file <- "nordic_biol_v23.nc"
  biol.prm.file <- "nordic_biol_incl_harv_v_007_3.prm"
  box.file <- "Nordic02.bgm"
  initial.conditions.file <- "nordic_biol_v23.nc"
  run.prm.file <- "nordic_run_v01.xml"
  scenario.name <- "nordic_runresults_01"
}
# NOBA note: output filenames in CCA and NEUS begin with "output" and the run_truth function is written to expect this. Need to check if default Atlantis output file nomenclature has changed or if NOBA is a special case. For now, NOBA filenames have been changed to include prefix "output"
```

```{r get_names, message=FALSE, warning=FALSE}
#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

```

```{r load_Rdata, message=FALSE, warning=FALSE}

if(initCCA) {
  d.name <- here("atlantisoutput","CalCurrentSummitScenario1")
  truth.file <- "outputCCV3run_truth.RData"
  load(file.path(d.name, truth.file))
  CCAresults <- result
}

if(initNEUS) {
  d.name <- here("atlantisoutput","NEUStest20160303")
  truth.file <- "outputneusDynEffort_Test1_run_truth.RData" 
  load(file.path(d.name, truth.file))
  NEUSresults <- result
}

if(initNOBA){
  d.name <- here("atlantisoutput","NOBACERESGlobalSustainability")
  truth.file <- "outputnordic_runresults_01run_truth.RData" 
  load(file.path(d.name, truth.file))
  NOBAresults <- result
}

```

```{r readsurvey1, eval=FALSE}

if(initCCA) {
  CCAsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
}

if(initNEUS) {
  NEUSsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
}

if(initNOBA) {
  NOBAsurveyB_frombio <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))
}

```

## Simulate a survey part 3: sample for length and age composition

Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/TrueLengthAgeCompTest.html).

The following settings should achieve a survey that samples all Atlantis model output timesteps, all fish and shark species, and all model polygons, with perfect efficiency and full selectivity for all ages:

```{r census-spec, message=FALSE, warning=FALSE, echo=TRUE}

# should return a perfectly scaled survey 
effic1 <- data.frame(species=funct.group.names,
                     efficiency=rep(1.0,length(funct.group.names)))

# should return all lengths fully sampled (Atlantis output is 10 age groups per spp)
# BUT CHECK if newer Atlantis models can do age-specific outputs
selex1 <- data.frame(species=rep(funct.group.names, each=10),
                     agecl=rep(c(1:10),length(funct.group.names)),
                     selex=rep(1.0,length(funct.group.names)*10))

# should return all model areas
boxpars <- load_box(d.name, box.file)
boxall <- c(0:(boxpars$nbox - 1))

# these are model specific, generalized above
# if(initCCA) boxall <- c(0:88) 
# if(initNEUS) boxall <- c(0:29)
# if(initNOBA) boxall <- c(0:59)

# should return all model output timesteps; need to generalize
# if(initCCA) timeall <- c(0:100) 
# if(initNEUS) timeall <- c(0:251)
# if(initNOBA) timeall <- c(0:560)

# generalized
runpar <- load_runprm(d.name, run.prm.file)
noutsteps <- runpar$tstop/runpar$outputstep
stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc

timeall <- c(0:noutsteps)

# define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
# for ecosystem indicator work test all species, e.g.
survspp <- funct.group.names

# to keep plots simpler, currently hardcoded for vertebrate/fished invert groups
# if(initCCA) survspp <- funct.group.names[c(1:44, 59:61, 65:68)] 
# if(initNEUS) survspp <- funct.group.names[1:21]
# if(initNOBA) survspp <- funct.group.names[1:36]

# for length and age groups lets just do fish and sharks
# NOBA model has InvertType, changed to GroupType in file, but check Atlantis default
if(initNOBA) funct.groups <- rename(funct.groups, GroupType = InvertType)

survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
                           funct.groups$GroupType %in% c("FISH", "SHARK")]

#if(initCCA) survspp <- survspp[!survspp %in% "Pisciv_T_Fish"]

```

Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age and length composition.

```{r stdsurveyNbased}

# this uses result$nums, but we are not creating a biomass index this time, so we don't need a weight at age conversion

if(initCCA) datN <- CCAresults$nums
if(initNEUS) datN <- NEUSresults$nums
if(initNOBA) datN <- NOBAresults$nums

survey_testNall <- create_survey(dat = datN,
                                 time = timeall,
                                 species = survspp,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

# consider saving this interim step if it takes a long time go generate

```

Next, get true annual (cohort) age comp from this census survey based on run truth. (is there a standard Atlantis output I can compare this to as we did for biomass?)

```{r truecohortagecomp}
# what is true composition? need annual by species, use code from sample_fish
# do tidyly
dat2 <- survey_testNall %>%
  group_by(species, agecl, time) %>%
  summarize(numAtAge = sum(atoutput))

#dat<-survey_testNall
#dat2 <- aggregate(dat$atoutput,list(dat$species,dat$agecl,dat$time),sum)
#names(dat2) <- c("species","agecl","time","numAtAge")

totN <- dat2 %>%
  group_by(species, time) %>%
  summarize(totN = sum(numAtAge))

#totN <- aggregate(dat2$numAtAge,list(dat2$species,dat2$time),sum )
#names(totN) <- c("species","time","totN")

dat2totN <- merge(dat2, totN)

ageclcomp <- ggplot(dat2totN, aes(x=agecl, y=numAtAge/totN, col=time)) +
  geom_point()

ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
#ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")

```

Then we use the `sample_fish` function and compare to true annual age comp calculated above as a test. These should match. We show compositions sampled for the initial time step, a time step in the middle of the run, and the last timestep.

```{r comptest1, message=FALSE, warning=FALSE}
# setting the effN higher than actual numbers results in sampling all
effNall <- data.frame(species=survspp, effN=rep(1e+15, length(survspp)))
# rmultinom broke with that sample size
#comptestall <- sample_fish(survey_testNall, effNall)
#names(comptestall) <- c("species","agecl","polygon", "layer","time","numAtAgesamp")

# this one is high but not equal to total for numerous groups
effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))

comptesthigh <- sample_fish(survey_testNall, effNhigh)
names(comptesthigh) <- c("species","agecl","polygon", "layer","time","numAtAgesamp")

comptesttot <- aggregate(comptesthigh$numAtAgesamp,list(comptesthigh$species,comptesthigh$time),sum )
names(comptesttot) <- c("species","time","totsamp")

comptestprop <- merge(comptesthigh, comptesttot)
```

```{r comptest1plot, message=FALSE, warning=FALSE}

# compare individual years, these proportions at age should match
comparecomps <- ggplot() +
  geom_point(data=subset(dat2totN, time==min(timeall)), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
  geom_point(data=subset(comptestprop, time==min(timeall)), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=paste0(scenario.name, " start"))

comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
#comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")

#ggsave("censuscomposition_time0.png", width=11, height=11)

comparecomps <- ggplot() +
  geom_point(data=subset(dat2totN, time==median(timeall)), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
  geom_point(data=subset(comptestprop, time==median(timeall)), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3)+
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=paste0(scenario.name, " midpoint"))

comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
#comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")

#ggsave("censuscomposition_time100.png", width=11, height=11)

comparecomps <- ggplot() +
  geom_point(data=subset(dat2totN, time==max(timeall)-1), aes(x=agecl, y=numAtAge/totN, color="true"), alpha = 0.3) +
  geom_point(data=subset(comptestprop, time==max(timeall)-1), aes(x=agecl, y=numAtAgesamp/totsamp, color="samp"), alpha = 0.3)+
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=paste0(scenario.name, " end"))

comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")
#comparecomps + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 5, scales="free")

#ggsave("censuscomposition_time250.png", width=11, height=11)

# and they do
# differences for baleen and toothed whales are due to small total numbers

```

Most match, with some exceptions to be discussed with the regional model developers. Mismatches seem to happen later in this CCA run?