### 2019-05-06 docs/index.Rmd
add comp testing to index

---
title: "Poseidon Adventure"
output: html_document
---

 <!-- ADD LINKS BELOW, note, using the RMD link syntax will open the presentation in the same tab, which is annoying. Use the html with the "_blank" to fix that behavior -->

These pages document initial testing of the atlantisom package in development at https://github.com/r4atlantis/atlantisom using three different [Atlantis](https://research.csiro.au/atlantis/) output datasets. Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.

### Testing the atlantisom code

<a href="https://sgaichas.github.io/poseidon-dev/TrueBioTest.html" target="_blank">Installation and initial tests - April 26, 2019, updated April 30, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NEUSExamples.html" target="_blank">Survey census visualization for NEUS -  April 30, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCAExamples.html" target="_blank">Survey census visualization for CCA - April 30, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NOBAExamples.html" target="_blank">Survey census visualization for NOBA - April 30, 2019</a>

### Create a standard survey

<a href="https://sgaichas.github.io/poseidon-dev/StandardSurveyTest.html" target="_blank">Initial survey tests - April 30, 2019, updated May 1, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NEUSStandardSurveyTest.html" target="_blank">Standard survey visualization for NEUS -  May 1, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCAStandardSurveyTest.html" target="_blank">Standard survey visualization for CCA - May 1, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NOBAStandardSurveyTest.html" target="_blank">Standard survey visualization for NOBA - May 1, 2019</a>

### Start age composition sampling: work in progress

<a href="https://sgaichas.github.io/poseidon-dev/TrueLengthAgeCompTest.html" target="_blank">Initial age comp tests - May 2, 2019, updated May 6, 2019</a>

CCA outputs had some 0 length at age, requiring changes to `atlantisom::sample_fish`

<a href="https://sgaichas.github.io/poseidon-dev/NEUSTrueLengthAgeCompTest.html" target="_blank">True and sample fish census comparison NEUS - May 6, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCATrueLengthAgeCompTest.html" target="_blank">True and sample fish census comparison CCA - May 6, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NOBATrueLengthAgeCompTest.html" target="_blank">True and sample fish census comparison NOBA - May 6, 2019</a>

Standard survey with composition sampling: impact of selectivity, `sample_fish`

<a href="https://sgaichas.github.io/poseidon-dev/StdSurvLengthAgeCompTest.html" target="_blank">Compare true cohort comp with survey + sampling - May 6, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NEUSStdSurvCohortCompTest.html" target="_blank">True cohort comp vs survey comparison NEUS - May 6, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCAStdSurvCohortCompTest.html" target="_blank">True cohort comp vs survey comparison CCA - May 6, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NOBAStdSurvCohortCompTest.html" target="_blank">True cohort comp vs survey comparison NOBA - May 6, 2019</a>

### More to come later!