### 2019-05-14 docs/index.Rmd
add lf from standard survey

---
title: "Poseidon Adventure"
output: html_document
---

 <!-- ADD LINKS BELOW, note, using the RMD link syntax will open the presentation in the same tab, which is annoying. Use the html with the "_blank" to fix that behavior -->

These pages document initial testing of the atlantisom package in development at https://github.com/r4atlantis/atlantisom using three different [Atlantis](https://research.csiro.au/atlantis/) output datasets. Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.

### Read this first on overall workflow

<a href="https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html" target="_blank">Big picture and new functions needed - May 10, 2019</a>

May 11: Changes to `atlantisom::sample_fish` and new `atlantisom::aggregateDensityData` functions seem to work. See below under [Length compositions].

### Testing the atlantisom code

<a href="https://sgaichas.github.io/poseidon-dev/TrueBioTest.html" target="_blank">Installation and initial tests - April 26, 2019, updated April 30, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NEUSExamples.html" target="_blank">Survey census visualization for NEUS -  April 30, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCAExamples.html" target="_blank">Survey census visualization for CCA - April 30, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NOBAExamples.html" target="_blank">Survey census visualization for NOBA - April 30, 2019</a>

### Create a standard survey

<a href="https://sgaichas.github.io/poseidon-dev/StandardSurveyTest.html" target="_blank">Initial survey tests - April 30, 2019, updated May 1, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NEUSStandardSurveyTest.html" target="_blank">Standard survey visualization for NEUS -  May 1, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCAStandardSurveyTest.html" target="_blank">Standard survey visualization for CCA - May 1, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NOBAStandardSurveyTest.html" target="_blank">Standard survey visualization for NOBA - May 1, 2019</a>

### Start age composition sampling: work in progress

<a href="https://sgaichas.github.io/poseidon-dev/TrueLengthAgeCompTest.html" target="_blank">Initial age comp tests - May 2, 2019, updated May 6, 2019</a>

CCA outputs had some 0 length at age, requiring changes to `atlantisom::sample_fish`

<a href="https://sgaichas.github.io/poseidon-dev/NEUSTrueLengthAgeCompTest.html" target="_blank">True and sample fish census comparison NEUS - May 6, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCATrueLengthAgeCompTest.html" target="_blank">True and sample fish census comparison CCA - May 6, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NOBATrueLengthAgeCompTest.html" target="_blank">True and sample fish census comparison NOBA - May 6, 2019</a>

Standard survey with composition sampling: impact of selectivity, `sample_fish`

<a href="https://sgaichas.github.io/poseidon-dev/StdSurvLengthAgeCompTest.html" target="_blank">Compare true cohort comp with survey + sampling - May 6, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NEUSStdSurvCohortCompTest.html" target="_blank">True cohort comp vs survey comparison NEUS - May 6, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCAStdSurvCohortCompTest.html" target="_blank">True cohort comp vs survey comparison CCA - May 6, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NOBAStdSurvCohortCompTest.html" target="_blank">True cohort comp vs survey comparison NOBA - May 6, 2019</a>

### Length compositions

Difficult when you don't have a good idea of the outputs you are using!

<a href="https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest_fail.html" target="_blank">Defeated by length compositions - May 8-9, 2019</a>

Better when you do:

<a href="https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html" target="_blank">Victory: length compositions - May 11, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCATrueLengthCompTest.html" target="_blank">CCA length composition examples - May 11, 2019</a>

But don't try running length comps at the subannual level for all species.

<a href="https://sgaichas.github.io/poseidon-dev/NOBATrueLengthCompTest.html" target="_blank">NOBA length composition examples - May 13, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NEUSTrueLengthCompTest.html" target="_blank">NEUS length composition examples - May 13, 2019</a>

Standard survey with length sampling: impact of selectivity, `sample_fish`

<a href="https://sgaichas.github.io/poseidon-dev/StdSurvLengthCompTest.html" target="_blank">Compare survey + sampled lengths with true - May 14, 2019</a>

### More to come later!