### 2019-05-15 TestSmallestLength.Rmd
looking for small turbot

---
title: "Testing atlantisom: do we ever get small fish in length comps?"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'knitr', 'rmarkdown', 'tidyr', 'dplyr', 'ggplot2',
  'data.table', 'here', 'ggforce', 'ggthemes'
), 'packages.bib')
```

## Introduction

Here we further test the function `calc_age2length` to see if small fish ever appear with full census sampling. If they don't, perhaps the default CV in the function needs modification? This question is triggered by the [sampling for NOBA Greenland halibut](https://sgaichas.github.io/poseidon-dev/NOBAStdSurvLengthCompTest.html), in which fish below 50 cm didn't show up.

For all setup, etc, please see previous files Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html) and  [here](https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html).

This page has visualizations for the NOBA model example, CERES Global Sustainability. For full explanation of methods, see the file linked at the beginning of each section.

```{r message=FALSE, warning=FALSE}
library(tidyr)
require(dplyr)
library(ggplot2)
library(data.table)
library(here)
library(ggforce)
library(ggthemes)
library(atlantisom)
```

```{r initialize}

initCCA <- FALSE
initNEUS <- FALSE
initNOBA <- TRUE

if(initCCA) source(here("config/CCConfig.R"))

if(initNEUS) source(here("config/NEUSConfig.R"))

if(initNOBA) source(here("config/NOBAConfig.R"))

```

```{r get_names, message=FALSE, warning=FALSE}
#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

```

```{r load_Rdata, message=FALSE, warning=FALSE}

if(initCCA) {
  truth.file <- "outputCCV3run_truth.RData"
  load(file.path(d.name, truth.file))
  truth <- result
}

if(initNEUS) {
  truth.file <- "outputneusDynEffort_Test1_run_truth.RData" 
  load(file.path(d.name, truth.file))
  truth <- result
}

if(initNOBA){
  truth.file <- "outputnordic_runresults_01run_truth.RData" 
  load(file.path(d.name, truth.file))
  truth <- result
}

```

## Simulate a survey part 4: sample for length composition (testing revised function)

Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/StdSurvLengthCompTest.html).

We will apply examples here to only one species, Greenland halibut in NOBA, which grows to a large size.

To create a census, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species. The following settings should achieve a survey that samples all Atlantis model output timesteps, all fish and shark species, and all model polygons, with perfect efficiency and full selectivity for all ages.

```{r census-spec, message=FALSE, warning=FALSE}

# should return a perfectly scaled survey 
effic1 <- data.frame(species=funct.group.names,
                     efficiency=rep(1.0,length(funct.group.names)))

# should return all lengths fully sampled (Atlantis output is 10 age groups per spp)
selex1 <- data.frame(species=rep(funct.group.names, each=10),
                     agecl=rep(c(1:10),length(funct.group.names)),
                     selex=rep(1.0,length(funct.group.names)*10))

# should return all model areas
boxpars <- load_box(d.name, box.file)
boxall <- c(0:(boxpars$nbox - 1))

# generalized timesteps all models
runpar <- load_runprm(d.name, run.prm.file)
noutsteps <- runpar$tstop/runpar$outputstep
stepperyr <- if(runpar$outputstepunit=="days") 365/runpar$toutinc

timeall <- c(0:noutsteps)

# define set of species we expect surveys to sample (e.g. fish only? vertebrates?)
# for ecosystem indicator work test all species, e.g.
survspp <- funct.group.names

# for length and age groups lets just do fish and sharks
# NOBA model has InvertType, changed to GroupType in file, but check Atlantis default
if(initNOBA) funct.groups <- rename(funct.groups, GroupType = InvertType)

survspp <- funct.groups$Name[funct.groups$IsTurnedOn==1 &
                           funct.groups$GroupType %in% c("FISH", "SHARK")]

```

Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition (for just one species in this case). The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call.

Because we don't want to wait 24 hours for this, we will look at only the first 112 time steps.

```{r stdsurveyNbased-GHR, echo=TRUE}

ss.name <- funct.group.names[funct.group.names == "Green_halibut"]

# get survey nums with full (no) selectivity
ss_survey_testNall <- create_survey(dat = truth$nums,
                                 time = c(0:111),
                                 species = ss.name,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

# this one is high but not equal to total for numerous groups
effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))

# apply default sample fish as before to get numbers
ss_numsallhigh <- sample_fish(ss_survey_testNall, effNhigh)

# aggregate true resn per survey design
ss_aggresnall <- aggregateDensityData(dat = truth$resn,
                                 time = c(0:111),
                                 species = ss.name,
                                 boxes = boxall)

# aggregate true structn per survey design
ss_aggstructnall <- aggregateDensityData(dat = truth$structn,
                                 time = c(0:111),
                                 species = ss.name,
                                 boxes = boxall)

#dont sample these, just aggregate them using median
ss_structnall <- sample_fish(ss_aggstructnall, effNhigh, sample = FALSE)

ss_resnall <-  sample_fish(ss_aggresnall, effNhigh, sample = FALSE)

```

Length sample with user specified max length bin (200 cm):

```{r userset-maxlen, echo=TRUE}

ss_length_census <- calc_age2length(structn = ss_structnall,
                                 resn = ss_resnall,
                                 nums = ss_numsallhigh,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 maxbin = 200,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

```

We should get the upper end of Greenland halibut with a 200cm max length bin. What I am concerned about here is whether we get very small fish at any time step:

```{r sslengthsamp2-test}

lfplot <- ggplot(ss_length_census$natlength, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte() +
  labs(subtitle = paste(scenario.name, ss_length_census$natlength$species))

lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 4, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 5, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 6, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 7, scales="free_y")

```

Are there fish <50 cm long at any output timestep? If not, does this make sense? There appear to be a few, such as timestep 4. Hard to see at this level.

Aggregate length comp over all of these time steps (first 22+ years):

```{r agglengthcomp}

lfplot

```

This suggests that the smallest Greenland turbot in the ecosystem are ~25 cm over all of the timesteps we checked. Does this make sense?

Or do we need to change something about the length comp estimation?