### 2019-05-20 CreateStockSynthesis.Rmd

---
title: "WritingSSfiles"
author: "Christine Stawitz"
date: "May 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
require(here)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/chris/Documents/GitHub/atlantisom/")
```

## Introduction

This page discusses reading in an existing stock assessment file and replacing the data values with those from the `atlantisom` package. To read in the Stock Synthesis files, first ensure you have the latest version of `r4ss`

```{r packages, eval=FALSE}
devtools::install_github("r4ss/r4ss", dependencies = FALSE)
```

## Sardine example

Dummy files are in the package directory; use `r4ss` to load the dummy data file

```{r readdata, echo=FALSE}
require(r4ss)
#Read in sardine data file
sardine.data <- r4ss::SS_readdat_3.30("./inst/extdata/Sardine_SS_files/sardEM_3_3.dat")
```

Below, we can see the structure of the catch and CPUE data
```{r}
sardine.data$catch

sardine.data$CPUE
```

We can next look at age composition data
```{r}
sardine.data$agecomp

sardine.data$lencomp
```

We may also need to access or modify the setup variables before the data matrices.

```{r}
#CPUE units, SD and error structure
sardine.data$CPUEinfo

#Number of length bins
sardine.data$N_lbins

#Length bin method
sardine.data$Lbin_method

#Vector of age bins
sardine.data$agebin_vector

```

## Writing index data
To edit index data, use the `atlantisom` sampling functions to extract survey biomass and numbers. These can then be written using the `SS_write_CPUE` function.

```{r}
results <- get(load("CalCurrentResults.RData"))
devtools::load_all()
species=c("Pacific_sardine")

#Sample all boxes
boxes <- unique(results$nums$polygon)

#Efficiency of the survey
effic <- data.frame(species=species, efficiency=0.5)

# Assume uniform selectivity
sel<- data.frame(species=rep(species,10),
               agecl=c(1:10),
              selex=rep(1,10))

#Sample survey
survey_out <- create_survey(dat=results$nums, time=seq(1,99,1), species=species, boxes=boxes, effic=effic, selex=sel)

#Need to replace with interp function
wtAtAge <- data.frame(species=rep("Pacific_sardine",10),
                agecl=1:10,
                wtAtAge=c(0.0542, 0.0837,	0.1103,	0.1323,	0.1497,	0.163,	0.1729,	0.1801,	0.1854,	0.1941))
cv <- data.frame(species="Pacific_sardine", cv=c(0.2))

#Sample survey biomass
survObsBiom <- sample_survey_biomass(dat=survey_out,cv=cv,wtAtAge)

#Sample survey numbers
survObsNum <- sample_survey_numbers(dat=survey_out,cv=cv)
nyears <- 50
burnin <- 30

#Test writing CPUE in biomass
sardine.data <- SS_write_CPUE(ss_data_list = sardine.data,
              cpue_data = list(survObsBiom$atoutput[burnin:(burnin+nyears-1)]),
              CVs = c(0.2),
              data_years = c(nyears),
              sampling_month = list(rep(1,nyears)),
              units = "biomass")

#Test writing CPUE in numbers
sardine.data <- SS_write_CPUE(ss_data_list = sardine.data,
                             cpue_data = list(survObsNum$atoutput[burnin:(burnin+nyears-1)]),
                             CVs = c(0.2),
                             data_years = c(nyears),
                             sampling_month = list(rep(1,nyears)),
                             units = "numbers")

sardine.data$CPUE
```