### 2019-06-03 SardinesOnlyatlantisom2SStest.Rmd
start adding survey-spec

---
title: "Sardines only end-to-end atlantisom test"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyr)
require(dplyr)
library(ggplot2)
library(data.table)
library(here)
library(ggforce)
library(ggthemes)
library(atlantisom)
```

## Introduction

Now we test the `atlantisom` to SS workflow with a new California Current run that includes climate forcing, recruitment variability, and the fishing scenario outlined for our project in Norway: "output_CC_2063_run11".

We will go from atlantis model output to SS input in this page for a single species, "sardine", that doesn't require any age class splitting.

As of 31 May 2019, the `run_truth` function has been modified to do the catch numbers fix for legacy models.

```{r initialize}

initCCA <- TRUE
initNEUS <- FALSE
initNOBA <- FALSE

species_ss <- c("Pacific_sardine")

#function to make a config file? need one for each atlantis run

if(initCCA) source(here("config/CC2Config.R"))

if(initNEUS) source(here("config/NEUSConfig.R"))

if(initNOBA) source(here("config/NOBAConfig.R"))

```

```{r get_names, message=FALSE, warning=FALSE}
#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

```

This can be run with `species_ss` in select_groups to get results only for that species. Here I'm running with all species; took about 50 minutes to return.

```{r get_truth_ss}

# default run_truth setup will save the file, so check for that first

if(!file.exists(file.path(d.name, 
                          paste0("output", scenario.name, "run_truth.RData")))){
  #Store all loaded results into an R object
  truth <- run_truth(scenario = scenario.name,
                     dir = d.name,
                     file_fgs = functional.groups.file,
                     file_bgm = box.file,
                     select_groups = funct.group.names,
                     file_init = initial.conditions.file,
                     file_biolprm = biol.prm.file,
                     file_runprm = run.prm.file,
                     verbose = TRUE
  )
} else {
  truth <- get(load(file.path(d.name,
                              paste0("output", scenario.name, "run_truth.RData"))))
}
```

<!--In this block, `test` was a vector of 0s, meaning the correction run separately and the new correction implemented in `run_truth` are producing the same results:

```{r catchnums-check, eval=FALSE}

# this was run with results in CalCurrent2013_OA_off folder and run_truth_sardines

catchnumfull <- readRDS(file.path(d.name, paste0(scenario.name, "catchnums_corrected.rds")))

catchnum_ss <- catchnumfull %>%
  filter(species == species_ss) %>%
  arrange(time, polygon, agecl)

new_catage <- truth$catch %>%
  arrange(time, polygon, agecl)

test <- catchnum_ss$atoutput - new_catage$atoutput

# test is all 0 so this worked the same way

```
-->

In the big picture, we will have the true population dynamics from Atlantis which we need for reference to compare with estimated population dynamics from the stock assessment estimation models.

Ideally we will have a function that gets the whole truth (and nothing but the truth) from Atlantis and stores it in a sensible way for later comparison and calculation of performance metrics.

Then we will have a function that implements the "survey" that produces the stock assessment inputs and sends those to the SS file writing functions.

## Truth from OM

The following settings should achieve a survey that samples all Atlantis model output timesteps, all species, and all model polygons, with perfect efficiency and full selectivity for all ages:

```{r census-spec, message=FALSE, warning=FALSE}

# make a function for this
source(here("config/census_spec.R"))

```

Get true total biomass, total numbers. This code gets results for all species and saves the output true B and N.

```{r trueB-N, eval=FALSE}

# this uses result$biomass_ages to sample biomass directly, no need for wt@age est

survey_testBall <- create_survey(dat = truth$biomass_ages,
                                 time = timeall,
                                 species = survspp,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

# make up a constant 0 cv for testing
surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))

# call sample_survey_biomass with a bunch of 1s for weight at age
# in the code it multiplies atoutput by wtatage so this allows us to use
# biomass directly
wtage <- data.frame(species=rep(survspp, each=10),
                    agecl=rep(c(1:10),length(survspp)),
                    wtAtAge=rep(1.0,length(survspp)*10))

surveyB_frombio <- sample_survey_biomass(survey_testBall, surv_cv, wtage)

#save for later use, takes a long time to generate
saveRDS(surveyB_frombio, file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))

# this uses result$nums and a new function to get survey index in numbers (abundance)

survey_testNall <- create_survey(dat = truth$nums,
                                 time = timeall,
                                 species = survspp,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

# as above, make up a constant 0 cv for testing
surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))

surveyN <- sample_survey_numbers(survey_testNall, surv_cv)

#save for later use, takes a long time to generate
saveRDS(surveyN, file.path(d.name, paste0(scenario.name, "surveyNcensus.rds")))

```

Plot true biomass and numbers for "sardines."

```{r plot-true-B-N}

trueB <- readRDS(file.path(d.name, paste0(scenario.name,"surveyBcensus.rds")))

trueB_ss <- trueB %>%
  filter(species==species_ss)

plotB <-ggplot() +
  geom_line(data=trueB_ss, aes(x=time/stepperyr,y=atoutput, color="survey census N"), 
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

plotB +
  facet_wrap(~species, scales="free")

trueN <- readRDS(file.path(d.name, paste0(scenario.name,"surveyNcensus.rds")))

trueN_ss <- trueN %>%
  filter(species==species_ss)

plotN <-ggplot() +
  geom_line(data=trueN_ss, aes(x=time/stepperyr,y=atoutput, color="survey census N"), 
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

plotN +
  facet_wrap(~species, scales="free")

```

Get true population age comp, length comp, and weight at age. Now we go to single species to speed things up.

```{r true-comps, warning=FALSE, message=FALSE, eval=FALSE}

# wrap these individual calls into a function
survey_N_ss <- create_survey(dat = truth$nums,
                                 time = timeall,
                                 species = species_ss,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

# apply default sample fish to get numbers, effNhigh samples all, see census_spec
numssshigh <- sample_fish(survey_N_ss, effNhigh)

# save age comps
saveRDS(numssshigh, file.path(d.name, paste0(scenario.name, "natage_census_sard.rds")))

# aggregate true resn per survey design
aggresnss <- aggregateDensityData(dat = truth$resn,
                                 time = timeall,
                                 species = species_ss,
                                 boxes = boxall)

# aggregate true structn per survey design
aggstructnss <- aggregateDensityData(dat = truth$structn,
                                 time = timeall,
                                 species = species_ss,
                                 boxes = boxall)

#dont sample these, just aggregate them using median
structnss <- sample_fish(aggstructnss, effNhigh, sample = FALSE)

resnss <-  sample_fish(aggresnss, effNhigh, sample = FALSE)

# this function does the length sampling and also weight at age
length_census_ss <- calc_age2length(structn = structnss,
                                 resn = resnss,
                                 nums = numssshigh,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

#save for later use, takes a long time to generate
saveRDS(length_census_ss, file.path(d.name, paste0(scenario.name, "lengthwt_census_sard.rds")))

```

Plot some of the age, length comps, and wtatage

```{r plot-true-agecomps}

Natage <- readRDS(file.path(d.name, paste0(scenario.name, "natage_census_sard.rds")))

Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
  geom_point() +
  theme_tufte() +
  labs(subtitle = paste(scenario.name,
                        Natage$species))

Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
Natageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")

```

```{r lf-plot}

lengthwt_census_ss <- readRDS(file.path(d.name, paste0(scenario.name, "lengthwt_census_sard.rds")))

len <- lengthwt_census_ss$natlength

lfplot <- ggplot(len, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte() +
  labs(subtitle = paste(scenario.name,
                        len$species))

lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 1, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 2, scales="free_y")
lfplot + facet_wrap_paginate(~time, ncol=4, nrow = 4, page = 3, scales="free_y")

```

```{r wtage-plot}

wtage_ann <- lengthwt_census_ss$muweight %>%
  filter(time %in% annualmidyear)

# reverse to show agecl time series of wt
wageplot <- ggplot(wtage_ann, aes(time, atoutput)) +
  geom_line(aes(colour = factor(agecl))) +
  theme_tufte() +
  theme(legend.position = "bottom") +
  xlab("model timestep (5 per year)") +
  ylab("average individual weight (g)") +
  labs(subtitle = paste0(scenario.name, " annual mid year sample"))

wageplot + facet_wrap(c("species"), scales="free_y")

```

Get true population parameters (M in progress, S-R, growth)

```{r true-poppars}

```

Get true catch (may not be different from input for our project)

```{r true-catchtot}

truecatchbio <- load_catch(d.name, file_catch = catch.file, fgs = funct.groups)

truecatchbio_ss <- truecatchbio[truecatchbio$species == species_ss,]

plotcatch <- ggplot() +
  geom_line(data=truecatchbio_ss, aes(x=time,y=atoutput, color="true catch bio"),
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

plotcatch +
  facet_wrap(~species, scales="free")

```

Get true catch at age and other comps (may not be different from input for our project)

```{r true-catage-lf-wtage}

# get survey nums with full (no) selectivity
catchN_ss <- create_fishery_subset(dat = truth$catch,
                                 time = timeall,
                                 species = species_ss,
                                 boxes = boxall)

# apply default sample fish as before to get numbers
catch_numssshigh <- sample_fish(catchN_ss, effNhigh)

# aggregate true resn per survey or fishery subset design
catch_aggresnss <- aggregateDensityData(dat = truth$resn,
                                 time = timeall,
                                 species = species_ss,
                                 boxes = boxall)

# aggregate true structn per survey or fishery subsetdesign
catch_aggstructnss <- aggregateDensityData(dat = truth$structn,
                                 time = timeall,
                                 species = species_ss,
                                 boxes = boxall)

#dont sample these, just aggregate them using median
catch_structnss <- sample_fish(catch_aggstructnss, effNhigh, sample = FALSE)

catch_resnss <-  sample_fish(catch_aggresnss, effNhigh, sample = FALSE)

# get catch lengths and wtage
catch_length_census_ss <- calc_age2length(structn = catch_structnss,
                                 resn = catch_resnss,
                                 nums = catch_numssshigh,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 maxbin = 150,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

#save for later use, takes a long time to generate
saveRDS(catch_length_census_ss, file.path(d.name, paste0(scenario.name, "catchlengthwt_census_sard.rds")))

```

This is true catch at age for sardine:

```{r catage1}

catchage <- catch_numssshigh %>%
  filter(species == "Pacific_sardine")

catageplot <- ggplot(catchage, aes(x=agecl, y=atoutput)) +
  geom_point() +
  theme_tufte() +
  labs(subtitle = paste(scenario.name,
                        catchage$species))

catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 1, scales="free")
catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 2, scales="free")
catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 3, scales="free")
catageplot + facet_wrap_paginate(~time, ncol=3, nrow = 3, page = 4, scales="free")

## "Data" from OM for EMs

Specify our survey sampling. This could come from other information such as the overlap of actual survey stations with OM polygons, experiments evaluating survey selectivity and efficiency, actual sample-based survey cv, etc.

```{r survey-spec}

```

## “Data” from OM for EMs

Specify our survey sampling. This could come from other information such as the overlap of actual survey stations with OM polygons, experiments evaluating survey selectivity and efficiency, actual sample-based survey cv, etc. Here we guess using Christine's specifications:

```{r sardine-survey-spec}
# from atlantisom/config/sardine_config.R, to be generalized here

#Set species name
species <- "Pacific_sardine" #change to species_ss to match

#Directory with SS files
model_dir <- "Sardine_SS_files/"

#Name of SS data file
datfile_name <- "sardEM_3_3.dat"

#Name of atlantis model output - this is an RData object from the runtruth() function
atlantis_output_name <- "outputCCV3run_truth.RData"

#Efficiency
eff_case <- 0.5

#Age classes of your species
age_classes <- 1:10

#Selectivity function - a vector equivalent to length (age classes)
sel_by_age <- rep(1,length(age_classes))

#Atlantis model timestep corresponding to the true output--now from census_spec.R
timestep <- stepperyr #5

#Which atlantis timestep does the survey run in?--now from census_spec.R
survey_sample_time <- midptyr #3

#The last timestep to sample
total_sample <- noutsteps #495

#Vector of indices of survey times to pull
survey_sample_full <- seq(survey_sample_time,
                          total_sample, by=timestep)

#Effective sample size for composition data
surveyEffN <- 1000
fisheryEffN <- 1000

#CVs for length at age, catch, and survey
CVs <- list("lenage"=0.1, "fishery"=0.01, "survey"=0.1)

#Number of years of data to pull
nyears <- 50

#Atlantis initialization period
burnin <- 30

#Maximum size of a fish in cm
maxbin <- 150

#Years to fish, assuming fishing output is annual
fish_years <- burnin:(burnin+nyears-1)

#Years to survey
survey_years <- survey_sample_full[burnin:(burnin+nyears-1)]

#Month of survey/fishing
survey_month <- 7
fishing_month <- 1

```

Get input survey biomass and catch biomass for SS (these have q, selectivity, cv)

```{r toSS-surveyfisheryts}

```

Get composition inputs for SS (survey and fishery catch at age, survey and fishery lengths, survey and fishey weight at age)

```{r toSS-surveyfisherycomps}

```

Get population parameters for SS

```{r toSS-poppars}

```