### 2019-06-04 CheckBiomassAges.Rmd

---
title: "Is Biomass Ages Broken?"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
require(dplyr)
library(ggplot2)
library(data.table)
library(here)
library(ggforce)
library(ggthemes)
library(atlantisom)

```

```{r constants}

initCCA <- TRUE
initNEUS <- FALSE
initNOBA <- FALSE

species_ss <- c("Pacific_sardine")

# make up a constant 0 cv for testing
surv_cv <- data.frame(species=species_ss, cv=rep(0.0,length(species_ss)))

# call sample_survey_biomass with a bunch of 1000s for weight at age
# in the code it multiplies atoutput by wtatage/1000 so this allows us to use
# biomass directly
wtage <- data.frame(species=rep(species_ss, each=10),
                    agecl=rep(c(1:10),length(species_ss)),
                    wtAtAge=rep(1000.0,length(species_ss)*10))

```

## This used to work

Before I cleverly changed sample_survey_biomass to have the right units to use weight at age. Now if we want biomass_ages to give a correct survey, the fake wtage we give it needs to be 1000 instead of 1. So much for a shortcut.

Test biomass_ages use in create survey in old CCA run vs new CCA run

```{r truth-oldCCA}

source(here("config/CC1Config.R"))

#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

source(here("config/census_spec.R"))

sardold <- run_truth(scenario = scenario.name,
                     dir = d.name,
                     file_fgs = functional.groups.file,
                     file_bgm = box.file,
                     select_groups = species_ss,
                     file_init = initial.conditions.file,
                     file_biolprm = biol.prm.file,
                     file_runprm = run.prm.file,
                     verbose = TRUE
)

survey_testBold <- create_survey(dat = sardold$biomass_ages,
                                 time = timeall,
                                 species = species_ss,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

surveyB_frombio_old <- sample_survey_biomass(survey_testBold, surv_cv, wtage)

# read Atlantis output files
atBtxt2old <- read.table(file.path(d.name, paste0("output", scenario.name, "BiomIndx.txt")), header=T)

# lookup the matching names, put in time, species, biomass column format
# WARNING hardcoded for output with last species group as DIN
groupslookup <- funct.groups %>%
  filter(IsTurnedOn > 0)

atBtxt2tidyold <- atBtxt2old %>%
  select(Time:DIN) %>%
  #select(Time, FPL:DIN) %>%
  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
  gather(species, biomass, -Time) %>%
  filter(species %in% species_ss)

```

Compare with BiomIndex.txt

```{r compareold}

surveyB_lastweek_old <- readRDS(file.path(d.name, paste0(scenario.name, "surveyBcensus.rds")))

surveyB_lastweek <- surveyB_lastweek_old %>%
  filter(species==species_ss)

compareB <-ggplot() +
  geom_line(data=surveyB_frombio_old, aes(x=time/5,y=atoutput, color="survey census B"), 
            alpha = 10/10) +
  geom_line(data=surveyB_lastweek, aes(x=time/5,y=atoutput, color="survey census B last week"), 
            alpha = 1/10) +
  geom_point(data=atBtxt2tidyold, aes(x=Time/365,y=biomass, color="txt output true B"),
           alpha = 10/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

compareB + 
  facet_wrap(~species, scales="free")

```

## Why doesnt it now? It does now.

```{r truth-newCCA}

source(here("config/CC2Config.R"))

#Load functional groups
funct.groups <- load_fgs(dir=d.name,
                         file_fgs = functional.groups.file)
#Get just the names of active functional groups
funct.group.names <- funct.groups %>% 
  filter(IsTurnedOn == 1) %>%
  select(Name) %>%
  .$Name

source(here("config/census_spec.R"))

sardnew <- run_truth(scenario = scenario.name,
                     dir = d.name,
                     file_fgs = functional.groups.file,
                     file_bgm = box.file,
                     select_groups = species_ss,
                     file_init = initial.conditions.file,
                     file_biolprm = biol.prm.file,
                     file_runprm = run.prm.file,
                     verbose = TRUE
)

survey_testBnew <- create_survey(dat = sardnew$biomass_ages,
                                 time = timeall,
                                 species = species_ss,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

surveyB_frombio_new <- sample_survey_biomass(survey_testBnew, surv_cv, wtage)

# read Atlantis output files
atBtxt2new <- read.table(file.path(d.name, paste0("output", scenario.name, "BiomIndx.txt")), header=T)

# lookup the matching names, put in time, species, biomass column format
# WARNING hardcoded for output with last species group as DIN
groupslookup <- funct.groups %>%
  filter(IsTurnedOn > 0)

atBtxt2tidynew <- atBtxt2new %>%
  select(Time:DIN) %>%
  #select(Time, FPL:DIN) %>%
  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
  gather(species, biomass, -Time) %>%
  filter(species %in% species_ss)

```

Compare with BiomIndex.txt

```{r comparenew}

compareB <-ggplot() +
  geom_line(data=surveyB_frombio_new, aes(x=time/5,y=atoutput, color="survey census B"), 
            alpha = 10/10) +
  geom_point(data=atBtxt2tidynew, aes(x=Time/365,y=biomass, color="txt output true B"),
             alpha = 10/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

compareB + 
  facet_wrap(~species, scales="free")

```