### 2019-06-07 docs/index.Rmd

---
title: "Poseidon Adventure"
output: html_document
---

 <!-- ADD LINKS BELOW, note, using the RMD link syntax will open the presentation in the same tab, which is annoying. Use the html with the "_blank" to fix that behavior -->

These pages document initial testing of the atlantisom package in development at https://github.com/r4atlantis/atlantisom using three different [Atlantis](https://research.csiro.au/atlantis/) output datasets. Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA.

### An example workflow for a single species (in progress)

<a href="https://sgaichas.github.io/poseidon-dev/FullSardineTruthEx.html" target="_blank">Sardines in the CalCurrent, all true data, May 22, 2019, updated May 23, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CreateStockSynthesis.html" target="_blank">Sardines in the CalCurrent, input to SS functions: May 24, 2019</a> - now generalized to run for any species in the cal current with a unique species config file

May 24: Sardines have a full dataset, but still working on converting to true ages for species with multiple ages per age class. See below under [True age compositions].

<a href="https://sgaichas.github.io/poseidon-dev/SardinesOnlyatlantisom2SStest.html" target="_blank">Sardines in the CalCurrent run with F and rec dev scenario, all true and SS data, May 31, 2019, updated June 7</a>

### Read this first on overall workflow

<a href="https://sgaichas.github.io/poseidon-dev/RethinkSamplingFunctions.html" target="_blank">Big picture and new functions needed - May 10, 2019, updated May 16, 2019</a>

May 11: Changes to `atlantisom::sample_fish` and new `atlantisom::aggregateDensityData` functions seem to work. See below under [Length compositions].

May 15: Changes to `atlantisom::calc_age2length` to allow user specified max length bin seem to work. Also listed under [Length compositions].

May 20-22: Fisheries, corrections to nums at age for Cal Current and decision to read in the catch.txt file for total catch in weight

### Testing the atlantisom code

<a href="https://sgaichas.github.io/poseidon-dev/TrueBioTest.html" target="_blank">Installation and initial tests - April 26, 2019, updated April 30, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NEUSExamples.html" target="_blank">Survey census visualization for NEUS -  April 30, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCAExamples.html" target="_blank">Survey census visualization for CCA - April 30, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NOBAExamples.html" target="_blank">Survey census visualization for NOBA - April 30, 2019</a>

### Create a standard survey

<a href="https://sgaichas.github.io/poseidon-dev/StandardSurveyTest.html" target="_blank">Initial survey tests - April 30, 2019, updated May 1, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NEUSStandardSurveyTest.html" target="_blank">Standard survey visualization for NEUS -  May 1, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCAStandardSurveyTest.html" target="_blank">Standard survey visualization for CCA - May 1, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NOBAStandardSurveyTest.html" target="_blank">Standard survey visualization for NOBA - May 1, 2019</a>

### Start age composition sampling: by output age class, not true age

<a href="https://sgaichas.github.io/poseidon-dev/TrueLengthAgeCompTest.html" target="_blank">Initial age comp tests - May 2, 2019, updated May 6, 2019</a>

CCA outputs had some 0 length at age, requiring changes to `atlantisom::sample_fish`

<a href="https://sgaichas.github.io/poseidon-dev/NEUSTrueLengthAgeCompTest.html" target="_blank">True and sample fish census comparison NEUS - May 6, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCATrueLengthAgeCompTest.html" target="_blank">True and sample fish census comparison CCA - May 6, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NOBATrueLengthAgeCompTest.html" target="_blank">True and sample fish census comparison NOBA - May 6, 2019</a>

Standard survey with composition sampling: impact of selectivity, `sample_fish`

<a href="https://sgaichas.github.io/poseidon-dev/StdSurvLengthAgeCompTest.html" target="_blank">Compare true cohort comp with survey + sampling - May 6, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NEUSStdSurvCohortCompTest.html" target="_blank">True cohort comp vs survey comparison NEUS - May 6, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCAStdSurvCohortCompTest.html" target="_blank">True cohort comp vs survey comparison CCA - May 6, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NOBAStdSurvCohortCompTest.html" target="_blank">True cohort comp vs survey comparison NOBA - May 6, 2019</a>

### Length compositions

Difficult when you don't have a good idea of the outputs you are using!

<a href="https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest_fail.html" target="_blank">Defeated by length compositions - May 8-9, 2019</a>

Better when you do:

<a href="https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html" target="_blank">Victory: length compositions - May 11, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCATrueLengthCompTest.html" target="_blank">CCA length composition examples - May 11, 2019</a>

But don't try running length comps at the subannual level for all species.

<a href="https://sgaichas.github.io/poseidon-dev/NOBATrueLengthCompTest.html" target="_blank">NOBA length composition examples - May 13, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NEUSTrueLengthCompTest.html" target="_blank">NEUS length composition examples - May 13, 2019</a>

Standard survey with length sampling: impact of selectivity, `sample_fish`

<a href="https://sgaichas.github.io/poseidon-dev/StdSurvLengthCompTest.html" target="_blank">Compare survey + sampled lengths with true - May 14, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/CCAStdSurvLengthCompTest.html" target="_blank">CCA compare survey + sampled lengths with true - May 14, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NOBAStdSurvLengthCompTest.html" target="_blank">NOBA compare survey + sampled lengths with true - May 15, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/NEUSStdSurvLengthCompTest.html" target="_blank">NEUS compare survey + sampled lengths with true - May 15, 2019</a>

Test new user specified max length bin for a large species:

<a href="https://sgaichas.github.io/poseidon-dev/TestBiggerMaxLen.html" target="_blank">Default 150 cm, try different max - May 15, 2019</a>

Do we see ever small fish for large species? Is this normal, or a problem with length estimation?

<a href="https://sgaichas.github.io/poseidon-dev/TestSmallestLength.html" target="_blank">Are length comps OK on the small end? - May 15, 2019</a>

May 16: Isaac says CCA fish look reasonable. NOBA Greenland halibut growth parameterization was reportedly "a nightmare" with conflicting science and datasets, so maybe we don't worry too much about them right now.

### Weight at age: by output age class, how to get to true age basis?

<a href="https://sgaichas.github.io/poseidon-dev/TestWeightAtStage.html" target="_blank">How much variability in weight at stage? - May 16, 2019</a>

Work in progress on interpolating weight at age in InterpolateWeightAtAge.Rmd

### Testing a full workflow through fisheries (sardine, agecl = true age)

<a href="https://sgaichas.github.io/poseidon-dev/Atlantisom2SSworkflowtest.html" target="_blank">Try for true datasets new CC model with climate - May 21, 2019, updated May 22, 2019 with catch correction</a>

Fishery catch problem, older codebase has a bug in output. Test with NOBA...

<a href="https://sgaichas.github.io/poseidon-dev/TestCatchNOBA.html" target="_blank">Suggests a workaround needed for catch in weight - May 21, 2019</a>

### True age compositions

<a href="https://sgaichas.github.io/poseidon-dev/FullBoccaccioTruthEx.html" target="_blank">Initial testing of true age functions, now running but output still wrong - May 24, 2019, updated May 27, 2019</a>

<a href="https://sgaichas.github.io/poseidon-dev/TestCalcZ.html" target="_blank">Test Z calc and compare with true annage NOBA - May 24, 2019</a>

May 24: Either `calc_Z` or `calc_stage2age` or both are incorrect if we expect them to exactly match atlantis annual age ouput where it is available. We match in timestep 0, then diverge. May 27 update: also true for Bocaccio in CCA, the interim numbers at age drop to 0.

June 7: `calc_Z` seems incorrect when compared to outputMort.txt; stepping through code needed

<a href="https://sgaichas.github.io/poseidon-dev/TestCalcZ2.html" target="_blank">Test Z calc and compare with output mort CCA - June 7, 2019</a>

### Rabbit holes that turned out ok

<a href="https://sgaichas.github.io/poseidon-dev/CheckBiomassAges.html" target="_blank">In which I broke the very first thing that worked, then fixed it - June 4, 2019</a>

### More to come later!