### 2020-01-13 WrapperDevelopment.Rmd
add details

---
title: "Develop atlantisom wrappers"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Too many steps!

We can run atlantisom with many functions to get stock assessment model inputs.

To make life easier, here are some wrapper functions to do the basic things we want.

Basic desired functionality:

1. initialize: locate model output (function to make a config file?), read in basic run parameters and species names, run_truth for all once and save output

  + User to make up front decision--want all species or just a subset?
  + May want multiple species that need to be split in a later step for models

1. get assessment input data: configure survey, run all survey functions, configure fishery, run all fishery functions (how to save output?)

  + Based on decision above, save interim steps for use later
  + Need a switch for single vs multispecies estimation models?

1. write input data for specific model: wrap functions in development for SS3, add functions for other models (ASAP, multispecies models)

## Cumbersome giant files eat all the memory!

Implement Christine's memory saving things from the NEMOW demo