### 2020-01-16 WrapperDevelopment.Rmd
begin wrapper functions

---
title: "Develop atlantisom wrappers"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Too many steps!

We can run atlantisom with many functions to get stock assessment model inputs.

To make life easier, here are some wrapper functions to do the basic things we want.

Basic desired functionality:

1. initialize: locate model output (function to make a config file?), read in basic run parameters and species names, run_truth for all once and save output 
    + User to make up front decision--want all species or just a subset?
    + May want multiple species that need to be split in a later step for models
    + wraps functions 
        + sourcing model config file--add function to make config file?
      + specifies top directory for source files, names key files for use later
      + load_fgs
      + load all other needed files--YOY, catch
      + run_truth
1. get assessment input data: configure survey, run all survey functions, configure fishery, run all fishery functions (how to save output?)
    + Based on decision above, save interim steps for use later
    + Need a switch for single vs multispecies estimation models?
    + Separate index and comp functions in output
    + indices:
      + create survey config file(s), default is a census of all see census_spec.R
        + survey index in biomass or numbers? (uses truth$biomass_ages vs truth$nums)
        + wraps functions create_survey, sample_survey_biomass
      + create fishery config file(s)
    + comps:
      + select age only or age and length
        + age only wraps create_survey using nums with sample_fish to get age comps
        + age and length (with weight at age)
          + create_survey followed by sample_fish, 
          + aggregateDensityData for resn and structn followed by sample_fish
          + inputs all to calc_age2length
      + catch comps wrap create_fishery_subset with sample_fish to get age comps, same as above for lengths
1. write input data for specific model: wrap functions in development for SS3, add functions for other models (ASAP, multispecies models)
    + wrap r4ss functions to read and write dat and ctl files
    + possibly start with set of dummy SS files with package (take from ss3sim?)
      + read in dat file
      + indices wrap SS_write_ts
      + comps wrap reformat_compositions, SS_write_comps, some of the checks for bin matches\
      + write full dat file
    + start with dummy ctl file (any from ss3sim?)
      + read in ctl file
      + wrap SS_write_bio using outputs of calc_Z, load_YOY, load_biolprm, load_runprm, etc
1. run specified model--automate for scenarios, save outputs 
1. skill assessment metrics: compare stored Atlantis truth with model output

### atlantisom initialize function

```{r ominit}

library(tidyverse)

om_init <- function(config = configfile){

  # Where are the atlantis output files? Consider filling with shiny app in future
  source(config)
  # needs these files, for example config file CC3config.R is:

  # d.name <- here::here("atlantisoutput","CC_2063_OA_OFF_22")
  # functional.groups.file <- "CalCurrentV3Groups.csv"
  # biomass.pools.file <- "DIVCalCurrentV3_Biol.nc"
  # biol.prm.file <- "CalCurrentV3_Biol.prm"
  # box.file <- "CalCurrentV3_utm.bgm"
  # initial.conditions.file <- "DIVCalCurrentV3_Biol.nc"
  # run.prm.file <- "CalCurrentV3_run.xml"
  # scenario.name <- "CCV3"
  # bioind.file <- "outputCCV3BiomIndx.txt"
  # catch.file <- "outputCCV3Catch.txt"

  #Load functional groups
  funct.groups <- atlantisom::load_fgs(dir=d.name,
                           file_fgs = functional.groups.file)
  #Get just the names of active functional groups
  funct.group.names <- funct.groups %>% 
    filter(IsTurnedOn == 1) %>%
    select(Name) %>%
    .$Name

  # load true catch in tons
  truecatchbio <- atlantisom::load_catch(d.name, file_catch = catch.file, fgs = funct.groups)

  # load YOY
  YOY <- atlantisom::load_yoy(d.name, paste0("output", scenario.name, "YOY.txt"))

  # load biol_prm
  biol <- atlantisom::load_biolprm(d.name, biol.prm.file)

  # load run_prm
  runpar <- atlantisom::load_runprm(d.name, run.prm.file)

  # default run_truth setup will save the file, so check for that first
  if(!file.exists(file.path(d.name, 
                            paste0("output", scenario.name, "run_truth.RData")))){
    #Store all loaded results into an R object
    truth <- atlantisom::run_truth(scenario = scenario.name,
                       dir = d.name,
                       file_fgs = functional.groups.file,
                       file_bgm = box.file,
                       select_groups = funct.group.names,
                       file_init = initial.conditions.file,
                       file_biolprm = biol.prm.file,
                       file_runprm = run.prm.file,
                       verbose = TRUE
    )
  } else {
    truth <- get(load(file.path(d.name,
                                paste0("output", scenario.name, "run_truth.RData"))))
  }

  omlist <-list("funct.groups" = funct.groups, 
                "funct.group.names" = funct.group.names,
                "truecatchbio" = truecatchbio,
                "YOY" = YOY,
                "biol" = biol,
                "runpar" = runpar,
                "truth" = truth)

  return(omlist)
}

```

Usage:

```{r exinit}

library(here)
CC3om <- om_init(here("config/CC3config.r"))

```

### Get assessment input data

Split to single species or subset of assessed species, get index data, get compositional data.

```{r inputs}

om_species <- function(species = spp, omlist){
  # spp format c("speciesname1", "speciesname2")
  if(!all(species %in% omlist$funct.group.names)) stop("species name not found") 
  species_ss <- species

  return(species_ss)
}

```

Usage:

```{r exsp}

om_species(c("Pacific_sardine"), CC3om)

```

## Cumbersome giant files eat all the memory!

Implement Christine's memory saving things from the NEMOW demo