### 2020-03-18 NOBAcod.Rmd
initial commit NOBA full age testing

---
title: "Generate data for NOBA cod ensemble modeling"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Standard atlantisom output test (age class not age)

Let's use the wrapper functions to get initial stock assessment model inputs.

Then we need to adjust the base atlantisom functions for true age classes, and develop options to call those within the wrappers.

We need config files for the NOBA model, our surveys and fisheries.

`NOBAconfig.R` looks like this (adjusted from Alfonso's original):

---
d.name <- here("atlantisoutput","NOBA_march_2020")
functional.groups.file <- "nordic_groups_v04.csv"
biomass.pools.file <- "nordic_biol_v23.nc"
biol.prm.file <- "nordic_biol_incl_harv_v_007_3.prm"
box.file <- "Nordic02.bgm"
initial.conditions.file <- "nordic_biol_v23.nc"
run.prm.file <- "nordic_run_v01.xml"
scenario.name <- "nordic_runresults_01"
bioind.file <- "nordic_runresults_01BiomIndx.txt"
catch.file <- "nordic_runresults_01Catch.txt"
---

`codsurvey.R` configures the fishery independent survey and looks like this:

---
timeall <- c(0:560)

funct.group.indices <- c(1:36)
---

`codfishery.R` configures the fishery dependent data and looks like this:

---

---

All the config files go in the config folder and we'll try running this:

```{r full-test}

library(here)
library(tidyverse)
library(atlantisom)

NOBAom <- om_init(here("config/NOBAconfig.R"))

NOBAom_cod <- om_species(c("Pacific_sardine"), NOBAom)

NOBAom_cod_ind <- om_index(usersurvey = here("config/sardinesurvey.R"), 
                           userfishery = here("config/sardinefishery.R"),
                           omlist_ss = CC3om_sardine, 
                           n_reps = 1, 
                           save = TRUE)

NOBAom_cod_comp <- om_comps(usersurvey = here("config/sardinesurvey.R"), 
                           userfishery = here("config/sardinefishery.R"),
                           omlist_ss = CC3om_sardine, 
                           n_reps = 1, 
                           save = TRUE)

```

### Write input data for specific assessment models

Use the saved outputs of above to write a Stock Synthesis dat file
This is not a function yet, copy in Christine's

**See corrections below, in fishery time fixed and nsamp added to length comps**

```{r writedat}

require(r4ss)
omlist_ss<-CC3om_sardine
source(here("/config/CC3Config_sardLWcorr.R"))
source(here("/config/omdimensions.R"))
source(here("/config/sardinesurvey.R"))
source(here("/config/sardinefishery.R"))

#Directory with SS files
model_dir <- "Sardine_SS_files/"

#Name of SS data file
datfile_name <- "sardEM_3_3.dat"

#CVs for length at age, catch, and survey
CVs <- list("lenage"=0.1, "fishery"=0.01, "survey"=0.1)

#Month of survey/fishing
survey_month <- 7
fishing_month <- 1

#Years to survey, assuming survey is once per year
#survey_years <- survey_sample_full[(burnin+1):(burnin+nyears)] # my original
survey_years <- survey_sample_full[burnin:(burnin+nyears-1)] #from Christine's new sardine_config.R

#read time series data
survObsBiom <- readRDS(file.path(d.name, paste0(scenario.name, "surveyB.rds")))
truecatchbio_ss <- readRDS(file.path(d.name, paste0(scenario.name, "fishCatch.rds")))

survObsBiom <- survObsBiom[[1]]
truecatchbio_ss <- truecatchbio_ss[[1]]

#add this to om_indices function so that this has years when read in
truecatchbio_ss$time <- as.integer(truecatchbio_ss$time/365)

#load dummy dat file
stocksynthesis.data <- r4ss::SS_readdat_3.30(paste0("./inst/extdata/",
model_dir,
datfile_name))

#Test writing CPUE in biomass
  stocksynthesis.data <- atlantisom::SS_write_ts(ss_data_list = stocksynthesis.data,
                ts_data = list(survObsBiom$atoutput[fish_years],
                  truecatchbio_ss$atoutput[truecatchbio_ss$time %in% fish_years]),
                CVs = c(CVs$survey,
                        CVs$fishery),
                data_years = list((survObsBiom$time[fish_years]-survey_sample_time)/timestep+1,             fish_years),
            sampling_month = list(rep(survey_month,nyears),
                                rep(fishing_month,nyears)),
                units = c("biomass","biomass"),
                data_type=c("CPUE","catch"),
                fleets = c(2,1))

#test  
stocksynthesis.data$CPUE

stocksynthesis.data$catch

#read in comp data
age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsAgeComp.rds")))
age_comp_data <- age_comp_data[[1]]

fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
fish_age_comp <- fish_age_comp[[1]]

#Get the age bins
age_bin_names <- names(stocksynthesis.data$agecomp)[10:length(names(stocksynthesis.data$agecomp))]
age_bins <- sub("a","",age_bin_names)

require(maditr)
# add dependency on maditr::dcast to atlantisom
age_comp_flat <- atlantisom::reformat_compositions(age_comp_data,
                     round.places = 4,
                     comp_type = "agecomp")

## Write age composition data for survey
stocksynthesis.data <- atlantisom::SS_write_comps(ss_data_list = stocksynthesis.data,
               comp_matrix = list(age_comp_flat[burnin:(burnin+nyears-1),]),
               data_rows = list(stocksynthesis.data$styr:(stocksynthesis.data$styr+nyears-1)),
               sampling_month = list(rep(survey_month,nyears)),
               data_type = c("agecomp"),
               fleet_number = c(1),
               bins = list(age_bins),
               caal_bool = c(FALSE))
stocksynthesis.data$agecomp

#length comps
len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsLenComp.rds")))
fish_len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsLenComp.rds")))

len_comp_data <- len_comp_data[[1]]
fish_len_comp_data <- fish_len_comp_data[[1]]

#add this to om_indices function so that this has years when read in
fish_len_comp_data$time <- as.integer(floor(fish_len_comp_data$time/fstepperyr))
fish_age_comp$time <- as.integer(floor(fish_age_comp$time/fstepperyr))

if(fstepperyr>1){
  fish_len_comp_anndata <- fish_len_comp_data %>%
    #mutate(yr = floor(time/fstepperyr)) %>%
    #group_by(species, agecl, lower.bins, upper.bins, time=as.integer(yr)) %>%
    group_by(species, agecl, lower.bins, upper.bins, time) %>% 
    summarise(annnatlength=sum(atoutput)) %>%
    rename(atoutput = annnatlength)
} else {
  fish_len_comp_anndata <- fish_len_comp_data
}

caal_comp_flat <- atlantisom::reformat_compositions(len_comp_data,                                round.places=4,
                comp_type="caalcomp")

#remove burnin
caal_comp_final <- filter(caal_comp_flat,
                         time %in% survey_years)

#Add over age classes to get sample size
len_comp_flat <- atlantisom::reformat_compositions(len_comp_data,
                                  round.places = 0,
                           comp_type="lencomp")
#remove burnin
len_comp_final <- filter(len_comp_flat,
                         time %in% survey_years)

length_bins <- as.integer(names(len_comp_final))
length_bins <- length_bins[!is.na(length_bins)]

# fishery length comps are still 5 timesteps per year
# need to aggregate to annual (done above)
# also,  make effN annual goal/fstepsperyr (done above)
fish_len_comp_flat <- atlantisom::reformat_compositions(fish_len_comp_anndata,
                                  round.places = 0,
                           comp_type="lencomp")

#remove burnin works after adjustment above
fish_len_comp_final <- filter(fish_len_comp_flat,
                         time %in% fish_years)

notbins <- c("time", "nsamp")

# fish_length_bins <- as.integer(names(fish_len_comp_final))
# fish_length_bins <- fish_length_bins[!is.na(fish_length_bins)]

# need to fill empty length bins with 0s to have same bins as survey for SS_write_comps
missing.lengths <- setdiff(length_bins, names(fish_len_comp_final)[!names(fish_len_comp_final) %in% notbins])
fish_len_comp_final[as.character(missing.lengths)] <- 0                    # Add them, filled with '0's
fish_len_comp_final <- fish_len_comp_final[c("time", length_bins, "nsamp")] #

# fishery age comps also 5 timesteps per year
if(fstepperyr>1){
  fish_age_comp_anndata <- fish_age_comp %>%
    #mutate(yr = floor(time/fstepperyr)) %>%
    #group_by(species, agecl, time=as.integer(yr)) %>%
    group_by(species, agecl, time) %>%
    summarise(annnatage=sum(atoutput)) %>%
    rename(atoutput = annnatage)
} else {
  fish_age_comp_anndata <- fish_age_comp
}

fish_age_comp_flat <- atlantisom::reformat_compositions(fish_age_comp_anndata,
                           comp_type="agecomp")

#remove burnin (not necessary?fish comps made with fish_years only) 
fish_age_comp_final <- filter(fish_age_comp_flat,
                         time %in% fish_years)

# #SS_write_comps breaking because fishery age bins start with 2 not 1; extracting bins from fish file may help?
# fish_age_bins <- names(fish_age_comp_flat)[!names(fish_age_comp_flat) %in% notbins]

# that leaves an empty column in data file, so instead fill with 0s
missing.ages <- setdiff(age_bins, names(fish_age_comp_final)[!names(fish_age_comp_final) %in% notbins])
fish_age_comp_final[missing.ages] <- 0                    # Add them, filled with '0's
fish_age_comp_final <- fish_age_comp_final[c("time", age_bins, "nsamp")]

comp_list <- list(caal_comp_final,len_comp_final, fish_age_comp_final, fish_len_comp_final)

apply_month <- list(rep(survey_month, nrow(comp_list[[1]])), 
                    rep(survey_month, nrow(comp_list[[2]])),
                    rep(fishing_month,nrow(comp_list[[3]])),
                    rep(fishing_month,nrow(comp_list[[4]])))

# This now runs by ensuring that survey and fishery compositions have the same bins 
# (filled with 0s for missing bins in fishery relative to survey)

# Write CAAL and length composition data
stocksynthesis.data <- atlantisom::SS_write_comps(ss_data_list = stocksynthesis.data, 
                                      comp_matrix = comp_list, 
                                      data_rows = list((comp_list[[1]]$time-survey_sample_time)/timestep + 1 , (survey_years-survey_sample_time)/timestep + 1,fish_years,fish_years),
                                      sampling_month = apply_month, 
                                      data_type = rep(c("agecomp", "lencomp"),2), 
                                      fleet_number = c(2,2,1,1),  
                                      bins = list(age_bins, 
                                                  length_bins, 
                                                  age_bins, 
                                                  length_bins), 
                                      caal_bool = c(TRUE, rep(FALSE,3)))

head(stocksynthesis.data$lencomp)
head(stocksynthesis.data$agecomp)

#Change length bin structure to match atlantis data
stocksynthesis.data$lbin_vector <- length_bins

#Get correct number of length bins
stocksynthesis.data$N_lbins <- length(length_bins)

#Set lbin_method to 1 - this makes the population length bins match the data bins 
#When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
stocksynthesis.data$lbin_method <- 1
stocksynthesis.data$binwidth <- "#"
stocksynthesis.data$minimum_size <- "#"
stocksynthesis.data$maximum_size <- "#"

#Change minimum sample size to 0.001 for CAAL data (SS won't let it go lower than this)
stocksynthesis.data$age_info$minsamplesize <- rep(0.001,2)

SS_writedat_3.30(stocksynthesis.data, outfile = paste0("./inst/extdata/",model_dir,
datfile_name),
                 overwrite=TRUE)

```