### 2020-03-18 NOBAcod.Rmd
start on wrapping NOBA cod

---
title: "Generate data for NOBA cod ensemble modeling"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Standard atlantisom output test (age class not age)

Let's use the wrapper functions to get initial stock assessment model inputs.

Then we need to adjust the base atlantisom functions for true age classes, and develop options to call those within the wrappers.

We need config files for the NOBA model, our surveys and fisheries. Surveys and fisheries are based on the ones we used for sardine in the California Current as a test.

*Special note: Atlantis output files usually have the prefix "output" and `atlantisom` expects this. NOBA output files use a different naming convention. I've appended "output" to the beginning of each NOBA output filename for `atlantisom` to work.*

`NOBA2config.R` looks like this (adjusted from Alfonso's original):

```
d.name <- here("atlantisoutput","NOBA_march_2020")
functional.groups.file <- "nordic_groups_v04.csv"
biomass.pools.file <- "nordic_biol_v23.nc"
biol.prm.file <- "nordic_biol_incl_harv_v_007_3.prm"
box.file <- "Nordic02.bgm"
initial.conditions.file <- "nordic_biol_v23.nc"
run.prm.file <- "nordic_run_v01.xml"
scenario.name <- "nordic_runresults_01"
bioind.file <- "outputnordic_runresults_01BiomIndx.txt"
catch.file <- "outputnordic_runresults_01Catch.txt"
```

`omdimensions.R` standardizes timesteps, etc.:

```
#survey species inherited from omlist_ss
survspp <- omlist_ss$species_ss
# survey season and other time dimensioning parameters
# generalized timesteps all models
noutsteps <- omlist_ss$runpar$tstop/omlist_ss$runpar$outputstep
timeall <- c(0:noutsteps)
stepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutinc
midptyr <- round(median(seq(0,stepperyr)))

# model areas, subset in surveyconfig
allboxes <- c(0:(omlist_ss$boxpars$nbox - 1))

# fishery output: learned the hard way this can be different from ecosystem outputs
fstepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutfinc

# survey selectivity (agecl based)
# should return all age classes fully sampled (Atlantis output is 10 age groups per spp)
n_age_classes <- omlist_ss$funct.group_ss$NumCohorts
age_classes <- 1:n_age_classes

n_annages <- n_age_classes * omlist_ss$funct.group_ss$NumAgeClassSize
annages <- 1:n_annages

```

`codsurvey.R` configures the fishery independent survey and looks like this:

```
# Default survey configuration here is a census
# Need to define survey season, area, efficiency, selectivity

#Atlantis model timestep corresponding to the true output--now from census_spec.R
timestep <- stepperyr #5

#Which atlantis timestep does the survey run in?--now from census_spec.R
# with 5 output steps per year, 0 is Jan-Feb-midMar, 1 is midMar-Apr-May, 
# 2 is June-July-midAug, 3 is midAug-Sept-Oct, 4 is Nov-Dec (ish)

survey_sample_time <- midptyr #2; defined in omdimensions.R

#The last timestep to sample
total_sample <- noutsteps-1 #495

#Vector of indices of survey times to pull
survey_sample_full <- seq(survey_sample_time,
                          total_sample, by=timestep)

survtime <- survey_sample_full

# survey area
# should return all model areas
survboxes <- allboxes

# survey efficiency (q)
# should return a perfectly efficient survey 
surveffic <- data.frame(species=survspp,
                     efficiency=rep(0.5,length(survspp)))

# survey selectivity (agecl based)
# this is by age class, need to change to use with ANNAGEBIO output
survselex <- data.frame(species=rep(survspp, each=n_age_classes),
                     agecl=rep(c(1:n_age_classes),length(survspp)),
                     selex=rep(1.0,length(survspp)*n_age_classes))

# effective sample size needed for sample_fish
# this effective N is high but not equal to total for numerous groups
surveffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))

# survey index cv needed for sample_survey_xxx
# cv = 0.1
surv_cv <- data.frame(species=survspp, cv=rep(0.1,length(survspp)))

# length at age cv for input into calc_age2length function
# function designed to take one cv for all species, need to change to pass it a vector
lenage_cv <- 0.1

# max size bin for length estimation, function defaults to 150 cm if not supplied
maxbin <- 150

```

`codfishery.R` configures the fishery dependent data and looks like this:

```
# Default fishery configuration here is a census
# Inherits species from input omlist_ss
fishspp <- omlist_ss$species_ss

#Number of years of data to pull
nyears <- 50

#Atlantis initialization period in years
burnin <- 30

# same time dimensioning parameters as in surveycensus.R
#Vector of indices of catch in numbers to pull (by timestep to sum)
fish_sample_full <- c(0:total_sample)  #total_sample defined in sardinesurvey.R
fish_burnin <- burnin*fstepperyr+1
fish_nyears <- nyears*fstepperyr
fish_times <- fish_sample_full[fish_burnin:(fish_burnin+fish_nyears-1)]
fish_timesteps <- seq(fish_times[5], max(fish_times), by=fstepperyr) #last timestep
#fish_years <- unique(floor(fish_times/fstepperyr)+1) # my original
fish_years <- unique(floor(fish_times/fstepperyr)) #from Christine's new sardine_config.R

fishtime <- fish_times

# fishery output: learned the hard way this can be different from ecosystem outputs
fstepperyr <- if(omlist_ss$runpar$outputstepunit=="days") 365/omlist_ss$runpar$toutfinc

# fishery sampling area
# should return all model areas, this assumes you see everything that it caught
fishboxes <- c(0:(omlist_ss$boxpars$nbox - 1))

# effective sample size needed for sample_fish
# this effective N is high but not equal to total for numerous groups
fisheffN <- data.frame(species=survspp, effN=rep(1000, length(survspp)))

#this adjusts for subannual fishery output so original effN is for whole year
fisheffN$effN <- fisheffN$effN/fstepperyr

# fishery catch cv can be used in sample_survey_biomass
# perfect observation
fish_cv <- data.frame(species=survspp, cv=rep(0.01,length(survspp)))
```

All the config files go in the config folder and we'll try running this:

```{r full-test, eval=FALSE}

library(here)
library(tidyverse)
library(atlantisom)

NOBAom <- om_init(here("config/NOBA2config.R"))

NOBAom_cod <- om_species(c("North_atl_cod"), NOBAom)

NOBAom_cod_ind <- om_index(usersurvey = here("config/codsurvey.R"), 
                           userfishery = here("config/codfishery.R"),
                           omlist_ss = NOBAom_cod, 
                           n_reps = 1, 
                           save = TRUE)

NOBAom_cod_comp <- om_comps(usersurvey = here("config/sardinesurvey.R"), 
                           userfishery = here("config/sardinefishery.R"),
                           omlist_ss = NOBAom_cod, 
                           n_reps = 1, 
                           save = TRUE)

```

### Write input data for specific assessment models

We will need to alter this for SAM, SS3, etc. Right now just makes SS3.

**THIS IS STILL THE CALIFORNIA CURRENT INFO, TO BE UPDATED**

```{r writedat, eval=FALSE}

require(r4ss)
omlist_ss<-CC3om_sardine
source(here("/config/CC3Config_sardLWcorr.R"))
source(here("/config/omdimensions.R"))
source(here("/config/sardinesurvey.R"))
source(here("/config/sardinefishery.R"))

#Directory with SS files
model_dir <- "Sardine_SS_files/"

#Name of SS data file
datfile_name <- "sardEM_3_3.dat"

#CVs for length at age, catch, and survey
CVs <- list("lenage"=0.1, "fishery"=0.01, "survey"=0.1)

#Month of survey/fishing
survey_month <- 7
fishing_month <- 1

#Years to survey, assuming survey is once per year
#survey_years <- survey_sample_full[(burnin+1):(burnin+nyears)] # my original
survey_years <- survey_sample_full[burnin:(burnin+nyears-1)] #from Christine's new sardine_config.R

#read time series data
survObsBiom <- readRDS(file.path(d.name, paste0(scenario.name, "surveyB.rds")))
truecatchbio_ss <- readRDS(file.path(d.name, paste0(scenario.name, "fishCatch.rds")))

survObsBiom <- survObsBiom[[1]]
truecatchbio_ss <- truecatchbio_ss[[1]]

#add this to om_indices function so that this has years when read in
truecatchbio_ss$time <- as.integer(truecatchbio_ss$time/365)

#load dummy dat file
stocksynthesis.data <- r4ss::SS_readdat_3.30(paste0("./inst/extdata/",
model_dir,
datfile_name))

#Test writing CPUE in biomass
  stocksynthesis.data <- atlantisom::SS_write_ts(ss_data_list = stocksynthesis.data,
                ts_data = list(survObsBiom$atoutput[fish_years],
                  truecatchbio_ss$atoutput[truecatchbio_ss$time %in% fish_years]),
                CVs = c(CVs$survey,
                        CVs$fishery),
                data_years = list((survObsBiom$time[fish_years]-survey_sample_time)/timestep+1,             fish_years),
            sampling_month = list(rep(survey_month,nyears),
                                rep(fishing_month,nyears)),
                units = c("biomass","biomass"),
                data_type=c("CPUE","catch"),
                fleets = c(2,1))

#test  
stocksynthesis.data$CPUE

stocksynthesis.data$catch

#read in comp data
age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsAgeComp.rds")))
age_comp_data <- age_comp_data[[1]]

fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
fish_age_comp <- fish_age_comp[[1]]

#Get the age bins
age_bin_names <- names(stocksynthesis.data$agecomp)[10:length(names(stocksynthesis.data$agecomp))]
age_bins <- sub("a","",age_bin_names)

require(maditr)
# add dependency on maditr::dcast to atlantisom
age_comp_flat <- atlantisom::reformat_compositions(age_comp_data,
                     round.places = 4,
                     comp_type = "agecomp")

## Write age composition data for survey
stocksynthesis.data <- atlantisom::SS_write_comps(ss_data_list = stocksynthesis.data,
               comp_matrix = list(age_comp_flat[burnin:(burnin+nyears-1),]),
               data_rows = list(stocksynthesis.data$styr:(stocksynthesis.data$styr+nyears-1)),
               sampling_month = list(rep(survey_month,nyears)),
               data_type = c("agecomp"),
               fleet_number = c(1),
               bins = list(age_bins),
               caal_bool = c(FALSE))
stocksynthesis.data$agecomp

#length comps
len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "survObsLenComp.rds")))
fish_len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsLenComp.rds")))

len_comp_data <- len_comp_data[[1]]
fish_len_comp_data <- fish_len_comp_data[[1]]

#add this to om_indices function so that this has years when read in
fish_len_comp_data$time <- as.integer(floor(fish_len_comp_data$time/fstepperyr))
fish_age_comp$time <- as.integer(floor(fish_age_comp$time/fstepperyr))

if(fstepperyr>1){
  fish_len_comp_anndata <- fish_len_comp_data %>%
    #mutate(yr = floor(time/fstepperyr)) %>%
    #group_by(species, agecl, lower.bins, upper.bins, time=as.integer(yr)) %>%
    group_by(species, agecl, lower.bins, upper.bins, time) %>% 
    summarise(annnatlength=sum(atoutput)) %>%
    rename(atoutput = annnatlength)
} else {
  fish_len_comp_anndata <- fish_len_comp_data
}

caal_comp_flat <- atlantisom::reformat_compositions(len_comp_data,                                round.places=4,
                comp_type="caalcomp")

#remove burnin
caal_comp_final <- filter(caal_comp_flat,
                         time %in% survey_years)

#Add over age classes to get sample size
len_comp_flat <- atlantisom::reformat_compositions(len_comp_data,
                                  round.places = 0,
                           comp_type="lencomp")
#remove burnin
len_comp_final <- filter(len_comp_flat,
                         time %in% survey_years)

length_bins <- as.integer(names(len_comp_final))
length_bins <- length_bins[!is.na(length_bins)]

# fishery length comps are still 5 timesteps per year
# need to aggregate to annual (done above)
# also,  make effN annual goal/fstepsperyr (done above)
fish_len_comp_flat <- atlantisom::reformat_compositions(fish_len_comp_anndata,
                                  round.places = 0,
                           comp_type="lencomp")

#remove burnin works after adjustment above
fish_len_comp_final <- filter(fish_len_comp_flat,
                         time %in% fish_years)

notbins <- c("time", "nsamp")

# fish_length_bins <- as.integer(names(fish_len_comp_final))
# fish_length_bins <- fish_length_bins[!is.na(fish_length_bins)]

# need to fill empty length bins with 0s to have same bins as survey for SS_write_comps
missing.lengths <- setdiff(length_bins, names(fish_len_comp_final)[!names(fish_len_comp_final) %in% notbins])
fish_len_comp_final[as.character(missing.lengths)] <- 0                    # Add them, filled with '0's
fish_len_comp_final <- fish_len_comp_final[c("time", length_bins, "nsamp")] #

# fishery age comps also 5 timesteps per year
if(fstepperyr>1){
  fish_age_comp_anndata <- fish_age_comp %>%
    #mutate(yr = floor(time/fstepperyr)) %>%
    #group_by(species, agecl, time=as.integer(yr)) %>%
    group_by(species, agecl, time) %>%
    summarise(annnatage=sum(atoutput)) %>%
    rename(atoutput = annnatage)
} else {
  fish_age_comp_anndata <- fish_age_comp
}

fish_age_comp_flat <- atlantisom::reformat_compositions(fish_age_comp_anndata,
                           comp_type="agecomp")

#remove burnin (not necessary?fish comps made with fish_years only) 
fish_age_comp_final <- filter(fish_age_comp_flat,
                         time %in% fish_years)

# #SS_write_comps breaking because fishery age bins start with 2 not 1; extracting bins from fish file may help?
# fish_age_bins <- names(fish_age_comp_flat)[!names(fish_age_comp_flat) %in% notbins]

# that leaves an empty column in data file, so instead fill with 0s
missing.ages <- setdiff(age_bins, names(fish_age_comp_final)[!names(fish_age_comp_final) %in% notbins])
fish_age_comp_final[missing.ages] <- 0                    # Add them, filled with '0's
fish_age_comp_final <- fish_age_comp_final[c("time", age_bins, "nsamp")]

comp_list <- list(caal_comp_final,len_comp_final, fish_age_comp_final, fish_len_comp_final)

apply_month <- list(rep(survey_month, nrow(comp_list[[1]])), 
                    rep(survey_month, nrow(comp_list[[2]])),
                    rep(fishing_month,nrow(comp_list[[3]])),
                    rep(fishing_month,nrow(comp_list[[4]])))

# This now runs by ensuring that survey and fishery compositions have the same bins 
# (filled with 0s for missing bins in fishery relative to survey)

# Write CAAL and length composition data
stocksynthesis.data <- atlantisom::SS_write_comps(ss_data_list = stocksynthesis.data, 
                                      comp_matrix = comp_list, 
                                      data_rows = list((comp_list[[1]]$time-survey_sample_time)/timestep + 1 , (survey_years-survey_sample_time)/timestep + 1,fish_years,fish_years),
                                      sampling_month = apply_month, 
                                      data_type = rep(c("agecomp", "lencomp"),2), 
                                      fleet_number = c(2,2,1,1),  
                                      bins = list(age_bins, 
                                                  length_bins, 
                                                  age_bins, 
                                                  length_bins), 
                                      caal_bool = c(TRUE, rep(FALSE,3)))

head(stocksynthesis.data$lencomp)
head(stocksynthesis.data$agecomp)

#Change length bin structure to match atlantis data
stocksynthesis.data$lbin_vector <- length_bins

#Get correct number of length bins
stocksynthesis.data$N_lbins <- length(length_bins)

#Set lbin_method to 1 - this makes the population length bins match the data bins 
#When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
stocksynthesis.data$lbin_method <- 1
stocksynthesis.data$binwidth <- "#"
stocksynthesis.data$minimum_size <- "#"
stocksynthesis.data$maximum_size <- "#"

#Change minimum sample size to 0.001 for CAAL data (SS won't let it go lower than this)
stocksynthesis.data$age_info$minsamplesize <- rep(0.001,2)

SS_writedat_3.30(stocksynthesis.data, outfile = paste0("./inst/extdata/",model_dir,
datfile_name),
                 overwrite=TRUE)

```