### 2020-05-11 Visualize_atlantisom.Rmd
to do next text

---
title: "Visualize atlantisom"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Relationships between `atlantsom` functions

First attempt is using the package [`DependenciesGraphs`](https://github.com/datastorm-open/DependenciesGraphs).

Other possibilities:

[`visNetwork`](https://github.com/datastorm-open/visNetwork)

[`drake`](https://docs.ropensci.org/drake/)

```{r installation, eval=FALSE}

devtools::install_github("datastorm-open/DependenciesGraphs")
devtools::install_github("datastorm-open/visNetwork") # on CRAN, this is dev version

```

## `DependenciesGraphs`

This is all the functions in `atlantisom`:

```{r depgraph, message=FALSE}

library(DependenciesGraphs)
library(atlantisom)

dep <- envirDependencies("package:atlantisom")
plot(dep)

```

This is just the `om_init` wrapper function:

```{r depgraphominit}

depominit <- funDependencies("package:atlantisom", "om_init")
plot(depominit)

```

While these map dependencies between functions, it would be nice to see what output of atlantis goes into atlantisom and what atlantisom outputs into SS3.