### 2020-06-03 InterpolateWeightAtAge.Rmd
testing for 3 species, range of true ages

---
title: "Interpolate the weight at agecl output to annual age"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(atlantisom)
library(ggthemes)

```

## Start with saved wrapper weight at age output for NOBA cod

```{r full-test, message=FALSE, warning=FALSE}

source(here("config/NOBA2Config.R"))
NOBAom_cod <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))

fgs <- NOBAom_cod$funct.group_ss
biolprm <- NOBAom_cod$biol

wtage <- readRDS(file.path(d.name, paste0(scenario.name, "survObsWtAtAge.rds")))
wtage <- wtage[[1]]

wageplot <- ggplot(wtage, aes(time, atoutput)) +
  geom_line(aes(colour = factor(agecl))) +
  theme_tufte() +
  theme(legend.position = "bottom") +
  xlab("model timestep (5 per year)") +
  ylab("average individual weight (g)") +
  labs(subtitle = paste0(scenario.name, " annual mid year sample"))

wageplot + facet_wrap(c("species"), scales="free_y")

```

If we were to do simple interpolation on this output, we can approximate weight at true age class:

```{r interpolate-wtage}
# a function to go into atlantisom, sketched here, reuse calc_stage2age as possible

wtagecl <- wtage
annages <- NOBAom_cod$truenumsage_ss %>%
  group_by(species, time, agecl) %>%
  rename(trueage = agecl) %>%
  summarise(truenatage = sum(atoutput))

#calc_wtTrueage <- function(wtagecl, annages, fgs) {
  # get number of ages per agecl from fgs.file
  #fgs <- atlantisom::load_fgs(dir=d.name, file_fgs = functional.groups.file)
  species.code <- fgs$Code
  #turnedon <- fgs[fgs$IsTurnedOn > 0, ]

  # Figure out the groups that have multiple ages classes in each stage (or
  # cohort) this is from the full model
  multiple_ages <- fgs[fgs$NumAgeClassSize>1, c(1,4,10)]

  # match species in wtagecl_data to those species in the fgs file
  sppwt <- levels(wtagecl$species)
  multiple_ages <- multiple_ages[multiple_ages$Name %in% sppwt, ]

  #now get indices
  names <- multiple_ages$Code
  num_multi_age <- dim(multiple_ages)[1]

  # timesteps from input wt file
  ntimesteps <- length(unique(wtagecl$time))

  # make a dataframe the right dimensions for all age classes
  # has species, time from wtage and trueage, truenatage from annages
  # add agecl column
  # finds the average age of each agecl each timestep based on truenatage
  wtage_out <- annages %>%
    dplyr::semi_join(wtagecl, by = c("species", "time")) %>%
    dplyr::left_join(multiple_ages, by = c("species" = "Name")) %>%
    dplyr::mutate(agecl = as.integer(ceiling(trueage/NumAgeClassSize))) %>%
    dplyr::group_by(species, time, agecl) %>%
    dplyr::mutate(avgage = weighted.mean(trueage, truenatage))

  wtage_avgage <- wtage_out %>%
    select(species, time, agecl, avgage) %>%
    distinct()

  # find weight increment for a timestep
  # muweight for agecl+1 - muweight for agecl
  wtagecl_inc <- wtagecl %>%
    dplyr::arrange(species, time, agecl) %>%
    #dplyr::left_join(multiple_ages, by = c("species" = "Name")) %>%
    dplyr::group_by(species, time) %>%
    dplyr::mutate(increment = case_when(
      agecl==1 ~ atoutput,
      agecl>1 ~ atoutput - dplyr::lag(atoutput)
      )) %>%
    dplyr::left_join(wtage_avgage) %>%
    dplyr::mutate(avgageinc = case_when(
      agecl==1 ~ avgage,
      agecl>1 ~ avgage - dplyr::lag(avgage)
      ))

  # complex, but works for 2 ages/agecl, need to test more, need to fix oldest
  wtage_out <- wtage_out %>%
    dplyr::select(species, time, trueage, agecl, avgage) %>%
    dplyr::left_join(wtagecl_inc) %>%
    dplyr::group_by(species, time) %>%
    dplyr::mutate(wtIntage = case_when(
      trueage==1 ~ trueage/avgage*increment,
      (trueage>1 & trueage<avgage) ~ 
        (1-(avgage-trueage)/avgageinc)*increment + lag(atoutput),
      (trueage>1 & trueage>avgage) ~ 
        ((trueage-avgage)/lead(avgageinc))*lead(increment) + atoutput
        ))

  # plot this--how bad would piecewise linear interpolation be?
  wtageclann <- ggplot(wtagecl_inc, aes(avgage, atoutput)) +
    geom_point() +
    geom_line(aes(colour = factor(time))) + 
    scale_x_continuous(minor_breaks = c(0:20)) +
    theme_tufte() +
    theme(legend.position = "none",
          panel.grid.minor.x = element_line(colour = "grey50"),
          panel.grid.major.x = element_line(colour = "grey50"))

  wtageclann + geom_point(data = wtage_out, 
             mapping = aes(trueage, wtIntage))
  # maybe not so bad

  # fix oldest age

  #clean up to have only standard columns
  finaldata <- data.frame("species" = wtage_out$species,
                          "agecl" = wtage_out$trueage, "polygon" = NA, "layer" = NA,
                          "time" = wtage_out$time, "atoutput" = wtage_out$wtIntage)

  # output full set of muweight at newly created agecl

#}
```

Run wrappers and plot interpolated results for multiple NOBA species:

```{r interp-wtage-test1}

NOBAom <- om_init(here("config/NOBA2config.R"))

# 3 species: 40, 20, 6 age classes
NOBAom_redcodcap <- om_species(c("Redfish", "North_atl_cod", "Capelin"), NOBAom)

rm(NOBAom) #free some memory

```

Fix bug in omdimensions.R; need to vectorize:

```{r omdimensions-bugfix}

omlist_ss <- NOBAom_redcodcap # multispecies call

n_age_classes <- omlist_ss$funct.group_ss$NumCohorts # works
age_classes <- 1:n_age_classes # nope, only first spp, needed in SS3 writing?

age_classes <- sapply(n_age_classes, seq)
names(age_classes)<-survspp

n_annages <- n_age_classes * omlist_ss$funct.group_ss$NumAgeClassSize # works
annages <- 1:n_annages # nope, only first spp, needed in SS3 writing?

annages <- sapply(n_annages, seq)
names(annages)<-survspp

# dimensioning survey in 3sppsurvey.R
survspp <- omlist_ss$species_ss # works
surveffic <- data.frame(species=survspp,
                     efficiency=rep(1.0,length(survspp)))  # works
survselex <- data.frame(species=rep(survspp, each=n_annages), # nope! 
                        agecl=rep(c(1:n_annages),length(survspp)),
                        selex=rep(1.0,length(survspp)*n_annages))
surveffN <- data.frame(species=survspp, effN=rep(1000, length(survspp))) # works
surv_cv <- data.frame(species=survspp, cv=rep(0.1,length(survspp))) # works

#dimensioned for annual ages rather than age classes
survselex <- data.frame(species=rep(survspp, n_annages), #  
                        agecl=unlist(sapply(n_annages,seq)),
                        selex=rep(1.0,sum(n_annages)))

```

New config files needed for multispecies pulls:

`3sppsurvey.R` configures the fishery independent survey and looks like this:

```{r, code = readLines("./config/3sppsurvey.R"), eval=F}
```

`3sppfishery.R` configures the fishery dependent data and looks like this:

```{r, code = readLines("./config/3sppfishery.R"), eval=F}
```

Fix multispecies bug in om_index function:
```{r omindex-bugfix}

# original  
wtage <- data.frame(species=rep(survspp, each=max(age_classes)), #nope
                      agecl=rep(c(1:max(age_classes)),length(survspp)), #nope
                      wtAtAge=rep(1000.0,length(survspp)*max(age_classes))) #nope

# age_classes is now a list 1:n_age_classes by species
# this is still dimensioned by agecl not true age
wtage <- data.frame(species=rep(survspp, n_age_classes), 
                      agecl=unlist(sapply(n_age_classes,seq)), 
                      wtAtAge=rep(1000.0,sum(n_age_classes)))

```

Try index step again:

```{r indcomp-test}

NOBAom_redcodcap_ind <- om_index(usersurvey = here("config/3sppsurvey.R"), 
                           userfishery = here("config/3sppfishery.R"),
                           omlist_ss = NOBAom_redcodcap, 
                           n_reps = 1, 
                           save = TRUE)

```

om_index works, om_comps runs for survey and produces age, length, weight output, produces fishery age output, but fails calling the age2length function.

check fishery observed age comp:

```{r debugcomps, eval=FALSE}

catch_age_comp <- readRDS(file.path(d.name, paste0(scenario.name, "fishObsAgeComp.rds")))
catch_age_comp <- catch_age_comp[[1]]

filter(catch_age_comp, species=="North_atl_cod")
filter(catch_age_comp, species=="Redfish")
filter(catch_age_comp, species=="Capelin")

#snippets from calc_age2length
nums <- filter(catch_age_comp, species=="Capelin")
#calculate mean length at age
mulen <- nums
muweight <- nums

#needed stuff from om_comps
#Get catch weights for length comp calc
  # aggregate true resn per fishery subset design
  catch_aggresnss <- atlantisom::aggregateDensityData(dat = omlist_ss$trueresn_ss,
                                                      time = fishtime,
                                                      species = survspp,
                                                      boxes = fishboxes)

  # aggregate true structn fishery subsetdesign
  catch_aggstructnss <- atlantisom::aggregateDensityData(dat = omlist_ss$truestructn_ss,
                                                         time = fishtime,
                                                         species = survspp,
                                                         boxes = fishboxes)

  #dont sample these, just aggregate them using median
  structn <- atlantisom::sample_fish(catch_aggstructnss, fisheffN, sample = FALSE)

  resn <- atlantisom::sample_fish(catch_aggresnss, fisheffN, sample = FALSE)

  # more snippets from calc_age2length
  # extract rows from structn and resn that match the rows in nums, which are only non-zeroes
resn.ind <- with(resn,paste(species,'.',agecl,'.',polygon,'.',layer,'.',time,sep=""))
num.ind <- with(nums,paste(species,'.',agecl,'.',polygon,'.',layer,'.',time,sep=""))
pick <- match(num.ind,resn.ind)
SRN <- resn$atoutput[pick] + structn$atoutput[pick]

# get weight-length parameters
li_a_use <- biolprm$wl[match(fgs$Code[match(nums$species,fgs$Name)],biolprm$wl[, 1]), 2]
li_b_use <- biolprm$wl[match(fgs$Code[match(nums$species,fgs$Name)],biolprm$wl[, 1]), 3]

#calc mean length and weight at age
mulen$atoutput <- ((biolprm$kgw2d*biolprm$redfieldcn*SRN)/(1000*li_a_use))^(1/li_b_use)
muweight$atoutput <- li_a_use*mulen$atoutput^li_b_use

upper.bins <- 1:maxbin
lower.bins <- c(0,upper.bins[-length(upper.bins)])
lenfreq <- NULL

# well, match does return NA for capelin, and this is what breaks the comparison using 
# the max function (returns NA). adding na.rm=TRUE to line 86
if(maxbin<max(mulen$atoutput)){
  print("Warning: maximum bin size is smaller than the longest fish in the sample. Fish above the maximum bin size will be removed from length compositions.")
}

groups <- unique(as.factor(structn$species)) 
CVlenage = 0.1
# but will NAs break the following?
for (irow in 1:nrow(mulen))
#for (irow in 1:500)
  {
  group <- nums$species[irow]
  igroup <- which(groups==group)
  box <- nums$polygon[irow]
  layer <- nums$layer[irow]
  time <- nums$time[irow]
  age <- nums$agecl[irow]

  sigma = sqrt(log((CVlenage^2)+1))
  muuse <- log(mulen$atoutput[irow]) - 0.5*(sigma^2)
  CumFracperbin <- plnorm(upper.bins,muuse,sigma)
  Fracperbin <- c(CumFracperbin[1],diff(CumFracperbin))
  natlength = Fracperbin*nums$atoutput[irow]
  results <- cbind(mulen[irow,],lower.bins,upper.bins, row.names = NULL) # add row.names = NULL?
  results$atoutput <- round(natlength,0)
  #results <- results[results$atoutput>0,]
  lenfreq <- rbind(lenfreq,results) # add row.names = NULL?
}

remove.zeroes = TRUE
#get rid of zero rows.
if (remove.zeroes) lenfreq <- lenfreq[lenfreq$atoutput>0,]

#output. natlength now has additional columns because of need to store length bin info
lenout <- NULL
lenout$mulen <- mulen
lenout$muweight <- muweight
lenout$natlength <- lenfreq

# ok, works but adds NA for Capelin age classes where no numbers were sampled.
# could crank up sample size or just live with this.

```

Try om_comps step again:

```{r comps-redo}

NOBAom_redcodcap_comp <- om_comps(usersurvey = here("config/3sppsurvey.R"), 
                           userfishery = here("config/3sppfishery.R"),
                           omlist_ss = NOBAom_redcodcap, 
                           n_reps = 1, 
                           save = TRUE)
```