### 2020-06-10 atlantisom_landingpage.Rmd
add detailed data flow section

---
title: "atlantisom: simulated ecosystem datasets for assessment model performance testing"
author: "Sarah Gaichas, Christine Stawitz, Kelli Johnson, Alexander Keth, Allan Hicks, Sean Lucey, Emma Hodgson, and Gavin Fay"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}

htmltools::img(src = knitr::image_uri(here::here("docs/images/atlantis-logo2.jpg")), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; height:175px; padding-right:20px;')

knitr::opts_chunk$set(echo = F,
                      warning = F,
                      message = F)
#Plotting and data libraries
library(tidyverse)
library(here)
library(ggrepel)
library(stringr)

library(atlantisom)

data.dir <- here::here("data")

```

## Overview

The purpose of `atlantisom` is to use existing [Atlantis](https://github.com/Atlantis-Ecosystem-Model/Atlantis_example_and_instructions) ecosystem model output to generate input datasets for a variety of simpler population models, so that the performance of these models can be evaluated against known (simulated) ecosystem dynamics. Atlantis models simulate a wide range of physical and ecological processes, include a full food web, and can be run using different climate forcing, fishing, and other scenarios.

Users of `atlantisom` specify fishery independent and fishery dependent sampling in space and time, as well as species-specific catchability, selectivty, and other observation processes to simulate survey and fishery "data" from stored Atlantis scenario output. `atlantisom` outputs internally consistent multispecies and ecosystem datasets with known observation error characteristics for use in individual model performance testing, comparing performance of alternative models, and performance testing of model ensembles against “true” Atlantis outputs.

Development of `atlantisom` began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA. Substantial progress since then was made possible by a NOAA NMFS International Fellowship and the Institute of Marine Research in Bergen, Norway.

## Get `atlantisom`  https://github.com/r4atlantis/atlantisom

```{r getit, echo=TRUE, eval=FALSE}
# install.packages("devtools")
devtools::install_github("r4atlantis/atlantisom")
```

## Detailed data flow

Have these Atlantis input files in a directory:

- The .bgm file defining model geometry
- The initial conditions/biomass pools .nc 
- The functional groups .csv
- The fishery groups .csv
- The biology .prm
- The run prm (.xml version)

At present,`atlantisom` uses the following Atlantis output files, where [...] is a model-specific prefix. They should be in the same directory as the input files above:

- [...]BiomIndx.txt
- [...]Catch.txt
- [...]CatchPerFishery.txt
- [...]DietCheck.txt
- [...]YOY.txt
- [...].nc
- [...]CATCH.nc
- [...]PROD.nc
- [...]ANNAGEBIO.nc (if available)
- [...]ANNAGECATCH.nc (if available)

Get the truth:
![](docs/images/atlantisomDataFlow_truth-link.png)

Generate the data:
![](docs/images/atlantisomDataFlow_data.png)

## Background

Assessing the ability of models to predict key processes is essential if the models are to be used in decision-making dependent on those processes. Skill assessment compares model predictions with observations; good agreement between predictions and observations indicates high skill (see, e.g. [Stow et al 2009](https://www.sciencedirect.com/science/article/abs/pii/S0924796308001103)). However, in many natural resrouce management contexts, the observations themselves are highly uncertain, so assessing model skill can be difficult. In these cases, creating an observational dataset with a simulation model can provide alternative information for model skill assessment.

### End-to-end ecosystem operating models as dataset generators

Atlantis is a spatially resolved mechanistic end-to-end ecosystem modeling framework: [Fulton et al. 2011](https://onlinelibrary.wiley.com/doi/full/10.1111/j.1467-2979.2011.00412.x), [Fulton and Smith 2004](https://www.ajol.info/index.php/ajms/article/view/33182). Atlantis models have been implemented for regional ecosystems around the world, including:

**Norwegian-Barents Sea**

[Hansen et al. 2016](https://www.imr.no/filarkiv/2016/04/fh-2-2016_noba_atlantis_model_til_web.pdf/nn-no), 
[2018](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0210419)

![NoBa model areas](docs/images/journal.pone.0210419.g001.png)

**California Current**

[Marshall et al. 2017](https://onlinelibrary.wiley.com/doi/full/10.1111/gcb.13594), 
[Kaplan et al. 2017](https://www.sciencedirect.com/science/article/pii/S0304380016308262?via%3Dihub)

![CCA model areas](docs/images/CCAspatial.png)

Atlantis models can incoroporate physical drivers from global change projections and simulate complex biological responses throughout the ecosystem:  [Hodgson et al. 2018](https://www.sciencedirect.com/science/article/pii/S0304380018301856?via%3Dihub), 
[Olsen et al. 2018](https://www.frontiersin.org/articles/10.3389/fmars.2018.00064/full)

## Why use Atlantis?

-  Mechanistic processes create internally consistent "truth"
-  Include cumulative effects of multiple processes:
    -  Climate drivers
    -  Species interactions 
    -  Spatial and seasonal variability
    -  Fisheries
    -  Oil spills, red tide, anything else Atlantis can do
-  Implemented for many ecosystems worldwide

## Why generate datasets instead of simulating within Atlantis?

-  Not all analyses need computationally expensive model interaction
-  Faster! 
-  Test many models or model configurations with the same dataset
-  Many dataset realizations from same "truth"; compare:
    -  Different observation error and bias
    -  Changing temporal and spatial survey coverage 
    -  Improved or degraded fishery observations

## How does `atlantisom` make "data" for stock assessments?

The `atlantisom` user must specify uncertainty in assessment "data":

1. Survey specification:

    1. timing and spatial coverage?

    1. which species are captured?

    1. species-specific survey efficiency ("q")?

    1. selectivity at age for each species?

1. Survey uncertainty:

    1. additional observation error (survey cv for index)?

    1. effective sample size for biological samples?

1. Fishery uncertainty:

    1. additional observation error (catch cv for total)?

    1. catch sampled for length/age in all areas?

    1. effective sample size for biological samples?

Using these specifications, `atlantisom` generates survey data from the true biomass outputs of the Atlantis model, and generates fishery data from the true fishery outputs of the Atlantis model. The simulated data can then be used as inputs to a variety of single species or multispecies assessment models.

## Make Atlantis output into assessment model input

Example atlantisom workflows:

1. Get true biomass, abundance, age composition, length composition, weight at age, fishery catch, fishery catch at age, fishery length composition, and fishery weight age age for a "sardine-like species": https://sgaichas.github.io/poseidon-dev/FullSardineTruthEx.html

1. Format these outputs and get other life history parameters for input into a stock assessment model ([Stock Synthesis](https://www.sciencedirect.com/science/article/abs/pii/S0165783612003293), using [`r4ss`](https://github.com/r4ss)):
https://sgaichas.github.io/poseidon-dev/CreateStockSynthesis.html

1. Get true and observed input data, format inputs, and run the assessment model: https://sgaichas.github.io/poseidon-dev/SardinesHakeatlantisom2SStest.html

1. In progress: compare assessment results with truth:
https://sgaichas.github.io/poseidon-dev/SkillAssessInit.html

1. Simplified dataset extraction with wrapper functions:
https://sgaichas.github.io/poseidon-dev/NOBAcod.html

## Visualizing `atlantisom` outputs

Survey census test NOBA 
```{r ex-output1, echo = F, warning = F, message=F, fig.align = "center"}
knitr::include_graphics("docs/images/NOBAcensus.png")
```

True length composition NOBA
```{r ex-output3, echo = F, warning = F, message=F, fig.align = "center"}
knitr::include_graphics("docs/images/NOBAherringtruelf.png")
```

Standard survey test CCA
```{r ex-output2, echo = F, warning = F, message=F, fig.align = "center"}
knitr::include_graphics("docs/images/CCsurveyex.png")
```

Survey length composition CCA
```{r ex-output4, echo = F, warning = F, message=F, fig.align = "center"}
knitr::include_graphics("docs/images/CCV3herringlfcomp.png")
```

## Example: a "sardine" assessment

Need: assessment model data inputs and life history parameters

*(model based on actual Sardine assessment in Stock Synthesis 3)*

Data:

  * survey biomass index 
  * survey length composition 
  * survey age composition (conditional catch at age)
  * fishery catch (tons) 
  * fishery length composition 
  * fishery age composition

Parameters:

  * natural mortality (from total mortality)
  * growth curve (from survey length at age)
  * maturity at age (true)
  * unfished recruitment and steepness (true)
  * weight-length curve (true)

## A "sardine" assessment: setup

  * California Current Atlantis run with and without climate signal
  * Input data generated (e.g. sardine survey, below in green)
  * Parameters derived; simpler recruitment distribution

```{r sardsurv, echo = F, warning = F, message=F, out.width = "90%", fig.align = "left"}
knitr::include_graphics("docs/images/CCsurvObsBiom.png")
```

## A "sardine" assessment: fits to data

![survey index fit](docs/images/index2_cpuefit_Survey.png)

![length fit](docs/images/comp_lenfit_flt2mkt0_page1.png)

![catch at age fit 1](docs/images/comp_agefit_flt1mkt0_page1.png)
![catch at age fit 2](docs/images/comp_agefit_flt1mkt0_page2.png)

## A "sardine" assessment: skill? (proof of concept)

Biomass
```{r skillB, echo = F, warning = F, message=F, fig.align = "center"}
knitr::include_graphics("docs/images/trueBss3B.png")
```

Fishing mortality
```{r skillF, echo = F, warning = F, message=F, fig.align = "center"}
knitr::include_graphics("docs/images/trueFss3F.png")
```

Recruitment
```{r skillR, echo = F, warning = F, message=F, fig.align = "center"}
knitr::include_graphics("docs/images/trueRss3R.png")
```

Key: <span style="color:teal">True</span> <span style="color:red">SS3 estimate</span>

## More information

* [Atlantis Model Documentation](https://github.com/Atlantis-Ecosystem-Model/Atlantis_example_and_instructions)

* [atlantisom R package](https://github.com/r4atlantis/atlantisom)

* [Testing atlantisom](https://github.com/sgaichas/poseidon-dev)