### 2022-06-30 NewCCASardines.Rmd
add new CC run

---
title: "New CCA sardines"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)  
library(atlantisom)
library(ggthemes)
library(FSA)

```

A test run for sardines with the updated California Current model, run 44:

```{r full-test, message=FALSE, warning=FALSE, eval=FALSE}

CCom <- atlantisom::om_init(here("config/CCConfigv4.R"))

CCom_sard <- atlantisom::om_species(c("Pacific_sardine"), CCom)

CCom_sard_ind <- atlantisom::om_index(usersurvey = here("config/sardinesurvey.R"), 
                           userfishery = here("config/sardinefishery.R"),
                           omlist_ss = CCom_sard, 
                           n_reps = 1, 
                           save = TRUE)

CCom_sard_comp <- atlantisom::om_comps(usersurvey = here("config/sardinesurvey.R"), 
                           userfishery = here("config/sardinefishery.R"),
                           omlist_ss = CCom_sard, 
                           n_reps = 1, 
                           save = TRUE)

```

Diagnose why survey functions resulting in NA outputs. Found it: change `sapply` to `lapply` in omdimensions to make the survey selectivity dataframe. The `sapply` function behaves inconsistently when the age classes are all 10, it orients the list incorrectly; `lapply` keeps the structure correct. This only worked for NOBA because I selected a species with fewer than 10 age classes.

```{r testomindex, eval=FALSE}

usersurvey = here("config/sardinesurvey.R")

omlist_ss <- CCom_sard

#this is om_index

#one script for dimension parameters to be used in multiple functions
  source("config/omdimensions.R", local = TRUE)

  # user options for survey--default is a census with mid-year sample
  # allows muliple surveys
  survObsBiomBs <- list()

  for (s in usersurvey)
  {
    source(s, local = TRUE)

    #biomass based fishery independent survey index
    # this uses result$biomass_ages to sample biomass directly, no need for wt@age est
    survey_B <- atlantisom::create_survey(dat = omlist_ss$truebio_ss,
                                          time = survtime,
                                          species = survspp,
                                          boxes = survboxes,
                                          effic = surveffic,
                                          selex = survselex)

    # call sample_survey_biomass with a bunch of 1000s for weight at age
    # in the code it multiplies atoutput by wtatage/1000 so this allows us to use
    # biomass directly
    wtage <- data.frame(species=rep(names(age_classes), n_age_classes),
                        agecl=unlist(sapply(n_age_classes,seq)),
                        wtAtAge=rep(1000.0,sum(n_age_classes)))

    # this is the step to repeat n_reps time if we want different realizations
    # of the same survey design specified above; only observation error differs
    # using the census cv of 0 will produce identical reps!
    survObsBiomB <- list()
    for(i in 1:n_reps){
      survObsBiomB[[i]] <- atlantisom::sample_survey_biomass(survey_B, surv_cv, wtage)
    }

    #save survey indices, takes a long time to generate with lots of reps/species
    if(save){
      saveRDS(survObsBiomB, file.path(d.name, paste0(scenario.name, "_",
                                                     survey.name, "surveyB.rds")))
    }

    survObsBiomBs[[survey.name]] <- survObsBiomB
  }

  #configure the fishery, a default is in config/fisherycensus.R
  #fishery configuration can specify only area and time of observation
  #fishery species inherited from omlist_ss
  #this is total catch not by fleet, so only one "fishery"
  source(userfishery, local = TRUE)

  #we are not currently subsetting fishery catch because we cannot correct catch.nc
  #  instead the catch in biomass from catch.txt is read in for the index
  #  we do not apply any cv to this, but we could this way (default cv=0)

  fishObsCatchB <- list()
  for(i in 1:n_reps){
    fishObsCatchB[[i]] <- atlantisom::sample_fishery_totcatch(omlist_ss$truecatchbio_ss, fish_cv)
  }

  if(save){
    saveRDS(fishObsCatchB, file.path(d.name, paste0(scenario.name, "_",
                                                    fishery.name, "fishCatch.rds")))
  }

  indices <- list("survObsBiomB" = survObsBiomBs,
                  "fishObsCatchB" = fishObsCatchB)

  return(indices)
}

```

Diagnose why comps aren't working. Survey worked, fishery didn't because the burnin period was longer than this run. Changed to 0 in the sardinefishery.R config file and runs.

```{r testomcomps, eval=FALSE}

usersurvey  <-  here("config/sardinesurvey.R")
userfishery <- here("config/sardinefishery.R")

omlist_ss <- CCom_sard

n_reps <- 1

#this is om_comps

 #one script for dimension parameters to be used in multiple functions
  source("config/omdimensions.R", local = TRUE)

  # user options for survey--default is a census with mid-year sample
  # allows muliple surveys
  age_comp_datas <- list()
  survObsLenComps <- list()
  survObsWtAtAges <- list()

  for (s in usersurvey)
  {
    source(s, local = TRUE)

    #numbers based fishery independent survey for age and length comps
    # same user specifications as indices
    survey_N <- atlantisom::create_survey(dat = omlist_ss$truenums_ss,
                                          time = survtime,
                                          species = survspp,
                                          boxes = survboxes,
                                          effic = surveffic,
                                          selex = survselex.agecl)

    #Sample fish for age composition
    # if we want replicates for obs error this sample function will generate them
    age_comp_data <- list()
    for(i in 1:n_reps){
      age_comp_data[[i]] <- atlantisom::sample_fish(survey_N, surveffN)
    }

    # save age comps
    if(save){
      saveRDS(age_comp_data, file.path(d.name, paste0(scenario.name, "_",
                                                      survey.name, "survObsAgeComp.rds")))
    }

    #weights needed for weight at age and length comp calcs
    # aggregate true resn per survey design
    survey_aggresn <- atlantisom::aggregateDensityData(dat = omlist_ss$trueresn_ss,
                                                       time = survtime,
                                                       species = survspp,
                                                       boxes = survboxes)

    # aggregate true structn per survey design
    survey_aggstructn <- atlantisom::aggregateDensityData(dat = omlist_ss$truestructn_ss,
                                                          time = survtime,
                                                          species = survspp,
                                                          boxes = survboxes)

    #dont sample these, just aggregate them using median
    structnss <- atlantisom::sample_fish(survey_aggstructn, surveffN, sample = FALSE)

    resnss <- atlantisom::sample_fish(survey_aggresn, surveffN, sample = FALSE)

    #this is all input into the length function, replicates follow age comp reps
    #  separating the length comps from the weight at age here
    survey_lenwt <- list()
    survObsLenComp <- list()
    survObsWtAtAge <- list()

    for(i in 1:n_reps){
      survey_lenwt[[i]] <- atlantisom::calc_age2length(structn = structnss,
                                                       resn = resnss,
                                                       nums = age_comp_data[[i]],
                                                       biolprm = omlist_ss$biol,
                                                       fgs = omlist_ss$funct.group_ss,
                                                       maxbin = maxbin,
                                                       CVlenage = lenage_cv,
                                                       remove.zeroes=TRUE)

      survObsLenComp[[i]] <- survey_lenwt[[i]]$natlength
      survObsWtAtAge[[i]] <- survey_lenwt[[i]]$muweight
    }

    if(save){
      saveRDS(survObsLenComp, file.path(d.name, paste0(scenario.name, "_",
                                                       survey.name, "survObsLenComp.rds")))
      saveRDS(survObsWtAtAge, file.path(d.name, paste0(scenario.name, "_",
                                                       survey.name, "survObsWtAtAge.rds")))
    }
    # add each survey to master list objects for survey data
    age_comp_datas[[survey.name]] <- age_comp_data
    survObsLenComps[[survey.name]] <- survObsLenComp
    survObsWtAtAges[[survey.name]] <- survObsWtAtAge

  }

  #now do fishery comps
  # user options for fishery--default is a census with mid-year sample
  # only one fishery, but multiple fleets possible within it
  source(userfishery, local = TRUE)

  #fishery catch at age each observed timestep summed over observed polygons
  # catch at age by area and timestep
  catch_numbers <-  atlantisom::create_fishery_subset(dat = omlist_ss$truecatchnum_ss,
                                                      time = fishtime,
                                                      species = survspp,
                                                      boxes = fishboxes)

  # if we want replicates for obs error this sample function will generate them
  catch_age_comp <- list()
  for(i in 1:n_reps){
    catch_age_comp[[i]] <- atlantisom::sample_fish(catch_numbers, fisheffN)
  }

  # save fishery age comps
  if(save){
    saveRDS(catch_age_comp, file.path(d.name, paste0(scenario.name, "_",
                                                     fishery.name, "fishObsAgeComp.rds")))
  }

  #Get catch weights for length comp calc
  # aggregate true resn per fishery subset design
  catch_aggresnss <- atlantisom::aggregateDensityData(dat = omlist_ss$trueresn_ss,
                                                      time = fishtime,
                                                      species = survspp,
                                                      boxes = fishboxes)

  # aggregate true structn fishery subsetdesign
  catch_aggstructnss <- atlantisom::aggregateDensityData(dat = omlist_ss$truestructn_ss,
                                                         time = fishtime,
                                                         species = survspp,
                                                         boxes = fishboxes)

  #dont sample these, just aggregate them using median
  catch_structnss <- atlantisom::sample_fish(catch_aggstructnss, fisheffN, sample = FALSE)

  catch_resnss <- atlantisom::sample_fish(catch_aggresnss, fisheffN, sample = FALSE)

  # these fishery lengths and weight at age are each output timestep
  #same structure as above for surveys, replicates follow age comp reps
  #  separating the length comps from the weight at age here
  fishery_lenwt <- list()
  fishObsLenComp <- list()
  fishObsWtAtAge <- list()

  for(i in 1:n_reps){
    fishery_lenwt[[i]] <- atlantisom::calc_age2length(structn = catch_structnss,
                                                      resn = catch_resnss,
                                                      nums = catch_age_comp[[i]],
                                                      biolprm = omlist_ss$biol,
                                                      fgs = omlist_ss$funct.group_ss,
                                                      maxbin = maxbin,
                                                      CVlenage = lenage_cv,
                                                      remove.zeroes=TRUE)

    fishObsLenComp[[i]] <- fishery_lenwt[[i]]$natlength
    fishObsWtAtAge[[i]] <- fishery_lenwt[[i]]$muweight
  }

  if(save){
    saveRDS(fishObsLenComp, file.path(d.name, paste0(scenario.name, "_",
                                                     fishery.name, "fishObsLenComp.rds")))
    saveRDS(fishObsWtAtAge, file.path(d.name, paste0(scenario.name, "_",
                                                     fishery.name, "fishObsWtAtAge.rds")))
  }

  if(!is.null(omlist_ss$truenumsage_ss)){
    #numbers based fishery independent survey for age and length comps
    #allows for mulitple surveys
    annage_comp_datas <- list()

    for (s in usersurvey)
    {
      source(s, local = TRUE)

      # same user specifications as indices
      survey_annageN <- atlantisom::create_survey(dat = omlist_ss$truenumsage_ss,
                                                  time = survtime,
                                                  species = survspp,
                                                  boxes = survboxes,
                                                  effic = surveffic,
                                                  selex = survselex)
      #Sample fish for age composition
      # if we want replicates for obs error this sample function will generate them
      annage_comp_data <- list()
      for(i in 1:n_reps){
        annage_comp_data[[i]] <- atlantisom::sample_fish(survey_annageN, surveffN)
      }

      # save survey annual age comps
      if(save){
        saveRDS(annage_comp_data, file.path(d.name, paste0(scenario.name, "_",
                                                           survey.name, "survObsFullAgeComp.rds")))
      }
      annage_comp_datas[[survey.name]] <- annage_comp_data
    }

  }else{annage_comp_datas <- NULL}

  if(!is.null(omlist_ss$truecatchage_ss)){
    #fishery catch at age each observed timestep summed over observed polygons
    # catch at age by area and timestep
    catch_annagenumbers <-  atlantisom::create_fishery_subset(dat = omlist_ss$truecatchage_ss,
                                                        time = fishtime,
                                                        species = survspp,
                                                        boxes = fishboxes)

    # if we want replicates for obs error this sample function will generate them
    # WARNING THIS AGGREGATES ACROSS FLEETS
    # TODO: need to change sample_fish to fix
    catch_annage_comp <- list()
    for(i in 1:n_reps){
      catch_annage_comp[[i]] <- atlantisom::sample_fish(catch_annagenumbers, fisheffN)
    }

    # save fishery annual age comps
    if(save){
    saveRDS(catch_annage_comp, file.path(d.name, paste0(scenario.name,"_",
                                                        fishery.name, "fishObsFullAgeComp.rds")))
    }
  }else{catch_annage_comp <- NULL}

  # call interpolate weight at age function to get survObsFullWtAtAge
  if(!is.null(omlist_ss$truenumsage_ss)){
    interp_survWtAtAges <- list()
    for (s in usersurvey)
    {
      source(s, local = TRUE)

      interp_survWtAtAge <- list()
      for(i in 1:n_reps){
        interp_survWtAtAge[[i]] <- calc_avgwtstage2age(wtagecl = survObsWtAtAges[[survey.name]][[i]],
                                                       annages = omlist_ss$truenumsage_ss,
                                                       fgs = omlist_ss$funct.group_ss)
      }
      if(save){
        saveRDS(interp_survWtAtAge, file.path(d.name, paste0(scenario.name, "_",
                                                             survey.name, "survObsFullWtAtAge.rds")))
      }
      interp_survWtAtAges[[survey.name]] <- interp_survWtAtAge
    }
  }else{interp_survWtAtAges <- NULL}

  # do we want fishery average weight at true age too? why not
  # call interpolate weight at age function to get fishObsFullWtAtAge
  # WARNING currently aggregates out fleet info, but no fleets in aggregate wtage
  if(!is.null(omlist_ss$truecatchage_ss)){
    interp_fishWtAtAge <- list()
    for(i in 1:n_reps){
      interp_fishWtAtAge[[i]] <- calc_avgwtstage2age(wtagecl = fishObsWtAtAge[[i]],
                                                     annages = omlist_ss$truecatchage_ss,
                                                     fgs = omlist_ss$funct.group_ss)
    }
    if(save){
      saveRDS(interp_fishWtAtAge, file.path(d.name, paste0(scenario.name,"_",
                                                           fishery.name, "fishObsFullWtAtAge.rds")))
    }
  }else{interp_fishWtAtAge <- NULL}

  comps <- list("survObsAgeComp" = age_comp_datas,
                "survObsLenComp" = survObsLenComps,
                "survObsWtAtAge" = survObsWtAtAges,
                "fishObsAgeComp" = catch_age_comp,
                "fishObsLenComp" = fishObsLenComp,
                "fishObsWtAtAge" = fishObsWtAtAge,
                "survObsFullAgeComp" = annage_comp_datas,
                "fishObsFullAgeComp" = catch_annage_comp,
                "survObsFullWtAtAge" = interp_survWtAtAges,
                "fishObsFullWtAtAge" = interp_fishWtAtAge
                )

  return(comps)
}

```

### Wrapper test results {.tabset}

#### Biomass index
```{r bioind-1}

source(here("config/CCConfigv4.R"))

omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))

source(here("config/omdimensions.R"))
source(here("config/sardinesurvey.R"))
source(here("config/sardinefishery.R"))

#read time series data
survObsBiom <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                survey.name,
                                                "surveyB.rds")))

survObsBiom <- survObsBiom[[1]]

plotB <-ggplot() +
  geom_line(data=survObsBiom, aes(x=time/stepperyr,y=atoutput, color="survey Biomass"), 
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

plotB +
  facet_wrap(~species, scales="free")

```

#### Catch time series
```{r catchind-1}
#read time series data
catchbio_ss <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                fishery.name, "fishCatch.rds")))

catchbio_ss <- catchbio_ss[[1]]

plotC <-ggplot() +
  geom_line(data=catchbio_ss, aes(x=time/365,y=atoutput, color="observed Catch"), 
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

plotC +
  facet_wrap(~species, scales="free")

```

#### Survey length composition
```{r lencomp-1}

#length comps
len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                survey.name, "survObsLenComp.rds")))
fish_len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                fishery.name,"fishObsLenComp.rds")))

len_comp_data <- len_comp_data[[1]]
fish_len_comp_data <- fish_len_comp_data[[1]]

#add this to om_indices function so that this has years when read in
fish_len_comp_data$time <- as.integer(floor(fish_len_comp_data$time/fstepperyr))

len <- len_comp_data

  lfplot <- ggplot(len, aes(upper.bins)) +
    geom_bar(aes(weight = atoutput)) +
    theme_tufte() +
    labs(subtitle = paste(scenario.name,
                          len$species))

  lfplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")

```

#### Survey age composition (age classes)
```{r agecomp-1}

#read in comp data
age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                survey.name, "survObsAgeComp.rds")))
age_comp_data <- age_comp_data[[1]]

Natage <- age_comp_data

Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
    geom_point() +
    theme_tufte() +
    labs(subtitle = paste(scenario.name,
                          Natage$species))

  Natageplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
```

#### Survey weight at age (age classes)
```{r wtageclass-1}

wtage <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                survey.name, "survObsWtAtAge.rds")))
wtage <- wtage[[1]]

wageplot <- ggplot(wtage, aes(time, atoutput)) +
  geom_line(aes(colour = factor(agecl))) +
  theme_tufte() +
  theme(legend.position = "bottom") +
  xlab("model timestep (5 per year)") +
  ylab("average individual weight (g)") +
  labs(subtitle = paste0(scenario.name, " annual mid year sample"))

wageplot + facet_wrap(c("species"), scales="free_y")
```

#### Fishery age composition (age classes)
```{r fishagecomp-1}

#read in comp data
fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                fishery.name, "fishObsAgeComp.rds")))
fish_age_comp <- fish_age_comp[[1]]

#add this to om_indices function so that this has years when read in
fish_age_comp$time <- fish_age_comp$time/fstepperyr

Natage <- fish_age_comp

Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
    geom_point() +
    theme_tufte() +
    labs(subtitle = paste(scenario.name,
                          Natage$species))

  Natageplot + facet_wrap(~time, ncol=6, scales="free_y")
```

### {-}