### 2022-09-16 NewCCASardines.Rmd
update CCA results WIP

---
title: "New CCA sardines, Sept 2022"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)  
library(atlantisom)
library(ggthemes)
library(FSA)

```

A test run for sardines with the updated California Current model, September 2022:

```{r full-test, message=FALSE, warning=FALSE, eval=FALSE}

CCom <- atlantisom::om_init(here("config/CCConfigSep22.R"))

CCom_sard <- atlantisom::om_species(c("Pacific_sardine"), CCom)

CCom_sard_ind <- atlantisom::om_index(usersurvey = here("config/sardinesurvey.R"), 
                           userfishery = here("config/sardinefishery.R"),
                           omlist_ss = CCom_sard, 
                           n_reps = 1, 
                           save = TRUE)

CCom_sard_comp <- atlantisom::om_comps(usersurvey = here("config/sardinesurvey.R"), 
                           userfishery = here("config/sardinefishery.R"),
                           omlist_ss = CCom_sard, 
                           n_reps = 1, 
                           save = TRUE)

```

Diagnose why survey functions resulting in NA outputs. Found it: change `sapply` to `lapply` in omdimensions to make the survey selectivity dataframe. The `sapply` function behaves inconsistently when the age classes are all 10, it orients the list incorrectly; `lapply` keeps the structure correct. This only worked for NOBA because I selected a species with fewer than 10 age classes.

```{r testomindex, eval=FALSE}

usersurvey = here("config/sardinesurvey.R")

omlist_ss <- CCom_sard

#this is om_index

#one script for dimension parameters to be used in multiple functions
  source("config/omdimensions.R", local = TRUE)

  # user options for survey--default is a census with mid-year sample
  # allows muliple surveys
  survObsBiomBs <- list()

  for (s in usersurvey)
  {
    source(s, local = TRUE)

    #biomass based fishery independent survey index
    # this uses result$biomass_ages to sample biomass directly, no need for wt@age est
    survey_B <- atlantisom::create_survey(dat = omlist_ss$truebio_ss,
                                          time = survtime,
                                          species = survspp,
                                          boxes = survboxes,
                                          effic = surveffic,
                                          selex = survselex)

    # call sample_survey_biomass with a bunch of 1000s for weight at age
    # in the code it multiplies atoutput by wtatage/1000 so this allows us to use
    # biomass directly
    wtage <- data.frame(species=rep(names(age_classes), n_age_classes),
                        agecl=unlist(sapply(n_age_classes,seq)),
                        wtAtAge=rep(1000.0,sum(n_age_classes)))

    # this is the step to repeat n_reps time if we want different realizations
    # of the same survey design specified above; only observation error differs
    # using the census cv of 0 will produce identical reps!
    survObsBiomB <- list()
    for(i in 1:n_reps){
      survObsBiomB[[i]] <- atlantisom::sample_survey_biomass(survey_B, surv_cv, wtage)
    }

    #save survey indices, takes a long time to generate with lots of reps/species
    if(save){
      saveRDS(survObsBiomB, file.path(d.name, paste0(scenario.name, "_",
                                                     survey.name, "surveyB.rds")))
    }

    survObsBiomBs[[survey.name]] <- survObsBiomB
  }

  #configure the fishery, a default is in config/fisherycensus.R
  #fishery configuration can specify only area and time of observation
  #fishery species inherited from omlist_ss
  #this is total catch not by fleet, so only one "fishery"
  source(userfishery, local = TRUE)

  #we are not currently subsetting fishery catch because we cannot correct catch.nc
  #  instead the catch in biomass from catch.txt is read in for the index
  #  we do not apply any cv to this, but we could this way (default cv=0)

  fishObsCatchB <- list()
  for(i in 1:n_reps){
    fishObsCatchB[[i]] <- atlantisom::sample_fishery_totcatch(omlist_ss$truecatchbio_ss, fish_cv)
  }

  if(save){
    saveRDS(fishObsCatchB, file.path(d.name, paste0(scenario.name, "_",
                                                    fishery.name, "fishCatch.rds")))
  }

  indices <- list("survObsBiomB" = survObsBiomBs,
                  "fishObsCatchB" = fishObsCatchB)

  return(indices)
}

```

Diagnose why comps aren't working. Survey worked, fishery didn't because the burnin period was longer than this run. Changed to 0 in the sardinefishery.R config file and runs.

```{r testomcomps, eval=FALSE}

usersurvey  <-  here("config/sardinesurvey.R")
userfishery <- here("config/sardinefishery.R")

omlist_ss <- CCom_sard

n_reps <- 1

#this is om_comps

 #one script for dimension parameters to be used in multiple functions
  source("config/omdimensions.R", local = TRUE)

  # user options for survey--default is a census with mid-year sample
  # allows muliple surveys
  age_comp_datas <- list()
  survObsLenComps <- list()
  survObsWtAtAges <- list()

  for (s in usersurvey)
  {
    source(s, local = TRUE)

    #numbers based fishery independent survey for age and length comps
    # same user specifications as indices
    survey_N <- atlantisom::create_survey(dat = omlist_ss$truenums_ss,
                                          time = survtime,
                                          species = survspp,
                                          boxes = survboxes,
                                          effic = surveffic,
                                          selex = survselex.agecl)

    #Sample fish for age composition
    # if we want replicates for obs error this sample function will generate them
    age_comp_data <- list()
    for(i in 1:n_reps){
      age_comp_data[[i]] <- atlantisom::sample_fish(survey_N, surveffN)
    }

    # save age comps
    if(save){
      saveRDS(age_comp_data, file.path(d.name, paste0(scenario.name, "_",
                                                      survey.name, "survObsAgeComp.rds")))
    }

    #weights needed for weight at age and length comp calcs
    # aggregate true resn per survey design
    survey_aggresn <- atlantisom::aggregateDensityData(dat = omlist_ss$trueresn_ss,
                                                       time = survtime,
                                                       species = survspp,
                                                       boxes = survboxes)

    # aggregate true structn per survey design
    survey_aggstructn <- atlantisom::aggregateDensityData(dat = omlist_ss$truestructn_ss,
                                                          time = survtime,
                                                          species = survspp,
                                                          boxes = survboxes)

    #dont sample these, just aggregate them using median
    structnss <- atlantisom::sample_fish(survey_aggstructn, surveffN, sample = FALSE)

    resnss <- atlantisom::sample_fish(survey_aggresn, surveffN, sample = FALSE)

    #this is all input into the length function, replicates follow age comp reps
    #  separating the length comps from the weight at age here
    survey_lenwt <- list()
    survObsLenComp <- list()
    survObsWtAtAge <- list()

    for(i in 1:n_reps){
      survey_lenwt[[i]] <- atlantisom::calc_age2length(structn = structnss,
                                                       resn = resnss,
                                                       nums = age_comp_data[[i]],
                                                       biolprm = omlist_ss$biol,
                                                       fgs = omlist_ss$funct.group_ss,
                                                       maxbin = maxbin,
                                                       CVlenage = lenage_cv,
                                                       remove.zeroes=TRUE)

      survObsLenComp[[i]] <- survey_lenwt[[i]]$natlength
      survObsWtAtAge[[i]] <- survey_lenwt[[i]]$muweight
    }

    if(save){
      saveRDS(survObsLenComp, file.path(d.name, paste0(scenario.name, "_",
                                                       survey.name, "survObsLenComp.rds")))
      saveRDS(survObsWtAtAge, file.path(d.name, paste0(scenario.name, "_",
                                                       survey.name, "survObsWtAtAge.rds")))
    }
    # add each survey to master list objects for survey data
    age_comp_datas[[survey.name]] <- age_comp_data
    survObsLenComps[[survey.name]] <- survObsLenComp
    survObsWtAtAges[[survey.name]] <- survObsWtAtAge

  }

  #now do fishery comps
  # user options for fishery--default is a census with mid-year sample
  # only one fishery, but multiple fleets possible within it
  source(userfishery, local = TRUE)

  #fishery catch at age each observed timestep summed over observed polygons
  # catch at age by area and timestep
  catch_numbers <-  atlantisom::create_fishery_subset(dat = omlist_ss$truecatchnum_ss,
                                                      time = fishtime,
                                                      species = survspp,
                                                      boxes = fishboxes)

  # if we want replicates for obs error this sample function will generate them
  catch_age_comp <- list()
  for(i in 1:n_reps){
    catch_age_comp[[i]] <- atlantisom::sample_fish(catch_numbers, fisheffN)
  }

  # save fishery age comps
  if(save){
    saveRDS(catch_age_comp, file.path(d.name, paste0(scenario.name, "_",
                                                     fishery.name, "fishObsAgeComp.rds")))
  }

  #Get catch weights for length comp calc
  # aggregate true resn per fishery subset design
  catch_aggresnss <- atlantisom::aggregateDensityData(dat = omlist_ss$trueresn_ss,
                                                      time = fishtime,
                                                      species = survspp,
                                                      boxes = fishboxes)

  # aggregate true structn fishery subsetdesign
  catch_aggstructnss <- atlantisom::aggregateDensityData(dat = omlist_ss$truestructn_ss,
                                                         time = fishtime,
                                                         species = survspp,
                                                         boxes = fishboxes)

  #dont sample these, just aggregate them using median
  catch_structnss <- atlantisom::sample_fish(catch_aggstructnss, fisheffN, sample = FALSE)

  catch_resnss <- atlantisom::sample_fish(catch_aggresnss, fisheffN, sample = FALSE)

  # these fishery lengths and weight at age are each output timestep
  #same structure as above for surveys, replicates follow age comp reps
  #  separating the length comps from the weight at age here
  fishery_lenwt <- list()
  fishObsLenComp <- list()
  fishObsWtAtAge <- list()

  for(i in 1:n_reps){
    fishery_lenwt[[i]] <- atlantisom::calc_age2length(structn = catch_structnss,
                                                      resn = catch_resnss,
                                                      nums = catch_age_comp[[i]],
                                                      biolprm = omlist_ss$biol,
                                                      fgs = omlist_ss$funct.group_ss,
                                                      maxbin = maxbin,
                                                      CVlenage = lenage_cv,
                                                      remove.zeroes=TRUE)

    fishObsLenComp[[i]] <- fishery_lenwt[[i]]$natlength
    fishObsWtAtAge[[i]] <- fishery_lenwt[[i]]$muweight
  }

  if(save){
    saveRDS(fishObsLenComp, file.path(d.name, paste0(scenario.name, "_",
                                                     fishery.name, "fishObsLenComp.rds")))
    saveRDS(fishObsWtAtAge, file.path(d.name, paste0(scenario.name, "_",
                                                     fishery.name, "fishObsWtAtAge.rds")))
  }

  if(!is.null(omlist_ss$truenumsage_ss)){
    #numbers based fishery independent survey for age and length comps
    #allows for mulitple surveys
    annage_comp_datas <- list()

    for (s in usersurvey)
    {
      source(s, local = TRUE)

      # same user specifications as indices
      survey_annageN <- atlantisom::create_survey(dat = omlist_ss$truenumsage_ss,
                                                  time = survtime,
                                                  species = survspp,
                                                  boxes = survboxes,
                                                  effic = surveffic,
                                                  selex = survselex)
      #Sample fish for age composition
      # if we want replicates for obs error this sample function will generate them
      annage_comp_data <- list()
      for(i in 1:n_reps){
        annage_comp_data[[i]] <- atlantisom::sample_fish(survey_annageN, surveffN)
      }

      # save survey annual age comps
      if(save){
        saveRDS(annage_comp_data, file.path(d.name, paste0(scenario.name, "_",
                                                           survey.name, "survObsFullAgeComp.rds")))
      }
      annage_comp_datas[[survey.name]] <- annage_comp_data
    }

  }else{annage_comp_datas <- NULL}

  if(!is.null(omlist_ss$truecatchage_ss)){
    #fishery catch at age each observed timestep summed over observed polygons
    # catch at age by area and timestep
    catch_annagenumbers <-  atlantisom::create_fishery_subset(dat = omlist_ss$truecatchage_ss,
                                                        time = fishtime,
                                                        species = survspp,
                                                        boxes = fishboxes)

    # if we want replicates for obs error this sample function will generate them
    # WARNING THIS AGGREGATES ACROSS FLEETS
    # TODO: need to change sample_fish to fix
    catch_annage_comp <- list()
    for(i in 1:n_reps){
      catch_annage_comp[[i]] <- atlantisom::sample_fish(catch_annagenumbers, fisheffN)
    }

    # save fishery annual age comps
    if(save){
    saveRDS(catch_annage_comp, file.path(d.name, paste0(scenario.name,"_",
                                                        fishery.name, "fishObsFullAgeComp.rds")))
    }
  }else{catch_annage_comp <- NULL}

  # call interpolate weight at age function to get survObsFullWtAtAge
  if(!is.null(omlist_ss$truenumsage_ss)){
    interp_survWtAtAges <- list()
    for (s in usersurvey)
    {
      source(s, local = TRUE)

      interp_survWtAtAge <- list()
      for(i in 1:n_reps){
        interp_survWtAtAge[[i]] <- calc_avgwtstage2age(wtagecl = survObsWtAtAges[[survey.name]][[i]],
                                                       annages = omlist_ss$truenumsage_ss,
                                                       fgs = omlist_ss$funct.group_ss)
      }
      if(save){
        saveRDS(interp_survWtAtAge, file.path(d.name, paste0(scenario.name, "_",
                                                             survey.name, "survObsFullWtAtAge.rds")))
      }
      interp_survWtAtAges[[survey.name]] <- interp_survWtAtAge
    }
  }else{interp_survWtAtAges <- NULL}

  # do we want fishery average weight at true age too? why not
  # call interpolate weight at age function to get fishObsFullWtAtAge
  # WARNING currently aggregates out fleet info, but no fleets in aggregate wtage
  if(!is.null(omlist_ss$truecatchage_ss)){
    interp_fishWtAtAge <- list()
    for(i in 1:n_reps){
      interp_fishWtAtAge[[i]] <- calc_avgwtstage2age(wtagecl = fishObsWtAtAge[[i]],
                                                     annages = omlist_ss$truecatchage_ss,
                                                     fgs = omlist_ss$funct.group_ss)
    }
    if(save){
      saveRDS(interp_fishWtAtAge, file.path(d.name, paste0(scenario.name,"_",
                                                           fishery.name, "fishObsFullWtAtAge.rds")))
    }
  }else{interp_fishWtAtAge <- NULL}

  comps <- list("survObsAgeComp" = age_comp_datas,
                "survObsLenComp" = survObsLenComps,
                "survObsWtAtAge" = survObsWtAtAges,
                "fishObsAgeComp" = catch_age_comp,
                "fishObsLenComp" = fishObsLenComp,
                "fishObsWtAtAge" = fishObsWtAtAge,
                "survObsFullAgeComp" = annage_comp_datas,
                "fishObsFullAgeComp" = catch_annage_comp,
                "survObsFullWtAtAge" = interp_survWtAtAges,
                "fishObsFullWtAtAge" = interp_fishWtAtAge
                )

  return(comps)
}

```

### Wrapper test results {.tabset}

#### Biomass index
```{r bioind-1}

source(here("config/CCConfigSep22.R"))

omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))

source(here("config/omdimensions.R"))
source(here("config/sardinesurvey.R"))
source(here("config/sardinefishery.R"))

#read time series data
survObsBiom <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                survey.name,
                                                "surveyB.rds")))

survObsBiom <- survObsBiom[[1]]

plotB <-ggplot() +
  geom_line(data=survObsBiom, aes(x=time/stepperyr,y=atoutput, color="survey Biomass"), 
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

plotB +
  facet_wrap(~species, scales="free")

```

#### Catch time series
```{r catchind-1}
#read time series data
catchbio_ss <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                fishery.name, "fishCatch.rds")))

catchbio_ss <- catchbio_ss[[1]]

plotC <-ggplot() +
  geom_line(data=catchbio_ss, aes(x=time/365,y=atoutput, color="observed Catch"), 
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

plotC +
  facet_wrap(~species, scales="free")

```

#### Survey length composition
```{r lencomp-1}

#length comps
len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                survey.name, "survObsLenComp.rds")))
fish_len_comp_data <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                fishery.name,"fishObsLenComp.rds")))

len_comp_data <- len_comp_data[[1]]
fish_len_comp_data <- fish_len_comp_data[[1]]

#add this to om_indices function so that this has years when read in
fish_len_comp_data$time <- as.integer(floor(fish_len_comp_data$time/fstepperyr))

len <- len_comp_data

len <- filter(len_comp_data, time %in% c(55:175))

  lfplot <- ggplot(len, aes(upper.bins)) +
    geom_bar(aes(weight = atoutput)) +
    theme_tufte() +
    labs(subtitle = paste(scenario.name,
                          len$species))

  lfplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")

```

#### Survey age composition (age classes)
```{r agecomp-1}

#read in comp data
age_comp_data <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                survey.name, "survObsAgeComp.rds")))
age_comp_data <- age_comp_data[[1]]

Natage <- age_comp_data

Natage <- filter(age_comp_data, time %in% c(150:270))

Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
    geom_point() +
    theme_tufte() +
    labs(subtitle = paste(scenario.name,
                          Natage$species))

  Natageplot + facet_wrap(~time/stepperyr, ncol=6, scales="free_y")
```

#### Survey weight at age (age classes)
```{r wtageclass-1}

wtage <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                survey.name, "survObsWtAtAge.rds")))
wtage <- wtage[[1]]

wageplot <- ggplot(wtage, aes(time, atoutput)) +
  geom_line(aes(colour = factor(agecl))) +
  theme_tufte() +
  theme(legend.position = "bottom") +
  xlab("model timestep (5 per year)") +
  ylab("average individual weight (g)") +
  labs(subtitle = paste0(scenario.name, " annual mid year sample"))

wageplot + facet_wrap(c("species"), scales="free_y")
```

#### Fishery age composition (age classes)
```{r fishagecomp-1}

#read in comp data
fish_age_comp <- readRDS(file.path(d.name, paste0(scenario.name,
                                                "_",
                                                fishery.name, "fishObsAgeComp.rds")))
fish_age_comp <- fish_age_comp[[1]]

#add this to om_indices function so that this has years when read in
fish_age_comp$time <- fish_age_comp$time/fstepperyr

Natage <- fish_age_comp

Natage <- filter(fish_age_comp, time %in% c(30:53))

Natageplot <- ggplot(Natage, aes(x=agecl, y=atoutput)) +
    geom_point() +
    theme_tufte() +
    labs(subtitle = paste(scenario.name,
                          Natage$species))

  Natageplot + facet_wrap(~time, ncol=6, scales="free_y")
```

### {-}

## Write data for Stock Synthesis

To read in the Stock Synthesis files, first ensure you have the latest version of `r4ss`. You will need to ensure the relevant Stock Synthesis files are stored in the inst/extdata/ folder, with each species example in its own folder.

```{r readdata}
#devtools::install_github("r4ss/r4ss", dependencies = FALSE)
require(r4ss)
#source("./config/sardine_config.R") this is above for illustration

#Directory with SS files
model_dir <- "Sardine_SS_files/"

#Name of SS data file
datfile_name <- "sardEM_3_3.dat"

stocksynthesis.data <- r4ss::SS_readdat_3.30(paste0("./inst/extdata/",
model_dir,
datfile_name))
```

*Now changing catch units to tons!*

Write time series biomass (tons) and catch (tons--taken directly from truth without error) data for SS using `SS_write_ts`:

```{r SS-write-ts}

#Month of survey/fishing
survey_month <- 7
fishing_month <- 1

#Test writing CPUE in biomass
  stocksynthesis.data <- SS_write_ts(ss_data_list = stocksynthesis.data,
                ts_data = list(survObsBiom$atoutput[fish_years],
                  omlist_ss$truecatchbio_ss$atoutput[omlist_ss$truecatchbio_ss$time %in% fish_years]),
                CVs = c(CVs$survey,
                        CVs$fishery),
                data_years = list((survObsBiom$time[fish_years]-survey_sample_time)/timestep+1,             fish_years),
            sampling_month = list(rep(survey_month,nyears),
                                rep(fishing_month,nyears)),
                units = c("biomass","biomass"),
                data_type=c("CPUE","catch"),
                fleets = c(2,1))

stocksynthesis.data$CPUE

stocksynthesis.data$catch
```

Writing age composition data

```{r SS-write-comps}
#age_comp_data

#Get the age bins
age_bin_names <- names(stocksynthesis.data$agecomp)[10:length(names(stocksynthesis.data$agecomp))]
age_bins <- sub("a","",age_bin_names)

age_comp_flat <- reformat_compositions(age_comp_data,
                     round.places = 4,
                     comp_type = "agecomp")

## Write age composition data for survey
stocksynthesis.data <- SS_write_comps(ss_data_list = stocksynthesis.data,
               comp_matrix = list(age_comp_flat[burnin:(burnin+nyears-1),]),
               data_rows = list(stocksynthesis.data$styr:(stocksynthesis.data$styr+nyears-1)),
               sampling_month = list(rep(survey_month,nyears)),
               data_type = c("agecomp"),
               fleet_number = c(1),
               bins = list(age_bins),
               caal_bool = c(FALSE))
stocksynthesis.data$agecomp

```

Once we have the natlength matrix, we can munge the data into the proper CAAL and length bin format for SS

```{r SS-comps-formatting}

#Check length compositions match age compositions
#ss_length_stdsurv$natlength %>%
#  group_by(time,agecl) %>%
#  summarise(sum(atoutput))

len_comp_data <- ss_length_stdsurv$natlength
fish_len_comp_data <- catch_lengthwt_samp$natlength

if(fstepperyr>1){
  fish_len_comp_anndata <- fish_len_comp_data %>%
    mutate(yr = floor(time/fstepperyr)) %>%
    group_by(species, agecl, lower.bins, upper.bins, time=as.integer(yr)) %>%
    summarise(annnatlength=sum(atoutput)) %>%
    rename(atoutput = annnatlength)
} else {
  fish_len_comp_anndata <- fish_len_comp_data
}

caal_comp_flat <- reformat_compositions(len_comp_data,                                round.places=4,
                comp_type="caalcomp")

#remove burnin
caal_comp_final <- filter(caal_comp_flat,
                         time %in% survey_years)

#Add over age classes to get sample size
len_comp_flat <- reformat_compositions(len_comp_data,
                                  round.places = 0,
                           comp_type="lencomp")
#remove burnin
len_comp_final <- filter(len_comp_flat,
                         time %in% survey_years)

length_bins <- as.integer(names(len_comp_final))
length_bins <- length_bins[!is.na(length_bins)]

# fishery length comps are still 5 timesteps per year
# need to aggregate to annual (done above)
# also,  make effN annual goal/fstepsperyr (done above)
fish_len_comp_flat <- reformat_compositions(fish_len_comp_anndata,
                                  round.places = 0,
                           comp_type="lencomp")

#remove burnin works after adjustment above
fish_len_comp_final <- filter(fish_len_comp_flat,
                         time %in% fish_years)

notbins <- c("time", "nsamp")

# fish_length_bins <- as.integer(names(fish_len_comp_final))
# fish_length_bins <- fish_length_bins[!is.na(fish_length_bins)]

# need to fill empty length bins with 0s to have same bins as survey for SS_write_comps
missing.lengths <- setdiff(length_bins, names(fish_len_comp_final)[!names(fish_len_comp_final) %in% notbins])
fish_len_comp_final[as.character(missing.lengths)] <- 0                    # Add them, filled with '0's
fish_len_comp_final <- fish_len_comp_final[c("time", length_bins, "nsamp")]

# fishery age comps also 5 timesteps per year
if(fstepperyr>1){
  fish_age_comp_anndata <- catch_numsss_samp %>%
    mutate(yr = floor(time/fstepperyr)) %>%
    group_by(species, agecl, time=as.integer(yr)) %>%
    summarise(annnatage=sum(atoutput)) %>%
    rename(atoutput = annnatage)
} else {
  fish_age_comp_anndata <- catch_numsss_samp
}

fish_age_comp_flat <- reformat_compositions(fish_age_comp_anndata,
                           comp_type="agecomp")

#remove burnin (not necessary?fish comps made with fish_years only) 
fish_age_comp_final <- filter(fish_age_comp_flat,
                         time %in% fish_years)

# #SS_write_comps breaking because fishery age bins start with 2 not 1; extracting bins from fish file may help?
# fish_age_bins <- names(fish_age_comp_flat)[!names(fish_age_comp_flat) %in% notbins]

# that leaves an empty column in data file, so instead fill with 0s
missing.ages <- setdiff(age_bins, names(fish_age_comp_final)[!names(fish_age_comp_final) %in% notbins])
fish_age_comp_final[missing.ages] <- 0                    # Add them, filled with '0's
fish_age_comp_final <- fish_age_comp_final[c("time", age_bins, "nsamp")]

comp_list <- list(caal_comp_final,len_comp_final, fish_age_comp_final, fish_len_comp_final)

apply_month <- list(rep(survey_month, nrow(comp_list[[1]])), 
                    rep(survey_month, nrow(comp_list[[2]])),
                    rep(fishing_month,nrow(comp_list[[3]])),
                    rep(fishing_month,nrow(comp_list[[4]])))

# This now runs by ensuring that survey and fishery compositions have the same bins 
# (filled with 0s for missing bins in fishery relative to survey)

# Write CAAL and length composition data
stocksynthesis.data <- SS_write_comps(ss_data_list = stocksynthesis.data, 
                                      comp_matrix = comp_list, 
                                      data_rows = list((comp_list[[1]]$time-survey_sample_time)/timestep + 1 , (survey_years-survey_sample_time)/timestep + 1,fish_years,fish_years),
                                      sampling_month = apply_month, 
                                      data_type = rep(c("agecomp", "lencomp"),2), 
                                      fleet_number = c(2,2,1,1),  
                                      bins = list(age_bins, 
                                                  length_bins, 
                                                  age_bins, 
                                                  length_bins), 
                                      caal_bool = c(TRUE, rep(FALSE,3)))

head(stocksynthesis.data$lencomp)
head(stocksynthesis.data$agecomp)

#Change length bin structure to match atlantis data
stocksynthesis.data$lbin_vector <- length_bins

#Get correct number of length bins
stocksynthesis.data$N_lbins <- length(length_bins)

#Set lbin_method to 1 - this makes the population length bins match the data bins 
#When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
stocksynthesis.data$lbin_method <- 1
stocksynthesis.data$binwidth <- "#"
stocksynthesis.data$minimum_size <- "#"
stocksynthesis.data$maximum_size <- "#"

#Change minimum sample size to 0.001 for CAAL data (SS won't let it go lower than this)
stocksynthesis.data$age_info$minsamplesize <- rep(0.001,2)

SS_writedat_3.30(stocksynthesis.data, outfile = paste0("./inst/extdata/",model_dir,
datfile_name),
                 overwrite=TRUE)
```

## Coming up with dynamic life history parameters (partial from Christine)

Below we get the weight-length relationship parameters from atlantis, calculate h and R0 from Atlantis $\alpha$ and $\beta$ parameters, use the \code{calc_Z} function to back out natural mortality (M) from Atlantis, and use the survey length-at-age from Atlantis to estimate a growth curve.

The following is under construction from CreateStockSynthesis.Rmd and not evaluated. Existing control file was modified by hand to make the number of recruitment settlement assignments 1 instead of two, with GPattern month area age vector 1 1 1 1.

```{r, eval=FALSE}
#Get biological parameters

#Load needed inputs for biological parameters
source(here("config/CC3Config.R"))
#biological parameters
biolprm <- load_biolprm(dir = d.name, file_biolprm = biol.prm.file)#"C:/Users/chris/Documents/GitHub/atlantisom", 
#d.name <- "C:/Users/chris/Documents/GitHub/atlantisom"
#functional groups
fgs <- truth$fgs
runprm <- load_runprm(dir = d.name, file_runprm = run.prm.file)
YOY <- load_yoy(d.name, paste0("output", scenario.name, "YOY.txt"))
truenums_ss <- truth$nums[results$nums$species==species,]

YOY_ss <- YOY %>%
  select(Time, "SAR.0")

fullresZ <- calc_Z(yoy = YOY_ss,
                   nums = truenums_ss,
                   fgs = fgs,
                   biolprm = biolprm,
                   toutinc = runprm$toutinc)

meanwt_spp <- ss_length_stdsurv$muweight %>% 
  filter(time>burnin) %>%
  group_by(agecl) %>%
  summarize(meanwt = mean(atoutput))

ctlfile_name <-  "sardEM_3_3.ctl"
sardine.ctl <-r4ss::SS_readctl_3.30(paste0("./inst/extdata/",
model_dir,
ctlfile_name))
sardine.ctl <- SS_write_biol(sardine.ctl, biolprm, "SAR", Z = mean(fullresZ$atoutput), wtsage=meanwt_spp)

li_a_use <- biolprm$wl[match(fgs$Code[match(species,fgs$Name)],biolprm$wl[, 1]), 2]/1000
li_b_use <- biolprm$wl[match(fgs$Code[match(species,fgs$Name)],biolprm$wl[, 1]), 3]

#maturity ogive
#check what R0 is

len_nonburn <- ss_length_stdsurv$mulen %>%
  filter(time > burnin*timestep)
plot(len_nonburn$atoutput~len_nonburn$agecl, col=len_nonburn$time)

length.data <- data.frame("Year"=(len_nonburn$time-survey_sample_time)/timestep+1, length=len_nonburn$atoutput, Weight=NA, Sex="Female", age=as.integer(len_nonburn$agecl))

#require(bbmle)
#vb_estimates <-sample_fit_vbgf(length.data, 25, 45, 0.4,
#  0.1, 0.1, 20, 40, 0.05, 0.01, 0.01,
#  30, 50, 0.6, 0.3, 0.3, 0.5, 10)

```

## Run Stock Synthesis
After the data and control file modifications, ensure you copy the local SS 3.3 executable into the folder containing the model files. Then run the following chunk.

```{r runSS}
#Run stock synthesis model 
# default runs with admb_options = "-noest"
#run_stocksynthesis(species = species, model_dir = model_dir, show_output = TRUE)

# for estimation
run_stocksynthesis(species = species, model_dir = model_dir, admb_options = "", show_output = TRUE)

```