### 2022-10-28 NewCCASardines.Rmd

---
title: "New CCA sardines, Sept 2022"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)
library(atlantisom)
library(ggthemes)
library(FSA)
```

A test run for sardines with the updated California Current model, September 2022:

```{r full-test, message=FALSE, warning=FALSE, eval=FALSE}
CCom <- atlantisom::om_init(here("config/CCConfigSep22.R"))

CCom_sard <- atlantisom::om_species(c("Pacific_sardine"), CCom)

CCom_sard_ind <- atlantisom::om_index(
  usersurvey = here("config/sardinesurvey.R"),
  userfishery = here("config/sardinefishery.R"),
  omlist_ss = CCom_sard,
  n_reps = 1,
  save = TRUE
)

CCom_sard_comp <- atlantisom::om_comps(
  usersurvey = here("config/sardinesurvey.R"),
  userfishery = here("config/sardinefishery.R"),
  omlist_ss = CCom_sard,
  n_reps = 1,
  save = TRUE
)
```

Diagnose why survey functions resulting in NA outputs. Found it: change `sapply` to `lapply` in omdimensions to make the survey selectivity dataframe. The `sapply` function behaves inconsistently when the age classes are all 10, it orients the list incorrectly; `lapply` keeps the structure correct. This only worked for NOBA because I selected a species with fewer than 10 age classes.

```{r testomindex, eval=FALSE}

usersurvey <- here("config/sardinesurvey.R")

omlist_ss <- CCom_sard

# this is om_index

# one script for dimension parameters to be used in multiple functions
source("config/omdimensions.R", local = TRUE)

###############################################
n_reps <- 1 # not sure about it, used the value from line 31
save <- FALSE # Sarah already provided output data
###############################################

# user options for survey--default is a census with mid-year sample
# allows muliple surveys
survObsBiomBs <- list()

for (s in usersurvey)
{
  source(s, local = TRUE)

  # biomass based fishery independent survey index
  # this uses result$biomass_ages to sample biomass directly, no need for wt@age est
  survey_B <- atlantisom::create_survey(
    dat = omlist_ss$truebio_ss,
    time = survtime,
    species = survspp,
    boxes = survboxes,
    effic = surveffic,
    selex = survselex
  )

  # call sample_survey_biomass with a bunch of 1000s for weight at age
  # in the code it multiplies atoutput by wtatage/1000 so this allows us to use
  # biomass directly
  wtage <- data.frame(
    species = rep(names(age_classes), n_age_classes),
    agecl = unlist(sapply(n_age_classes, seq)),
    wtAtAge = rep(1000.0, sum(n_age_classes))
  )

  # this is the step to repeat n_reps time if we want different realizations
  # of the same survey design specified above; only observation error differs
  # using the census cv of 0 will produce identical reps!
  survObsBiomB <- list()
  for (i in 1:n_reps) {
    survObsBiomB[[i]] <- atlantisom::sample_survey_biomass(survey_B, surv_cv, wtage)
  }

  # save survey indices, takes a long time to generate with lots of reps/species
  if (save) {
    saveRDS(survObsBiomB, file.path(d.name, paste0(
      scenario.name, "_",
      survey.name, "surveyB.rds"
    )))
  }

  survObsBiomBs[[survey.name]] <- survObsBiomB
}

# configure the fishery, a default is in config/fisherycensus.R
# fishery configuration can specify only area and time of observation
# fishery species inherited from omlist_ss
# this is total catch not by fleet, so only one "fishery"

##################################################
# Use information from lines 29 and 35
userfishery <- here("config/sardinefishery.R")
##################################################
source(userfishery, local = TRUE)

# we are not currently subsetting fishery catch because we cannot correct catch.nc
#  instead the catch in biomass from catch.txt is read in for the index
#  we do not apply any cv to this, but we could this way (default cv=0)

fishObsCatchB <- list()
for (i in 1:n_reps) {
  fishObsCatchB[[i]] <- atlantisom::sample_fishery_totcatch(omlist_ss$truecatchbio_ss, fish_cv)
}

if (save) {
  saveRDS(fishObsCatchB, file.path(d.name, paste0(
    scenario.name, "_",
    fishery.name, "fishCatch.rds"
  )))
}

indices <- list(
  "survObsBiomB" = survObsBiomBs,
  "fishObsCatchB" = fishObsCatchB
)

#  return(indices)
# }
```

Diagnose why comps aren't working. Survey worked, fishery didn't because the burnin period was longer than this run. Changed to 0 in the sardinefishery.R config file and runs.

```{r testomcomps, eval=FALSE}

usersurvey <- here("config/sardinesurvey.R")
userfishery <- here("config/sardinefishery.R")

omlist_ss <- CCom_sard

n_reps <- 1

# this is om_comps

# one script for dimension parameters to be used in multiple functions
source("config/omdimensions.R", local = TRUE)

# user options for survey--default is a census with mid-year sample
# allows muliple surveys
age_comp_datas <- list()
survObsLenComps <- list()
survObsWtAtAges <- list()

for (s in usersurvey)
{
  source(s, local = TRUE)

  # numbers based fishery independent survey for age and length comps
  # same user specifications as indices
  survey_N <- atlantisom::create_survey(
    dat = omlist_ss$truenums_ss,
    time = survtime,
    species = survspp,
    boxes = survboxes,
    effic = surveffic,
    selex = survselex.agecl
  )

  # Sample fish for age composition
  # if we want replicates for obs error this sample function will generate them
  age_comp_data <- list()
  for (i in 1:n_reps) {
    age_comp_data[[i]] <- atlantisom::sample_fish(survey_N, surveffN)
  }

  # save age comps
  if (save) {
    saveRDS(age_comp_data, file.path(d.name, paste0(
      scenario.name, "_",
      survey.name, "survObsAgeComp.rds"
    )))
  }

  # weights needed for weight at age and length comp calcs
  # aggregate true resn per survey design
  survey_aggresn <- atlantisom::aggregateDensityData(
    dat = omlist_ss$trueresn_ss,
    time = survtime,
    species = survspp,
    boxes = survboxes
  )

  # aggregate true structn per survey design
  survey_aggstructn <- atlantisom::aggregateDensityData(
    dat = omlist_ss$truestructn_ss,
    time = survtime,
    species = survspp,
    boxes = survboxes
  )

  # dont sample these, just aggregate them using median
  structnss <- atlantisom::sample_fish(survey_aggstructn, surveffN, sample = FALSE)

  resnss <- atlantisom::sample_fish(survey_aggresn, surveffN, sample = FALSE)

  # this is all input into the length function, replicates follow age comp reps
  #  separating the length comps from the weight at age here
  survey_lenwt <- list()
  survObsLenComp <- list()
  survObsWtAtAge <- list()

  for (i in 1:n_reps) {
    survey_lenwt[[i]] <- atlantisom::calc_age2length(
      structn = structnss,
      resn = resnss,
      nums = age_comp_data[[i]],
      biolprm = omlist_ss$biol,
      fgs = omlist_ss$funct.group_ss,
      maxbin = maxbin,
      CVlenage = lenage_cv,
      remove.zeroes = TRUE
    )

    survObsLenComp[[i]] <- survey_lenwt[[i]]$natlength
    survObsWtAtAge[[i]] <- survey_lenwt[[i]]$muweight
  }

  if (save) {
    saveRDS(survObsLenComp, file.path(d.name, paste0(
      scenario.name, "_",
      survey.name, "survObsLenComp.rds"
    )))
    saveRDS(survObsWtAtAge, file.path(d.name, paste0(
      scenario.name, "_",
      survey.name, "survObsWtAtAge.rds"
    )))
  }
  # add each survey to master list objects for survey data
  age_comp_datas[[survey.name]] <- age_comp_data
  survObsLenComps[[survey.name]] <- survObsLenComp
  survObsWtAtAges[[survey.name]] <- survObsWtAtAge
}

# now do fishery comps
# user options for fishery--default is a census with mid-year sample
# only one fishery, but multiple fleets possible within it
source(userfishery, local = TRUE)

# fishery catch at age each observed timestep summed over observed polygons
# catch at age by area and timestep
catch_numbers <- atlantisom::create_fishery_subset(
  dat = omlist_ss$truecatchnum_ss,
  time = fishtime,
  species = survspp,
  boxes = fishboxes
)

# if we want replicates for obs error this sample function will generate them
catch_age_comp <- list()
for (i in 1:n_reps) {
  catch_age_comp[[i]] <- atlantisom::sample_fish(catch_numbers, fisheffN)
}

# save fishery age comps
if (save) {
  saveRDS(catch_age_comp, file.path(d.name, paste0(
    scenario.name, "_",
    fishery.name, "fishObsAgeComp.rds"
  )))
}

# Get catch weights for length comp calc
# aggregate true resn per fishery subset design
catch_aggresnss <- atlantisom::aggregateDensityData(
  dat = omlist_ss$trueresn_ss,
  time = fishtime,
  species = survspp,
  boxes = fishboxes
)

# aggregate true structn fishery subsetdesign
catch_aggstructnss <- atlantisom::aggregateDensityData(
  dat = omlist_ss$truestructn_ss,
  time = fishtime,
  species = survspp,
  boxes = fishboxes
)

# dont sample these, just aggregate them using median
catch_structnss <- atlantisom::sample_fish(catch_aggstructnss, fisheffN, sample = FALSE)

catch_resnss <- atlantisom::sample_fish(catch_aggresnss, fisheffN, sample = FALSE)

# these fishery lengths and weight at age are each output timestep
# same structure as above for surveys, replicates follow age comp reps
#  separating the length comps from the weight at age here
fishery_lenwt <- list()
fishObsLenComp <- list()
fishObsWtAtAge <- list()

for (i in 1:n_reps) {
  fishery_lenwt[[i]] <- atlantisom::calc_age2length(
    structn = catch_structnss,
    resn = catch_resnss,
    nums = catch_age_comp[[i]],
    biolprm = omlist_ss$biol,
    fgs = omlist_ss$funct.group_ss,
    maxbin = maxbin,
    CVlenage = lenage_cv,
    remove.zeroes = TRUE
  )

  fishObsLenComp[[i]] <- fishery_lenwt[[i]]$natlength
  fishObsWtAtAge[[i]] <- fishery_lenwt[[i]]$muweight
}

if (save) {
  saveRDS(fishObsLenComp, file.path(d.name, paste0(
    scenario.name, "_",
    fishery.name, "fishObsLenComp.rds"
  )))
  saveRDS(fishObsWtAtAge, file.path(d.name, paste0(
    scenario.name, "_",
    fishery.name, "fishObsWtAtAge.rds"
  )))
}

if (!is.null(omlist_ss$truenumsage_ss)) {
  # numbers based fishery independent survey for age and length comps
  # allows for mulitple surveys
  annage_comp_datas <- list()

  for (s in usersurvey)
  {
    source(s, local = TRUE)

    # same user specifications as indices
    survey_annageN <- atlantisom::create_survey(
      dat = omlist_ss$truenumsage_ss,
      time = survtime,
      species = survspp,
      boxes = survboxes,
      effic = surveffic,
      selex = survselex
    )
    # Sample fish for age composition
    # if we want replicates for obs error this sample function will generate them
    annage_comp_data <- list()
    for (i in 1:n_reps) {
      annage_comp_data[[i]] <- atlantisom::sample_fish(survey_annageN, surveffN)
    }

    # save survey annual age comps
    if (save) {
      saveRDS(annage_comp_data, file.path(d.name, paste0(
        scenario.name, "_",
        survey.name, "survObsFullAgeComp.rds"
      )))
    }
    annage_comp_datas[[survey.name]] <- annage_comp_data
  }
} else {
  annage_comp_datas <- NULL
}

if (!is.null(omlist_ss$truecatchage_ss)) {
  # fishery catch at age each observed timestep summed over observed polygons
  # catch at age by area and timestep
  catch_annagenumbers <- atlantisom::create_fishery_subset(
    dat = omlist_ss$truecatchage_ss,
    time = fishtime,
    species = survspp,
    boxes = fishboxes
  )

  # if we want replicates for obs error this sample function will generate them
  # WARNING THIS AGGREGATES ACROSS FLEETS
  # TODO: need to change sample_fish to fix
  catch_annage_comp <- list()
  for (i in 1:n_reps) {
    catch_annage_comp[[i]] <- atlantisom::sample_fish(catch_annagenumbers, fisheffN)
  }

  # save fishery annual age comps
  if (save) {
    saveRDS(catch_annage_comp, file.path(d.name, paste0(
      scenario.name, "_",
      fishery.name, "fishObsFullAgeComp.rds"
    )))
  }
} else {
  catch_annage_comp <- NULL
}

# call interpolate weight at age function to get survObsFullWtAtAge
if (!is.null(omlist_ss$truenumsage_ss)) {
  interp_survWtAtAges <- list()
  for (s in usersurvey)
  {
    source(s, local = TRUE)

    interp_survWtAtAge <- list()
    for (i in 1:n_reps) {
      interp_survWtAtAge[[i]] <- calc_avgwtstage2age(
        wtagecl = survObsWtAtAges[[survey.name]][[i]],
        annages = omlist_ss$truenumsage_ss,
        fgs = omlist_ss$funct.group_ss
      )
    }
    if (save) {
      saveRDS(interp_survWtAtAge, file.path(d.name, paste0(
        scenario.name, "_",
        survey.name, "survObsFullWtAtAge.rds"
      )))
    }
    interp_survWtAtAges[[survey.name]] <- interp_survWtAtAge
  }
} else {
  interp_survWtAtAges <- NULL
}

# do we want fishery average weight at true age too? why not
# call interpolate weight at age function to get fishObsFullWtAtAge
# WARNING currently aggregates out fleet info, but no fleets in aggregate wtage
if (!is.null(omlist_ss$truecatchage_ss)) {
  interp_fishWtAtAge <- list()
  for (i in 1:n_reps) {
    interp_fishWtAtAge[[i]] <- calc_avgwtstage2age(
      wtagecl = fishObsWtAtAge[[i]],
      annages = omlist_ss$truecatchage_ss,
      fgs = omlist_ss$funct.group_ss
    )
  }
  if (save) {
    saveRDS(interp_fishWtAtAge, file.path(d.name, paste0(
      scenario.name, "_",
      fishery.name, "fishObsFullWtAtAge.rds"
    )))
  }
} else {
  interp_fishWtAtAge <- NULL
}

comps <- list(
  "survObsAgeComp" = age_comp_datas,
  "survObsLenComp" = survObsLenComps,
  "survObsWtAtAge" = survObsWtAtAges,
  "fishObsAgeComp" = catch_age_comp,
  "fishObsLenComp" = fishObsLenComp,
  "fishObsWtAtAge" = fishObsWtAtAge,
  "survObsFullAgeComp" = annage_comp_datas,
  "fishObsFullAgeComp" = catch_annage_comp,
  "survObsFullWtAtAge" = interp_survWtAtAges,
  "fishObsFullWtAtAge" = interp_fishWtAtAge
)

#  return(comps)
# }
```

### Wrapper test results {.tabset}

#### Biomass index
```{r bioind-1}

source(here("config/CCConfigSep22.R"))

omlist_ss <- readRDS(file.path(d.name, paste0(scenario.name, "omlist_ss.rds")))

source(here("config/omdimensions.R"))
source(here("config/sardinesurvey.R"))
source(here("config/sardinefishery.R"))

# read time series data
survObsBiom <- readRDS(file.path(d.name, paste0(
  scenario.name,
  "_",
  survey.name,
  "surveyB.rds"
)))

survObsBiom <- survObsBiom[[1]]

plotB <- ggplot() +
  geom_line(
    data = survObsBiom, aes(x = time / stepperyr, y = atoutput, color = "survey Biomass"),
    alpha = 10 / 10
  ) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour = scenario.name)

plotB +
  facet_wrap(~species, scales = "free")
```

#### Catch time series
```{r catchind-1}
# read time series data
catchbio_ss <- readRDS(file.path(d.name, paste0(
  scenario.name,
  "_",
  fishery.name, "fishCatch.rds"
)))

catchbio_ss <- catchbio_ss[[1]]

plotC <- ggplot() +
  geom_line(
    data = catchbio_ss, aes(x = time / 365, y = atoutput, color = "observed Catch"),
    alpha = 10 / 10
  ) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour = scenario.name)

plotC +
  facet_wrap(~species, scales = "free")
```

#### Survey length composition
```{r lencomp-1}

# length comps
len_comp_data <- readRDS(file.path(d.name, paste0(
  scenario.name,
  "_",
  survey.name, "survObsLenComp.rds"
)))
fish_len_comp_data <- readRDS(file.path(d.name, paste0(
  scenario.name,
  "_",
  fishery.name, "fishObsLenComp.rds"
)))

len_comp_data <- len_comp_data[[1]]
fish_len_comp_data <- fish_len_comp_data[[1]]

# add this to om_indices function so that this has years when read in
fish_len_comp_data$time <- as.integer(floor(fish_len_comp_data$time / fstepperyr))

len <- len_comp_data

len <- filter(len_comp_data, time %in% c(55:175))

lfplot <- ggplot(len, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte() +
  labs(subtitle = paste(
    scenario.name,
    len$species
  ))

lfplot + facet_wrap(~ time / stepperyr, ncol = 6, scales = "free_y")
```

#### Survey age composition (age classes)
```{r agecomp-1}

# read in comp data
age_comp_data <- readRDS(file.path(d.name, paste0(
  scenario.name,
  "_",
  survey.name, "survObsAgeComp.rds"
)))
age_comp_data <- age_comp_data[[1]]

Natage <- age_comp_data

Natage <- filter(age_comp_data, time %in% c(150:270))

Natageplot <- ggplot(Natage, aes(x = agecl, y = atoutput)) +
  geom_point() +
  theme_tufte() +
  labs(subtitle = paste(
    scenario.name,
    Natage$species
  ))

Natageplot + facet_wrap(~ time / stepperyr, ncol = 6, scales = "free_y")
```

#### Survey weight at age (age classes)
```{r wtageclass-1}

wtage <- readRDS(file.path(d.name, paste0(
  scenario.name,
  "_",
  survey.name, "survObsWtAtAge.rds"
)))
wtage <- wtage[[1]]

wageplot <- ggplot(wtage, aes(time, atoutput)) +
  geom_line(aes(colour = factor(agecl))) +
  theme_tufte() +
  theme(legend.position = "bottom") +
  xlab("model timestep (5 per year)") +
  ylab("average individual weight (g)") +
  labs(subtitle = paste0(scenario.name, " annual mid year sample"))

wageplot + facet_wrap(c("species"), scales = "free_y")
```

#### Fishery age composition (age classes)
```{r fishagecomp-1}

# read in comp data
fish_age_comp <- readRDS(file.path(d.name, paste0(
  scenario.name,
  "_",
  fishery.name, "fishObsAgeComp.rds"
)))
fish_age_comp <- fish_age_comp[[1]]

# add this to om_indices function so that this has years when read in
fish_age_comp$time <- fish_age_comp$time / fstepperyr

Natage <- fish_age_comp

Natage <- filter(fish_age_comp, time %in% c(30:53))

Natageplot <- ggplot(Natage, aes(x = agecl, y = atoutput)) +
  geom_point() +
  theme_tufte() +
  labs(subtitle = paste(
    scenario.name,
    Natage$species
  ))

Natageplot + facet_wrap(~time, ncol = 6, scales = "free_y")
```

### {-}

## Write data for Stock Synthesis

To read in the Stock Synthesis files, first ensure you have the latest version of `r4ss`. You will need to ensure the relevant Stock Synthesis files are stored in the inst/extdata/ folder, with each species example in its own folder.

```{r readdata}
# devtools::install_github("r4ss/r4ss", dependencies = FALSE)
require(r4ss)
# source("./config/sardine_config.R") this is above for illustration

# Set species name
species <- "Pacific_sardine" # change to species_ss if we want both

# Directory with SS files
model_dir <- "Sardine_SS_files_Sep2022/"

# Name of SS data file
datfile_name <- "sardEM_3_3.dat"

stocksynthesis.data <- r4ss::SS_readdat_3.30(paste0(
  "./inst/extdata/",
  model_dir,
  datfile_name
))
```

```{r sardine-survey-spec}
# Years to survey, assuming survey is once per year
# survey_years <- survey_sample_full[(burnin+1):(burnin+nyears)] # my original
survey_years <- survey_sample_full[burnin:(burnin + nyears - 1)] # from Christine's new sardine_config.R

# Month of survey/fishing
survey_month <- 7
fishing_month <- 1
```

*Now changing catch units to tons!*

Write time series biomass (tons) and catch (tons--taken directly from truth without error) data for SS using `SS_write_ts`:

```{r SS-write-ts}

# Test writing CPUE in biomass

stocksynthesis.data <- SS_write_ts(
  ss_data_list = stocksynthesis.data,
  ts_data = list(
    survObsBiom$atoutput[fish_years],
    omlist_ss$truecatchbio_ss$atoutput[(omlist_ss$truecatchbio_ss$time / 365) %in% fish_years]
  ),
  CVs = c(
    surv_cv$cv,
    fish_cv$cv
  ),
  data_years = list((survObsBiom$time[fish_years] - survey_sample_time) / timestep + 1, fish_years),
  sampling_month = list(
    rep(survey_month, nyears),
    rep(fishing_month, nyears)
  ),
  units = c("biomass", "biomass"),
  data_type = c("CPUE", "catch"),
  fleets = c(2, 1)
)

stocksynthesis.data$CPUE

stocksynthesis.data$catch
```

Writing age composition data

```{r SS-write-comps}
# age_comp_data

# Get the age bins
age_bin_names <- names(stocksynthesis.data$agecomp)[10:length(names(stocksynthesis.data$agecomp))]
age_bins <- sub("a", "", age_bin_names)

age_comp_flat <- reformat_compositions(age_comp_data,
  round.places = 4,
  comp_type = "agecomp"
)

## Write age composition data for survey
stocksynthesis.data <- SS_write_comps(
  ss_data_list = stocksynthesis.data,
  comp_matrix = list(age_comp_flat[burnin:(burnin + nyears - 1), ]),
  data_rows = list(stocksynthesis.data$styr:(stocksynthesis.data$styr + nyears - 1)),
  sampling_month = list(rep(survey_month, nyears)),
  data_type = c("agecomp"),
  fleet_number = c(2),
  bins = list(age_bins),
  caal_bool = c(FALSE)
)

stocksynthesis.data$agecomp[, age_bin_names] <- as.data.frame(age_comp_flat)[burnin:(burnin + nyears - 1), age_bins]
```

Once we have the natlength matrix, we can munge the data into the proper CAAL and length bin format for SS

```{r SS-comps-formatting}

# Check length compositions match age compositions
# ss_length_stdsurv$natlength %>%
#  group_by(time,agecl) %>%
#  summarise(sum(atoutput))

len_comp_data <- survey_lenwt[[1]]$natlength
fish_len_comp_data <- fishery_lenwt[[1]]$natlength

if (fstepperyr > 1) {
  fish_len_comp_anndata <- fish_len_comp_data %>%
    mutate(yr = floor(time / fstepperyr)) %>%
    group_by(species, agecl, lower.bins, upper.bins, time = as.integer(yr)) %>%
    summarise(annnatlength = sum(atoutput)) %>%
    rename(atoutput = annnatlength)
} else {
  fish_len_comp_anndata <- fish_len_comp_data
}

caal_comp_flat <- reformat_compositions(len_comp_data,
  round.places = 4,
  comp_type = "caalcomp"
)

# remove burnin
caal_comp_final <- filter(
  caal_comp_flat,
  time %in% survey_years
)
caal_comp_final <- as.data.frame(caal_comp_final)

# Add over age classes to get sample size
len_comp_flat <- reformat_compositions(len_comp_data,
  round.places = 0,
  comp_type = "lencomp"
)
# remove burnin
len_comp_final <- filter(
  len_comp_flat,
  time %in% survey_years
)
len_comp_final <- as.data.frame(len_comp_final)

length_bins <- as.integer(names(len_comp_final))
length_bins <- length_bins[!is.na(length_bins)]

# fishery length comps are still 5 timesteps per year
# need to aggregate to annual (done above)
# also,  make effN annual goal/fstepsperyr (done above)
fish_len_comp_flat <- reformat_compositions(fish_len_comp_anndata,
  round.places = 0,
  comp_type = "lencomp"
)

# remove burnin works after adjustment above
fish_len_comp_final <- filter(
  fish_len_comp_flat,
  time %in% fish_years
)
fish_len_comp_final <- as.data.frame(fish_len_comp_final)

notbins <- c("time", "nsamp")

# fish_length_bins <- as.integer(names(fish_len_comp_final))
# fish_length_bins <- fish_length_bins[!is.na(fish_length_bins)]

# need to fill empty length bins with 0s to have same bins as survey for SS_write_comps
missing.lengths <- setdiff(length_bins, names(fish_len_comp_final)[!names(fish_len_comp_final) %in% notbins])
fish_len_comp_final[as.character(missing.lengths)] <- 0 # Add them, filled with '0's
fish_len_comp_final <- fish_len_comp_final[c("time", length_bins, "nsamp")]

# fishery age comps also 5 timesteps per year
if (fstepperyr > 1) {
  fish_age_comp_anndata <- catch_age_comp[[1]] %>%
    mutate(yr = floor(time / fstepperyr)) %>%
    group_by(species, agecl, time = as.integer(yr)) %>%
    summarise(annnatage = sum(atoutput)) %>%
    rename(atoutput = annnatage)
} else {
  fish_age_comp_anndata <- catch_age_comp[[1]]
}

fish_age_comp_flat <- reformat_compositions(fish_age_comp_anndata,
  comp_type = "agecomp"
)

# remove burnin (not necessary?fish comps made with fish_years only)
fish_age_comp_final <- filter(
  fish_age_comp_flat,
  time %in% fish_years
)
fish_age_comp_final <- as.data.frame(fish_age_comp_final)

# #SS_write_comps breaking because fishery age bins start with 2 not 1; extracting bins from fish file may help?
# fish_age_bins <- names(fish_age_comp_flat)[!names(fish_age_comp_flat) %in% notbins]

# that leaves an empty column in data file, so instead fill with 0s
missing.ages <- setdiff(age_bins, names(fish_age_comp_final)[!names(fish_age_comp_final) %in% notbins])
fish_age_comp_final[missing.ages] <- 0 # Add them, filled with '0's
fish_age_comp_final <- fish_age_comp_final[c("time", age_bins, "nsamp")]

comp_list <- list(caal_comp_final, len_comp_final, fish_age_comp_final, fish_len_comp_final)

apply_month <- list(
  rep(survey_month, nrow(comp_list[[1]])),
  rep(survey_month, nrow(comp_list[[2]])),
  rep(fishing_month, nrow(comp_list[[3]])),
  rep(fishing_month, nrow(comp_list[[4]]))
)

# This now runs by ensuring that survey and fishery compositions have the same bins
# (filled with 0s for missing bins in fishery relative to survey)

# Write CAAL and length composition data
stocksynthesis.data <- SS_write_comps(
  ss_data_list = stocksynthesis.data,
  comp_matrix = comp_list,
  data_rows = list((comp_list[[1]]$time - survey_sample_time) / timestep + 1, (survey_years - survey_sample_time) / timestep + 1, fish_years, fish_years),
  sampling_month = apply_month,
  data_type = rep(c("agecomp", "lencomp"), 2),
  fleet_number = c(2, 2, 1, 1),
  bins = list(
    age_bins,
    length_bins,
    age_bins,
    length_bins
  ),
  caal_bool = c(TRUE, rep(FALSE, 3))
)

head(stocksynthesis.data$lencomp)
head(stocksynthesis.data$agecomp)

# Change length bin structure to match atlantis data
stocksynthesis.data$lbin_vector <- length_bins

# Get correct number of length bins
stocksynthesis.data$N_lbins <- length(length_bins)

# Set lbin_method to 1 - this makes the population length bins match the data bins
# When lbin_method==1, we just comment out the binwidth, minimum, and maximum size arguments since they aren't used
stocksynthesis.data$lbin_method <- 1
stocksynthesis.data$binwidth <- "#"
stocksynthesis.data$minimum_size <- "#"
stocksynthesis.data$maximum_size <- "#"

# Change minimum sample size to 0.001 for CAAL data (SS won't let it go lower than this)
stocksynthesis.data$age_info$minsamplesize <- rep(0.001, 2)

SS_writedat_3.30(
  datlist = stocksynthesis.data,
  outfile = paste0(
    "./inst/extdata/", model_dir,
    datfile_name
  ),
  overwrite = TRUE
)
```

## Coming up with dynamic life history parameters (partial from Christine)

Below we get the weight-length relationship parameters from atlantis, calculate h and R0 from Atlantis $\alpha$ and $\beta$ parameters, use the \code{calc_Z} function to back out natural mortality (M) from Atlantis, and use the survey length-at-age from Atlantis to estimate a growth curve.

The following is under construction from CreateStockSynthesis.Rmd and not evaluated. Existing control file was modified by hand to make the number of recruitment settlement assignments 1 instead of two, with GPattern month area age vector 1 1 1 1.

```{r, eval=FALSE}

# # Need to clone atlantisom and load the package locally because the
# # installed package from remote repository does not include SS_write_biol()
# devtools::load_all("C:/Users/bai.li/Documents/sardine/atlantisom")
# 
# # Get biological parameters
# 
# # Load needed inputs for biological parameters
# 
# # biological parameters
# biolprm <- load_biolprm(dir = d.name, file_biolprm = biol.prm.file) # "C:/Users/chris/Documents/GitHub/atlantisom"
# # functional groups
# 
# fgs <- CCom$truth$fgs
# runprm <- load_runprm(dir = d.name, file_runprm = run.prm.file)
# YOY <- load_yoy(d.name, paste0(scenario.name, "YOY.txt"))
# truenums_ss <- CCom$truth$nums[CCom$truth$nums$species == species, ]
# 
# YOY_ss <- YOY %>%
#   select(Time, "SAR.0")
# 
# fullresZ <- calc_Z(
#   yoy = YOY_ss,
#   nums = truenums_ss,
#   fgs = fgs,
#   biolprm = biolprm,
#   toutinc = runprm$toutinc
# )
# Z <- mean(fullresZ$atoutput)
# 
# meanwt_spp <- survey_lenwt[[1]]$muweight %>%
#   filter(time > burnin) %>%
#   group_by(agecl) %>%
#   summarize(meanwt = mean(atoutput))
# 
# ctlfile_name <- "sardEM_3_3.ctl"
# sardine.ctl <- r4ss::SS_readctl_3.30(paste0(
#   "./inst/extdata/",
#   model_dir,
#   ctlfile_name
# ),
# datlist = paste0(
#   "./inst/extdata/", model_dir,
#   datfile_name
# )
# )
# 
# lensage <- data.frame(
#   agecl = age_classes[[species]],
#   meanln = seq(15.188, 48.39, length = length(age_classes[[species]])),
#   cvln = seq(15.188 * 0.25, 0.9, length = length(age_classes[[species]]))
# )
# cvo_guess <- 48.39
# 
# sardine.ctl <- atlantisom::SS_write_biol(sardine.ctl,
#   biolprm,
#   "SAR",
#   wtsage = meanwt_spp,
#   M_est = rep(0.4, length(age_classes[[species]])),
#   lensage = lensage
# )
# 
# li_a_use <- biolprm$wl[match(fgs$Code[match(species, fgs$Name)], biolprm$wl[, 1]), 2] / 1000
# li_b_use <- biolprm$wl[match(fgs$Code[match(species, fgs$Name)], biolprm$wl[, 1]), 3]
# 
# # maturity ogive
# # check what R0 is
# 
# 
# 
# len_nonburn <- survey_lenwt[[1]]$mulen %>%
#   filter(time > burnin * timestep)
# plot(len_nonburn$atoutput ~ len_nonburn$agecl, col = len_nonburn$time)
# 
# length.data <- data.frame("Year" = (len_nonburn$time - survey_sample_time) / timestep + 1, length = len_nonburn$atoutput, Weight = NA, Sex = "Female", age = as.integer(len_nonburn$agecl))
# 
# # require(bbmle)
# # vb_estimates <-sample_fit_vbgf(length.data, 25, 45, 0.4,
# #  0.1, 0.1, 20, 40, 0.05, 0.01, 0.01,
# #  30, 50, 0.6, 0.3, 0.3, 0.5, 10)
```

## Run Stock Synthesis
After the data and control file modifications, ensure you copy the local SS 3.3 executable into the folder containing the model files. Then run the following chunk.

```{r runSS}
# Run stock synthesis model
# default runs with admb_options = "-noest"
# run_stocksynthesis(species = species, model_dir = model_dir, show_output = TRUE)

# for estimation
run_stocksynthesis(species = species, model_dir = model_dir, admb_options = "", show_output = TRUE)
```