### 2023-05-17 NewCatchWorkflow.Rmd
test with NOBA sacc30-climate run

---
title: "New catch functions from CATCH.nc file"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_fold: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(here)
library(stringr)
library(magrittr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(atlantisom)
```

## Intro

Total catch in tons has been pulled from the Atlantis `CATCH.txt` output file from 2019-2023 due to difficulties scaling catch in numbers from the `CATCH.nc` file to total weight. See comparisons [here](https://sgaichas.github.io/poseidon-dev/TestCatchNOBA.html).

For testing multispecies models, there is a need for subannual catch information, which is not provided in the annual `CATCH.txt` output. In addition, there is no polygon information in the `CATCH.txt` output, so spatial subsetting is not possible.

In order to get catch in tons subannually and by area, we can use the fleet-specific catch information in the `CATCH.nc` file. These outputs are reported in tons by species in aggregate over all age classes for each defined fleet.

## Methods

Workflow

1.  Modify `atlantisom::load_nc` to get catch in tons from different `CATCH.nc` outputs. New function because current function already gets catch in numbers, would be complicated to use the same "Catch" variable to get tons. Name new function `atlantisom::load_nc_catchtons`

2.  Test new `atlantisom::load_nc_catchtons` with NOBA, NEUS, CC by summing output over polygons and fleets to each year and comparing with annual output of current `atlantisom::load_catch` based on `CATCH.txt`

3.  Incorporate `atlantisom::load_nc_catchtons` into `atlantisom::run_truth`. Add a new catchtons object keeping polygon, timestep, and fleet information to the output list. Consider removing the `CATCH.txt` output in this function, currently called `catch_all` but not used? Or keep it, since catchtons can be summed to catch_all this could be used as an internal test as was started in the run_truth function.

4.  Adjust `atlantisom::om_init` for the new catchtons output. Should we keep reading in the `CATCH.txt` here? It may be ok to keep it just to keep existing functions for aggregate annual catch working.

5.  Adjust `atlantisom::om_species` for the new catchtons output.

6.  Adjust `atlantisom::om_index` for the new catchtons output. Create a new index that will be consistent across the subannual and annual datasets: this means applying the cv at the level of reporting? So possibly by fleet and timestep, then summing to annual? This involves including and updating two functions: `atlantisom::create_fishery_subset` parallel to the `atlantisom::create_survey` function to subset areas, currently used only on fishery comps but can now be applied to the index as well, and updating another `atlantisom::sample_fishery_totalcatch` for the cv application. This will save perhaps one new output and update one existing output, or just one output? Will we have the same current `censusfishCatch.rds` output that is annually aggregated plus a disaggregated version? Or have a single output with the same name `censusfishCatch.rds` that is disaggregated, to be summed separately in mskeyrun?

7. Adjust `mskeyrun` functions creating aggregated and subannual catch datasets. Do we need a fleet specific index?

### 1. New `atlantisom::load_nc_catchtons` function

Here is the new function, tested with snippets in mskeyrun sarah_wgsamsim branch and implemented in atlantisom dev branch

```{r, code = readLines("https://raw.githubusercontent.com/r4atlantis/atlantisom/dev/R/load_nc_catchtons.R"), eval=F}
```

### 2. Test `load_nc_catchtons` with the NOBA, NEUS, CC model outputs

Only NOBA first, this one sacc30 with climate, different from sacc38 in mskeyrun without

```{r}

# define for each model

## NOBA (different from mskeyrun version)

atlmod <- here("config/NOBAsacc30Config.R")

select_groups <- c("Long_rough_dab",
                   "Green_halibut",
                   "Mackerel",
                   "Haddock",
                   "Saithe",
                   "Redfish",
                   "Blue_whiting",
                   "Norwegian_ssh",
                   "North_atl_cod",
                   "Polar_cod",
                   "Capelin")

stepperyr <- 5

## CC

## NEUS

##### aggregate over polygons to fleets #######
# make this a function

source(atlmod)

dir <- d.name
nc_catch <- paste0(scenario.name, 'CATCH.nc')
file_fgs <- functional.groups.file
file_init <- initial.conditions.file
file_fish <- fisheries.file

# Get the boundary boxes
allboxes <- atlantisom::load_box(dir = d.name, file_bgm = box.file)
boxes <- atlantisom::get_boundary(allboxes)

  # Read in information
  # Read in the functional groups csv since that is used by many functions
  fgs <- load_fgs(dir = dir, file_fgs = file_fgs)
  # Read in the biomass pools
  bps <- load_bps(dir = dir, fgs = file_fgs, file_init = file_init)

  catchtons <- load_nc_catchtons(dir = dir,
                                 file_nc = nc_catch,
                                 file_fish = file_fish,
                                 bps = bps,
                                 fgs = fgs,
                                 select_groups = select_groups,
                                 select_variable = "Catch",
                                 check_acronyms = TRUE,
                                 bboxes = boxes)
  #if(verbose) message("Catch tons read in.")

# result is output of load_nc_catchtons

aggcatchtons <- catchtons %>%
  dplyr::select(-agecl) %>%
  dplyr::group_by(species, fleet, time) %>%
  dplyr::summarise(totcatch = sum(atoutput)) %>%
  dplyr::mutate(year = ceiling(time/stepperyr))

# compare annual catch in tons to CATCH.txt output
yearcatchtons <- aggcatchtons %>%
  dplyr::group_by(species, year) %>%
  dplyr::summarise(yeartot = sum(totcatch))

txtCatchtons <- atlantisom::load_catch(d.name, catch.file, fgs) %>%
  dplyr::filter(species %in% select_groups) %>%
  dplyr::mutate(year = time/365) %>%
  dplyr::left_join(dplyr::select(fgs, Code, Name), by = c("species" = "Name"))

comparetons <- txtCatchtons %>%
  dplyr::left_join(yearcatchtons, by = c("Code" = "species", "year" = "year")) %>%
  ggplot2::ggplot() +
  geom_line(aes(year, atoutput)) +
  geom_point(aes(year, yeartot), color = "blue", alpha = 0.2) +
  facet_wrap(~species, scales="free_y")

# looks like a match!               
comparetons
```