### 2023-06-14 NewBiomassPoolsWorkflow.Rmd
start biomass pools for rpath sim data

---
title: "New Biomass Pools Workflow"
author: "Sarah Gaichas"
date: "`r Sys.Date()`"
output:
  html_document:
    code_fold: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(here)
library(stringr)
library(magrittr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(atlantisom)
```

## Intro

Current `atlantisom` functions are designed primarily for age structured modeled groups, because the package was originally scoped to generate simulated data to test fishery stock assessments. It is now clear that simulated data testing can also benefit other model types, including multispecies models and food web models. Therefore, there is a need to develop functions to develop simulated data for lower trophic level biomass pools modeled in Atlantis and also included in food web models such as Rpath.

Where is biomass information for non-age structured groups output? To be parallel to survey sampling for fish, it would be good to have it by model output timestep toutinc and polygon, but depth layer is not necessary at this time.

First check content of the `.nc` output file:

```{r}

dir <- here("atlantisoutput/NOBA_sacc_30")
file_nc <- "nordic_runresults_01.nc"

file.nc <- file.path(dir, file_nc)

 # Load ATLANTIS output!
  at_out <- RNetCDF::open.nc(con = file.nc)
  on.exit(RNetCDF::close.nc(at_out))

    # Get info from netcdf file! (Filestructure and all variable names)
  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length

```

Which species are biomass pools, are they identified in the groups.csv?

No, biomass pools are defined as the combination of NumCohorts == 1 and GroupType not FISH, FISH_INVERT, SHARK, BIRD, MAMMAL, REPTILE. Biomass pools groups are among those below according to the Atlantis manual, ch 6. p 106:

>BIOMASS POOL GROUPS
Phytoplankton
7) Small Phytoplankton SM_PHY
8) Large Phytoplankton LG_PHY (used in combination with IsSiliconDep
to define diatoms)
9) Trichodesmium and Cyanobacteria TRICHO (N fixers)
Other primary producers
10) Microphtybenthos MICROPHTYBENTHOS
11) Dinoflagellates DINOFLAG
12) Phytoben PHYTOBEN (typically used to represent macroalgae) 
13) Seagrass SEAGRASS
14) Turf TURF
Zooplankton
15) Small Zooplankton SM_ZOO
16) Medium Zooplankton MED_ZOO
17) Large Zooplankton LG_ZOO
18) Jellyfish JELLIES (in the past represented as LG_ZOO)
Large pelagic invetebrates
19)Cephalopod CEP 
20) Prawns PWN
Infauna
21) Small Infauna SM_INF 
22) Large Infauna LG_INF
Epibenthic organisms
23) Sediment epibenthic filter feeders SED_EP_FF (often used for bivalves, sponges)

For NOBA Atlantis, these are the biomass pools:

```{r}

fgs <- atlantisom::load_fgs(here("atlantisoutput/NOBA_sacc_30"), "nordic_groups_v04.csv")

pools <- fgs %>%
  dplyr::filter(NumCohorts==1) %>%
  dplyr::select(Code, Name, Long.Name, NumCohorts, InvertType)

knitr::kable(pools)

```

Here are the relevant outputs in the output.nc file for these groups:

```{r}
# return everything in var_names_ncdf matching Name in pools

var_names_ncdf[str_detect(var_names_ncdf, str_c(pools$Name, collapse ="|"))]

```
So we want "N" nitrogen output. In theory this can be achieved with `atlantisom::load_nc`with `select_variable = "N"`.

```{r}

select_groups <- pools$Name
nc_out <- file_nc
bps <- load_bps(dir = dir, fgs = "nordic_groups_v04.csv", file_init = "nordic_biol_v23.nc")
  # Get the boundary boxes
  allboxes <- load_box(dir = dir, file_bgm = "Nordic02.bgm")
  boxes <- get_boundary(allboxes)

testNpools <- load_nc(dir = dir,
                  file_nc = nc_out,
                  bps = bps,
                  fgs = fgs,
                  select_groups = select_groups,
                  select_variable = "N",
                  check_acronyms = TRUE,
                  bboxes = boxes)
```

I don't think we have a function that converts N to biomass but there are pieces in other functions.