### 2023-08-23 NewBiomassPoolsWorkflow.Rmd
correction for benthos

---
title: "New Biomass Pools Workflow"
author: "Sarah Gaichas"
date: "`r Sys.Date()`"
output:
  html_document:
    code_fold: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(here)
library(stringr)
library(magrittr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(atlantisom)
```

## Intro

Current `atlantisom` functions are designed primarily for age structured modeled groups, because the package was originally scoped to generate simulated data to test fishery stock assessments. It is now clear that simulated data testing can also benefit other model types, including multispecies models and food web models. Therefore, there is a need to develop functions to develop simulated data for lower trophic level biomass pools modeled in Atlantis and also included in food web models such as Rpath.

Where is biomass information for non-age structured groups output? To be parallel to survey sampling for fish, it would be good to have it by model output timestep toutinc and polygon, but depth layer is not necessary at this time.

First check content of the `.nc` output file:

```{r}

dir <- here("atlantisoutput/NOBA_sacc_30")
file_nc <- "nordic_runresults_01.nc"

file.nc <- file.path(dir, file_nc)

 # Load ATLANTIS output!
  at_out <- RNetCDF::open.nc(con = file.nc)
  #on.exit(RNetCDF::close.nc(at_out))

    # Get info from netcdf file! (Filestructure and all variable names)
  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length

```

Which species are biomass pools, are they identified in the groups.csv?

No, biomass pools are defined as the combination of NumCohorts == 1 and GroupType not FISH, FISH_INVERT, SHARK, BIRD, MAMMAL, REPTILE. Biomass pools groups are among those below according to the Atlantis manual, ch 6. p 106:

>BIOMASS POOL GROUPS  
Phytoplankton  
7) Small Phytoplankton SM_PHY  
8) Large Phytoplankton LG_PHY (used in combination with IsSiliconDep to define diatoms)  
9) Trichodesmium and Cyanobacteria TRICHO (N fixers)  
Other primary producers  
10) Microphtybenthos MICROPHTYBENTHOS  
11) Dinoflagellates DINOFLAG  
12) Phytoben PHYTOBEN (typically used to represent macroalgae)  
13) Seagrass SEAGRASS  
14) Turf TURF  
Zooplankton  
15) Small Zooplankton SM_ZOO  
16) Medium Zooplankton MED_ZOO  
17) Large Zooplankton LG_ZOO  
18) Jellyfish JELLIES (in the past represented as LG_ZOO)  
Large pelagic invetebrates  
19)Cephalopod CEP  
20) Prawns PWN  
Infauna  
21) Small Infauna SM_INF  
22) Large Infauna LG_INF  
Epibenthic organisms  
23) Sediment epibenthic filter feeders SED_EP_FF (often used for bivalves, sponges)  
24) Benthic grazers SED_EP_OTHER  
25) Mobile epibenthos MOB_EP_OTHER (often used for crabs, lobster, octopus)   
26) Corals CORAL  
27)Sponges SPONGE  
Bacteria   
28) Pelagic Bacteria PL_BACT  
29) Sediment Bacteria SED_BACT    
Obligatory detritus groups   
30) Carrion CARRION   
31) Labile detritus LAB_DET   
32) Refractory detritus REF_DET   
33) Additional ice and land based groups are also available.   
Ice dwelling   
34) ICE_BACT   
35) ICE_MIXOTROPHS   
36) ICE_DIATOMS   
37) ICE_ZOOBIOTA   
Land dwelling (vegetation only for now)    
38) MARSH   
39) MANGROVE

For NOBA Atlantis, these are the biomass pools:

```{r}

fgs <- atlantisom::load_fgs(here("atlantisoutput/NOBA_sacc_30"), "nordic_groups_v04.csv")

pools <- fgs %>%
  dplyr::filter(NumCohorts==1) %>%
  dplyr::select(Code, Name, Long.Name, NumCohorts, InvertType)

knitr::kable(pools)

```

Here are the relevant outputs in the output.nc file for these groups:

```{r}
# return everything in var_names_ncdf matching Name in pools

var_names_ncdf[str_detect(var_names_ncdf, str_c(pools$Name, collapse ="|"))]

```
So we want "N" nitrogen output. In theory this can be achieved with `atlantisom::load_nc`with `select_variable = "N"`. Looks like it works, output is N for our selected biomass pool species over all polygons, layers, and times. Output of `unique(testNpools$species)`:

```{r}

select_groups <- pools$Name
nc_out <- file_nc
bps <- load_bps(dir = dir, fgs = "nordic_groups_v04.csv", file_init = "nordic_biol_v23.nc")
  # Get the boundary boxes
  allboxes <- load_box(dir = dir, file_bgm = "Nordic02.bgm")
  boxes <- get_boundary(allboxes)

testNpools <- load_nc(dir = dir,
                  file_nc = nc_out,
                  bps = bps,
                  fgs = fgs,
                  select_groups = select_groups,
                  select_variable = "N",
                  check_acronyms = TRUE,
                  bboxes = boxes)

unique(testNpools$species)
```

I don't think we have a function that aggregates over layers and converts N to biomass but there are pieces in other functions.

New `atlantisom` function: `calc_biomass_pool` based on `calc_biomass_age`, but adding the expansion to volume used in `calc_pred_cons`.

```{r}

# we need the volume of each layer as input to this
vol <- load_nc_physics(dir = dir,
                         file_nc = nc_out,
                         physic_variables = "volume",
                         aggregate_layers = FALSE,
                         bboxes = boxes)

# we need area for the benthic inverts instead of volume
# THANK YOU JOE for this tip
area <- load_boxarea(dir = dir,
                     file_bgm = "Nordic02.bgm")

# we need a list of which group types to expand to area instead of volume
benthos <- c("SED_EP_FF", "SED_EP_OTHER", "SED_BAC", 
             "MOB_EP_OTHER", "LG_INF", "SM_INF", 
             "MICROPHYTBENTHOS", "SEAGRASS")

calc_biomass_pool <- function(pooln, vol, area, fgs, biolprm){
  datalist <- list(pooln)

  # Conversion factor from mg N to t wet-weight
  # should only use conversion for non vertebrates
  # also need volume of cell info
  bio_conv <- biolprm$redfieldcn * biolprm$kgw2d / 1000000000

  # get fgs info
    # use grouptype column to allocate
  colnames(fgs) <- tolower(colnames(fgs))

  # check for GroupType or InvertType
  if (!("grouptype" %in% colnames(fgs) | "inverttype"%in% colnames(fgs))) {
    stop(paste("The columns GroupType or InvertType ars not in your functional\n",
               "groups file."))
  }

  # change inverttype to grouptype, contents should be the same
  if("inverttype" %in% names(fgs)) names(fgs)[names(fgs) == 'inverttype'] <- 'grouptype'

  fgs$grouptype <- tolower(fgs$grouptype)

  pooltype <- fgs |>
    dplyr::select(species=name, grouptype) |>
    dplyr::mutate(pooltype = dplyr::case_when((grouptype %in% c("lg_inf",
                                                                "sm_inf",
                                                                "sed_bact",
                                                                "sed_ep_ff",
                                                                "sed_ep_other",
                                                                "mobile_ep_other",
                                                                "coral",
                                                                "sponge")) ~ "benthos",
                                              TRUE ~ "nonbenth"))

  data_names <- c("species", "agecl", "polygon", "layer", "time", "atoutput")

  if (all(sapply(datalist, function(x) all(is.element(names(x), data_names))))){

    names(vol)[names(vol) == "atoutput"] <- "vol"

    pooln <- merge(pooln, vol,
                   by = c("time", "polygon", "layer"))

    pooln <- merge(pooln, area,
                   by = c("polygon"))

    pooln <- merge(pooln, pooltype,
                   by = c("species"))

    #pooln$atoutput <- with(pooln, vol * atoutput * bio_conv)

    pooln <- pooln |>
      dplyr::mutate(atoutput = dplyr::case_when(pooltype == "nonbenth" ~ vol * atoutput * bio_conv,
                                                pooltype == "benthos" ~ 
                                                  ifelse(layer==7, area * atoutput * bio_conv, 0)))

    # Sum over layers 
    biomass_pools <- aggregate(atoutput ~ species + agecl + time + polygon,
      data = pooln, sum)
  } else {
    stop(paste("Dataframe names do not match with", data_names))
  }

  return(biomass_pools)
}

biolprm <- load_biolprm(dir = dir, file_biolprm = "nordic_biol_incl_harv_v_011_1skg.prm")

biomass_pools <- calc_biomass_pool(pooln = testNpools,
                                   vol = vol,
                                   area = area,
                                   fgs = fgs,
                                   biolprm = biolprm)

```

Do the new biomass pools match the `BiomIndx.txt` output when aggregated over polygon? If so they can be the input to the survey function for time and spatial subsetting.

```{r, fig.width=10}

aggbiopools <- biomass_pools %>%
  dplyr::group_by(species, agecl, time) %>%
  dplyr::summarise(totpool = sum(atoutput))

txtbiopools <- atlantisom::load_bioind(dir, "nordic_runresults_01BiomIndx.txt", fgs) %>%
  dplyr::mutate(time = time/73) %>%
  dplyr::filter(species %in% unique(aggbiopools$species)) %>%
  dplyr::select(species, time, atoutput)

ggplot(txtbiopools, aes(x=time, y=atoutput)) +
  geom_line() +
  geom_point(data = aggbiopools, aes(x = time, y = totpool), colour = "blue", alpha=0.1) +
  theme_bw() +
  facet_wrap(~species, scales = "free_y")

```

## Diagnose mismatches between biopools and BiomInd.txt output

Who matches exactly (enough). Patterns in absolute mismatch are interesting, some species are really close and some always way off. Volume wrong, depth layers wrong? Mine are always higher than the txt output when they are off, so maybe too much expansion somewhere.

Joe pointed out that benthic groups should be expanded to area rather than volume. That correction has now been applied, and matches are better, but still diverge over time for some groups.

Groups that match seem more likely to be pelagic (dinoflagellates, all zooplankton, pelagic bacteria, labile and refractive detritus) although large phytoplankton is way off.

```{r, warning=FALSE, fig.width=10}
mismatch <- txtbiopools |>
  dplyr::left_join(aggbiopools) |>
  dplyr::mutate(mismatch = atoutput - totpool,
                bad = ifelse(abs(mismatch) > 0.01*(atoutput), 1, 0))

ggplot(mismatch, aes(x=time, y=mismatch)) +
  geom_line() +
  theme_bw() +
  facet_wrap(~species, scales = 'free_y')
```

This is mismatch coded as 0 if less than 1% of txt output, 1 if more. Those exceeding are carrion, predatory benthos, detritivore benthos, king crab, benthic bacteria, large phytoplankton and sometimes small zooplankton.

```{r, warning=FALSE}

ggplot(mismatch, aes(x=time, y=bad)) +
  geom_line() +
  theme_bw() +
  facet_wrap(~species)
```

What is up with the phytoplankton groups? Small matches well visually but large doesn't. Large phytoplankton is the largest magnitude mismatch across all groups.

```{r}
ggplot(txtbiopools |> filter(species %in% c("Large_phytop", "Small_phytop")), 
       aes(x=time, y=atoutput)) +
  geom_line() +
  geom_point(data = aggbiopools |> filter(species %in% c("Large_phytop", "Small_phytop")), 
             aes(x = time, y = totpool), colour = "blue") +
  theme_bw() +
  facet_wrap(~species, scales = "free_y")

```

Predatory benthos is off by what appears to be a scalar, Detrivore ben converges as they increase.

```{r}
ggplot(txtbiopools |> filter(species %in% c("Predatory_ben", "Detrivore_ben")), 
       aes(x=time, y=atoutput)) +
  geom_line() +
  geom_point(data = aggbiopools |> filter(species %in% c("Predatory_ben", "Detrivore_ben")), 
             aes(x = time, y = totpool), colour = "blue", alpha=0.1) +
  theme_bw() +
  facet_wrap(~species, scales = "free_y")

```

## Remaining changes to workflow, once corrected

### Add `calc_biomass_pool` to `run_truth`

### Test wrapper functions with biomass pool groups